---
title: "Figuremaking_New12May2026"
author: "Anthony Fischer"
date: "2026-05-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load mpra_tools.R and additional libraries, badAlt, and set theme

```{r}
source("../../code/mpra_tools/mpra_tools.R")
library(umap)
library(RColorBrewer)
library(ggforce)
library(fplot)
library(pdftools)
library(gkmSVM)
library(ggbreak)
library(xtable)
library(knitr)
library(readxl)
library(openxlsx)
library(ggrain)
library(tidyverse)
library(ggrastr)
cbPalette <- brewer.pal(8,"Dark2")
#A list of elements we do not consider from Library 1.0 as they have bad alts or something
badAlt<-c('ZNF626_wgs','ASIC2_wgs', 'CDKN2B_wgs', 'ELMO1_wgs', 'ENPP5_wgs', 'GDPD4_ssc', 'INSR_wgs', 'LMBRD2_wgs', 'LRP2BP_wgs', 'SCARB2.1_wgs', 'TLK1_wgs', 'TRIM37_wgs', 'ZCCHC11_wgs')
#Set the theme for all ggplots
theme_set(
  theme_classic2(base_family = "sans")+
theme(plot.title=element_text(hjust=0.5, size = 14, face = "bold"),
      plot.subtitle = element_text(hjust=0.5, size = 12, face = "bold"),
  axis.title = element_text(size = 12),
  axis.text = element_text(size = 10),
  legend.title = element_text(size = 14, face = "bold",hjust=0.5),
  legend.text = element_text(size = 10, face = "bold"),
  panel.border = element_rect(color = "black", fill = NA, linewidth = 1)))
cbPalette <- brewer.pal(8,"Dark2")
```

## Explanation of New ActD Assay

```{r New ISOMPRA Data 12May2026}
# As part of the review process, we performed an additional screen of the variants
# in an ActinomycinD experiment, and made new pre- and post-splicing libraries
# to compare them to to not run into batch effects from the older experiment.
# We decided to replace the figure panels for the cell culture experiments based
# on this new data, as the old pre-splice data was pulled from a TRAP experiment
# where we didn't save the nuclear pellet, so these pre-splice levels are presumably
# whatever wound up in the TRAP input supernatant - not ideal.  This section
# will specifically deal with remaking those figure panels (now Figure 4) by
#recycling code from figuremaking.Rmd
```

```{r Figure Panels}
# We are going to separate by figure panels that I have identified need to be fixed
# for the resubmission.  We will place them in order here, but this is not the order
# I will create them (some panels are easier than others to recreate)
# List of panels/tables that need to be remade can be found at
# https://docs.google.com/spreadsheets/d/1E8q9C-_C8j2ejS-qITyMVZh_kqwEg37Ju8EMv59rqpQ/edit
################################################################################
```

## Figure 2 Panels
### Figure 2L - New Batch Corrected MDS Plot
#### Not Including in Paper, see original figuremaking.Rmd for updated plot

```{r Figure 2L}


################################################################################
#Panel 2L - UMAP Plot
# New ActD data
d <- fread("../../data/formatted/ActD_Trials_08Apr2026/formatted.txt") %>% filter(,!Element %in% badAlt)

#Separate allele names
d <- d %>%
  separate_wider_regex(
    cols = Element,
    patterns = c(
      Element = ".*",        # Match as much as possible (greedy)
      "_",                   # Match the last underscore
      Allele = "[^_]+"       # Match everything remaining that isn't an underscore
    )
  )
d <- as.data.table(d)

d_old <- fread("../../data/formatted/Library1and2BCLess_ISOMPRA_6-30-2022/formatted_Reads_compiled_06-30-2022.tsv")[,c(2,3,5,6,9,10)] %>% filter(Comment %in% c("Input Splicing Agnostic mRNA","TRAP Exonic (mature) mRNA","Plasmid DNA"))
d_tmp <- filter(d,Fraction=="0hr")[,!c(3,4)]
setnames(d_tmp,4,"Comment")
d_tmp$Comment[which(d_tmp$Comment=="pre")] <- "Input Intronic (pre) mRNA"
d_tmp$Comment[which(d_tmp$Comment=="post")] <- "Input Exonic (mature) mRNA"
d_tmp$TechRep <- 1
umap_data <- bind_rows(
  d_old,
  d_tmp
)
#Add a pseudocount to CPM
umap_data$CPM <- umap_data$CPM + 1

umap_CPM <- as.data.table(filter(umap_data,
                   Comment %in% c("Input Intronic (pre) mRNA",
                                  "Input Exonic (mature) mRNA",
                                  "Input Splicing Agnostic mRNA",
                                  "Plasmid DNA",
                                  "TRAP Exonic (mature) mRNA")) %>% 
  pivot_wider(id_cols = c(Element,Allele),values_from = "CPM",names_from = c(Comment,BiolRep,TechRep)))

umap_results <- as.data.table(umap(select(umap_CPM,!1:2))$data,keep.rownames = TRUE) %>% tidyr::separate(rn,c("Library","BiolRep"),sep="_")
umap_results$Library <- factor(umap_results$Library,levels = c("Input Intronic (pre) mRNA",
                                                               "Input Exonic (mature) mRNA",
                                                               "Input Splicing Agnostic mRNA",
                                                               "TRAP Exonic (mature) mRNA",
                                                               "Plasmid DNA"))
p1 <- ggplot(umap_results,aes(x=X1,y=X2,fill=Library))+
  geom_point(shape=21,size=1)+
  geom_polygon(stat = "ellipse",aes(fill=Library),color="black",alpha=0.5)+
  scale_fill_manual(values=cbPalette)+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank())+
  labs(title="UMAP Analysis\nof Library CPM")+
  theme(plot.title = element_text())
print(p1)

export_graph_start("./Figures for Paper/Figure1/UMAP Plot.pdf",width=0.75)
print(p1)
export_graph_end()

#Oooh, that changed it quite a bit.  I think we should try something different.
#UMAP may be inappropriate for this analysis
#limma and edgeR packages have plotMDS functions that should work here, they
#want log2 CPM data

BioBlock <- c(
  paste0("RNA_Rep", 1:8), # Pre
  paste0("RNA_Rep", 1:8), # Mature
  paste0("RNA_Rep", 1:8), # Agnostic
  paste0("DNA_Tech", 1:8), # Plasmid DNA (unpaired background)
  paste0("RNA_Rep", 1:7)  # TRAP (missing rep 8)
)

corrected <- removeBatchEffect(umap_CPM[,-c(1,2)],batch = BioBlock)

mds_res <- limma::plotMDS(umap_CPM_wide)

mds_df <- data.frame(
  SampleID = colnames(umap_CPM_wide),
  Dim1 = mds_res$x,
  Dim2 = mds_res$y
)
mds_df <- tidyr::separate(mds_df,SampleID,c("Library","BiolRep","TechRep"),sep="_")

# Calculate the total variance
total_var <- sum(mds_res$eigen.values)

# Calculate percentage for Dim 1 and Dim 2
var_pc1 <- (mds_res$eigen.values[1] / total_var) * 100
var_pc2 <- (mds_res$eigen.values[2] / total_var) * 100

export_graph_start("New MDS Plot CPM.pdf")
ggplot(mds_df,aes(x=Dim1,y=Dim2,fill=Library))+geom_point(shape=21,size=3)+
  geom_polygon(stat = "ellipse",alpha=0.2,color="black")+
  labs(x=paste0("Leading logFC dim 1 (",round(var_pc1,0),"%)"),
       y=paste0("Leading logFC dim 2 (",round(var_pc2,0),"%)"))
export_graph_end()

# Talking with Joe, we are most certainly seeing batch effects here.  Since the
# numbers didn't come from the same RNA, it's inappropriate to display it as seen
# here.  Joe thinks we should make separate MDS plots, keep the old (change to
# limma MDS plot?), and generate a new one here.

#Add a pseudocount to CPM
umap_data <- as.data.table(d)
umap_data$CPM <- umap_data$CPM + 1
umap_data[,Fraction := factor(Fraction,
                              levels = c("0hr","5hr","11hr"),
                              labels = c("t0","t5","t11"))]
umap_data[,Splicing := factor(Splicing,
                              levels=c("pre","post"),
                              labels=c("Pre-Splicing","Post-Splicing"))]


umap_CPM_wide <- umap_data %>% pivot_wider(id_cols = c(Element,Allele),values_from = "CPM",names_from = c(Fraction,Splicing,BiolRep)) %>% as.data.table()

mds_res <- limma::plotMDS(umap_CPM_wide)

mds_df <- data.frame(
  SampleID = colnames(umap_CPM_wide),
  Dim1 = mds_res$x,
  Dim2 = mds_res$y
)
mds_df <- tidyr::separate(mds_df,SampleID,c("Time","Splicing","BiolRep"),sep="_")

# Calculate the total variance
total_var <- sum(mds_res$eigen.values)

# Calculate percentage for Dim 1 and Dim 2
var_pc1 <- (mds_res$eigen.values[1] / total_var) * 100
var_pc2 <- (mds_res$eigen.values[2] / total_var) * 100

export_graph_start("./Figures for Paper/Figure1/New MDS Plot CPM Only New Data.pdf")
ggplot(mds_df,aes(x=Dim1,y=Dim2,fill=interaction(Time,Splicing)))+geom_point(shape=21,size=3)+
  geom_polygon(stat = "ellipse",alpha=0.1,color="black")+
  labs(x=paste0("Leading logFC dim 1 (",round(var_pc1,0),"%)"),
       y=paste0("Leading logFC dim 2 (",round(var_pc2,0),"%)"))+
  labs(fill="Libary")+
  geom_text(aes(label=BiolRep),size=2)
export_graph_end()


#I want to see everything, with corrected batch effects
d_tmp <- d
d_old <- fread("../../data/formatted/Library1and2BCLess_ISOMPRA_6-30-2022/formatted_Reads_compiled_06-30-2022.tsv")[,c(2,3,5,6,9,10)]
setnames(d_tmp,4,"Comment")
d_tmp$Comment[which(d_tmp$Comment=="0hr" & d_tmp$Splicing=="pre")] <- "Zero Hour Intronic (pre) mRNA"
d_tmp$Comment[which(d_tmp$Comment=="0hr" & d_tmp$Splicing=="post")] <- "Zero Hour Exonic (mature) mRNA"
d_tmp$Comment[which(d_tmp$Comment=="5hr" & d_tmp$Splicing=="post")] <- "Five Hour Exonic (mature) mRNA"
d_tmp$Comment[which(d_tmp$Comment=="11hr" & d_tmp$Splicing=="post")] <- "Eleven Hour Exonic (mature) mRNA"

d_tmp$TechRep <- 1
d_all <- as.data.table(bind_rows(d_tmp[,-c(3,6)],d_old))
d_all[,Comment := factor(Comment,levels = c(
  "Input Intronic (pre) mRNA",
  "Input Exonic (mature) mRNA",
  "Input Splicing Agnostic mRNA",
  "TRAP Exonic (mature) mRNA",
  "TRAP Splicing Agnostic mRNA",
  "Plasmid DNA",
  "Zero Hour Intronic (pre) mRNA",
  "Zero Hour Exonic (mature) mRNA",
  "Five Hour Exonic (mature) mRNA",
  "Eleven Hour Exonic (mature) mRNA"))]
setorderv(d_all,c("Comment","BiolRep","TechRep"))

umap_CPM <- pivot_wider(d_all,id_cols = c(Element,Allele),values_from = "CPM",names_from = c(Comment,BiolRep,TechRep))
colnames(umap_CPM)

umap_CPM[is.na.data.frame(umap_CPM)] <- 0
umap_CPM <- umap_CPM[,-c(1:2)]+1
BioBlock <- c(
  paste0("RNA_Rep", 1:8), # Pre
  paste0("RNA_Rep", 1:8), # Mature
  paste0("RNA_Rep", 1:8), # Agnostic
  paste0("RNA_Rep", 1:7),  # TRAP mature (missing rep 8)
  paste0("RNA_Rep", 1:7),  # TRAP SA (missing rep 8)
  paste0("DNA_Tech", 1:8), # Plasmid DNA (unpaired background)
  paste0("RNA_Rep_0hr", 1:6),
  paste0("RNA_Rep_0hr", 1:6),
  paste0("RNA_Rep_5hr", 1:6),
  paste0("RNA_Rep_11hr", 1:6)
)

all_corrected <- removeBatchEffect(umap_CPM,BioBlock)

mds_res_all <- limma::plotMDS(all_corrected,gene.selection = "common")

mds_df_all <- data.frame(
  SampleID = colnames(umap_CPM),
  Dim1 = mds_res_all$x,
  Dim2 = mds_res_all$y
)

mds_df_all <- tidyr::separate(mds_df_all,SampleID,c("Library","BiolRep","TechRep"),sep="_")

# Calculate the total variance
total_var <- sum(mds_res_all$eigen.values)

# Calculate percentage for Dim 1 and Dim 2
var_pc1 <- (mds_res_all$eigen.values[1] / total_var) * 100
var_pc2 <- (mds_res_all$eigen.values[2] / total_var) * 100

#Change the Plasmid DNA title
mds_df_all$Library[mds_df_all$Library=="Plasmid DNA"] <- "Plasmid DNA (Technical Replicates)"
mds_df_all <- as.data.table(mds_df_all)
mds_df_all[,Library := factor(Library,levels = c(
  "Input Intronic (pre) mRNA",
  "Input Exonic (mature) mRNA",
  "Input Splicing Agnostic mRNA",
  "TRAP Exonic (mature) mRNA",
  "TRAP Splicing Agnostic mRNA",
  "Plasmid DNA (Technical Replicates)",
  "Zero Hour Intronic (pre) mRNA",
  "Zero Hour Exonic (mature) mRNA",
  "Five Hour Exonic (mature) mRNA",
  "Eleven Hour Exonic (mature) mRNA"))]
setorderv(mds_df_all,c("Library","BiolRep","TechRep"))


ggplot(filter(mds_df_all,!Library %in% c("Five Hour Exonic (mature) mRNA","Eleven Hour Exonic (mature) mRNA")),
       aes(x=Dim1,y=Dim2,fill=Library))+geom_point(shape=21,size=3)+
  geom_polygon(stat = "ellipse",alpha=0.2,color="black")+
  labs(x=paste0("Leading logFC dim 1 (",round(var_pc1,0),"%)"),
       y=paste0("Leading logFC dim 2 (",round(var_pc2,0),"%)"))+
  scale_fill_manual(values = c("red","orange","yellow","green","cyan","blue","purple","brown","white","black"))

#That doesn't really help much, I don't think we should worry about it, the ActD
#throws a huge variable in there in the form of a dying library, don't worry about it for right now

```

### Figure 2M

```{r Figure-2M}
#Panel 2M - Raincloud Plots

#New Raincloud plot for stability sequences
d <- fread("../../data/formatted/ActD_Trials_08Apr2026/formatted.txt") %>% filter(,!Element %in% badAlt)

#Separate allele names
d <- d %>%
  separate_wider_regex(
    cols = Element,
    patterns = c(
      Element = ".*",        # Match as much as possible (greedy)
      "_",                   # Match the last underscore
      Allele = "[^_]+"       # Match everything remaining that isn't an underscore
    )
  )
d <- as.data.table(d)

#Grab the DNA info
dna <- fread("../../data/formatted/Library1and2BCLess_ISOMPRA_6-30-2022/formatted_Reads_compiled_06-30-2022.tsv") %>% filter(Fraction=="DNA")
dna <- dna %>% group_by(Element,Allele) %>% summarize(CPM=mean(CPM)) %>% as.data.table()
#Correct the controls
dna$Allele[dna$Element=="PREx4"] <- "ref"
dna$Element[dna$Element=="PREx4"] <- "PREx4_dnvctl"
dna$Allele[dna$Element=="SREx4"] <- "ref"
dna$Element[dna$Element=="SREx4"] <- "SREx4_dnvctl"
dna$Allele[dna$Element=="WPRE-alpha"] <- "ref"
dna$Element[dna$Element=="WPRE-alpha"] <- "WPRE-alpha_dnvctl"

#Put the DNA info in the table, adjoining to correct element
d <- as.data.table(left_join(d,dna,by=c("Element","Allele")))
#Calculate expression
d[, Expression := log2(CPM.x/CPM.y)]
#Get rid of non-finite values
d <- filter(d,is.finite(Expression))

d_stab <- inner_join(
  filter(d,Fraction=="0hr",Splicing=="pre"),
  filter(d,Fraction=="0hr",Splicing=="post"),
  by=c("Element","Allele","BiolRep")) %>% mutate(Stability=Expression.y-Expression.x)

View(bind_rows(
    filter(d_stab,Element %in% c("PREx4_dnvctl","SREx4_dnvctl"),is.finite(Stability)),
    filter(d_stab,Allele=="shuf",is.finite(Stability))%>%mutate(Element="Shuffled Elements"))
  %>%mutate(Element=factor(Element,levels=c("Shuffled Elements","PREx4_dnvctl","SREx4_dnvctl"),labels=c("Shuffled Elements","Pumilio Response Element","Smaug Response Element"))))

p2 <- ggplot(
  bind_rows(
    filter(d_stab,Element %in% c("PREx4_dnvctl","SREx4_dnvctl"),is.finite(Stability)),
    filter(d_stab,Allele=="shuf",is.finite(Stability))%>%mutate(Element="Shuffled Elements"))
  %>%mutate(Element=factor(Element,levels=c("Shuffled Elements","PREx4_dnvctl","SREx4_dnvctl"),labels=c("Shuffled Elements","Pumilio Response Element","Smaug Response Element"))),
    aes(x=Element,y=Stability,fill=Element))+
  geom_rain(point.args = list(shape=21, alpha = 0.2),
            violin.args.pos = list(scale="width",width=1.5,position=position_nudge(x=0.15)),
            rain.side = "r")+
  geom_signif(comparisons = list(
    c("Pumilio Response Element","Shuffled Elements"),
    c("Smaug Response Element","Shuffled Elements")
    ),map_signif_level = TRUE,test = "t.test",y_position = c(10,11),tip_length = 0)+
  labs(title="Stability of Control Sequences",y="Stability\n(Post-Splicing Expression\nMinus Pre-Splicing Expression)")+
  theme(plot.title = element_text(hjust=0.5),
        legend.position = "none",
        axis.title.x = element_blank())+
  #scale_y_continuous(limits = c(-2.5,2.7))+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 13))+
  scale_fill_manual(values=c("#454545","#808080","#bababa"))
print(p2)

p2 <- ggplot(
  rbind(
    filter(d,Allele=="dnvctl",Element!="QREx4",is.finite(Stability)),
    filter(d,Allele=="shuf",is.finite(Stability))%>%mutate(Element="Shuffled Elements"))
  %>%mutate(Element=factor(Element,levels=c("Shuffled Elements","PREx4","SREx4","WPRE-alpha"),labels=c("Shuffled","PRE","SRE","WPRE"))),
    aes(x=Element,y=Stability,fill=Element))+
  geom_rain(point.args = list(shape=21, alpha = 0.2))+
  geom_signif(comparisons = list(
    c("PRE","Shuffled"),
    c("SRE","Shuffled"),
    c("WPRE","Shuffled")),map_signif_level = TRUE,test = "t.test",y_position = c(1.5,1.8,2.1),tip_length = 0)+
  labs(title="Stability of Control Sequences",y="Stability\n(Post-Splicing Expression\nMinus Pre-Splicing Expression)")+
  theme(plot.title = element_text(hjust=0.5),legend.position = "none")+
  scale_y_continuous(limits = c(-2.5,2.7))+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 13))+
  scale_fill_manual(values=c("#454545","#808080","#bababa","#f5f5f5"))

print(p2)

export_graph_start("./Figures for Paper/Figure1/Control Stability - RainCloud_New.pdf",sideways = FALSE,width = 1.5)
print(p2+coord_fixed(ratio = .15))
export_graph_end()


# Oooh, now PRE and SRE are higher than the average shuffled.  I don't think this
# is a particularly infrmative graph, talk to Joe about removing.
# Joe suggested removal
```

## Figure 4 Panels

```{r Figure 4}
# Panel 4A, 4B, 4C - Volcano Plots RefShuf
# We will recycle data for 4D as this came from the TRAP and no new data has been generated
intron_shuf <- fread("../ActD Trials/Full_Library_08Apr2026/Pre-Splicing_RefvsShuf_Welchs.tsv") %>% .[,plotcolor := as.factor(case_when(
    (abs(logFC)>log2(1.25) & fdr<0.1) ~ 1,
    TRUE ~ 2))]

exon_shuf <- fread("../ActD Trials/Full_Library_08Apr2026/Post-Splicing_RefvsShuf_Welchs.tsv") %>% .[,plotcolor := as.factor(case_when(
    (abs(logFC)>log2(1.25) & fdr<0.1) ~ 1,
    TRUE ~ 2))]

stability_shuf <- fread("../ActD Trials/Full_Library_08Apr2026/Stability_RefvsShuf_Welchs.tsv") %>%
  .[,plotcolor := as.factor(case_when(
    (abs(logFC)>log2(1.25) & fdr<0.1) ~ 1,
    TRUE ~ 2))]

translational_shuf <- fread("./Translation_Splicing-Agnostic_RefvsShuf_Welchs.tsv") %>% .[,plotcolor := as.factor(case_when(
    (abs(logFC)>log2(1.25) & fdr<0.1) ~ 1,
    TRUE ~ 2))]

p1_shuf <- ggplot(intron_shuf)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor,alpha = plotcolor),shape=21,size=2)+
  scale_fill_manual(values=c(cbPalette[1],"lightgray"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  scale_alpha_manual(values = c(1,1,0.1,0.1))+
  labs(fill="Expression Change",
       title="Reference vs. Shuffled Expression Change\nin Pre-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.05),linetype="dashed")+
  theme(legend.position = "none",aspect.ratio=1)

p2_shuf <- ggplot(exon_shuf)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor),shape=21,size=2)+
  scale_fill_manual(values=c(cbPalette[1],"lightgray"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  labs(fill="Expression Change",
       title="Reference vs. Shuffled Expression Change\nin Post-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.05),linetype="dashed")+
  theme(legend.position = "none",aspect.ratio=1)

p3_shuf <- ggplot(stability_shuf)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor),shape=21,size=2)+
  scale_fill_manual(values=c(cbPalette[1],"lightgrey"),
                    labels=c("Increased Stability","Decreased Stability",
                             "Not Significant"))+
  labs(fill="Stability Change",
       title="Reference vs. Shuffled Stability Change",
       x="Log2 Fold Change in Stability",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.05),linetype="dashed")+
  theme(legend.position = "none",aspect.ratio=1)

p4_shuf <- ggplot(translational_shuf)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor),shape=21,size=2)+
  scale_fill_manual(values=c(cbPalette[1],"lightgrey"),
                    labels=c("Increased Translation","Decreased Translation",
                             "Not Significant"))+
  labs(fill="Translational Change",
       title="Reference vs. Shuffled Translational Change",
       x="Log2 Fold Change in Translation",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.05),linetype="dashed")+
  theme(legend.position = "none",aspect.ratio=1)

export_graph_start("./Figures for Paper/Figure3/AllRefvsShuf_VolcanoPlot_Patchwork_New.pdf",sideways = TRUE,height = 0.5)
print(p1_shuf+theme(legend.position = "none")+labs(title=NULL,subtitle="Pre-Splicing",x="Log2FC Expression")+
        p2_shuf+theme(legend.position = "none")+labs(title=NULL,subtitle="Post-Splicing",x="Log2FC Expression")+
        p3_shuf+theme(legend.position = "none")+labs(title=NULL,subtitle="Stability",x="Log2FC Stability")+
        p4_shuf+theme(legend.position = "none")+labs(title=NULL,subtitle="Translational",x="Log2FC TE")+
        plot_layout(ncol=4)+plot_annotation(title="Reference vs. Shuffled"))
export_graph_end()
################################################################################
#Figure 4F - Run 4G code first
exon_stability <- left_join(exon_stability,plottable)
exon_stability <- unique.data.frame(exon_stability)
p5_shuf_dotplot <- ggplot(exon_stability,aes(x=logFC.x,y=logFC.y,fill=Category))+
  geom_point(shape=21,alpha=0.5,size=3)+
  scale_fill_manual(values = c("Not Significant" = "lightgrey",
                               "True Positive" = "#136d52",
                               "False Positive" = "#21c494",
                               "False Negative" = "#67e5bf"))+
  labs(x="Log Fold Change\n(Post-Splice)",
       y="Log Fold Change\n(Stability)",
       )+
  theme_bw()+
  theme(legend.position = "none",aspect.ratio = 1)
print(p5_shuf_dotplot)

export_graph_start("./Figures for Paper/Figure3/logfcpostvsstab_new_refshuf.pdf")
print(p5_shuf_dotplot + coord_fixed(ratio=1.979))
export_graph_end()
################################################################################
#Figure 4G
#False positive analysis
exon_stability <- inner_join(exon_shuf,stability_shuf,"Element")

#Significant in both and same direction
true_positives <- filter(exon_stability,
                         fdr.x<0.1,
                         fdr.y<0.1,
                         ((logFC.x>0 & logFC.y>0) | (logFC.x<0 & logFC.y<0))
                         )$Element
#Only significant in stability, not post-splicing
false_negatives <- filter(exon_stability,
                         fdr.x>0.1,
                         fdr.y<0.1)$Element
#Either only significant in post-splicing, or significant in both but opposite
#directions
false_positives <- c(
  filter(exon_stability,
         fdr.x<0.1,
         fdr.y>0.1)$Element,
  filter(exon_stability,
         fdr.x<0.1,
         fdr.y<0.1,
         ((logFC.x>0 & logFC.y<0) | (logFC.x<0 & logFC.y>0)))$Element)
  




all_else <- filter(stability_shuf,
                   !Element %in% false_positives,
                   !Element %in% false_negatives,
                   !Element %in% true_positives)$Element

plottable <- rbind(
  data.table("Element"=true_positives,"Category"="True Positive"),
  data.table("Element"=false_positives,"Category"="False Positive"),
  data.table("Element"=false_negatives,"Category"="False Negative"),
  data.table("Element"=all_else,"Category"="Not Significant")
)

plottable[,Category := factor(Category,levels=c("True Positive","False Positive","False Negative","Not Significant"))]

summary_data <- plottable %>%
  group_by(Category) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

# Separate the data for labels
label_data <- summary_data %>%
  filter(Category != "Not Significant") %>%
  mutate(cumulative_percentage = 100 - cumsum(percentage) + percentage/2)

# Create the plot

p5_shuf <- ggplot(summary_data, aes(x = "", y = percentage, fill = Category)) +
  geom_col(width = 0.4, color = "black") +
  scale_fill_manual(values = c("Not Significant" = "lightgrey",
                               "True Positive" = "#136d52",
                               "False Positive" = "#21c494",
                               "False Negative" = "#67e5bf")) +
  labs(y = "Percentage",
       fill = "Category",
       x=NULL) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        legend.position = "right",
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

export_graph_start("./Figures for Paper/Figure3/RefShuf_Accuracy_New.pdf",width=0.75)
print(p5_shuf+theme_minimal())
ggplot(summary_data,aes(Category,count,fill=as.factor(count)))+
  geom_col()+
  geom_text(aes(y=count-100,label=paste0(round(percentage,2),"\n",count)))+
  theme(axis.text.x = element_text(angle = 45,hjust=1))
export_graph_end()

#Keep the exon_stability and plottable for Figure S6 down below

################################################################################
# Panel 4H, 4I, 4J - Volcano Plots RefAlt
# We will recycle data for 4K as this came from the TRAP and no new data has been generated

intron <- fread("../ActD Trials/Full_Library_08Apr2026/PreSplicing_RefAlt_LM.txt") %>% .[,plotcolor := as.factor(case_when(
    (abs(logFC)>log2(1.25) & fdr<0.1) ~ 1,
    TRUE ~ 2))]

exon <- fread("../ActD Trials/Full_Library_08Apr2026/PostSplicing_RefAlt_LM.txt") %>% .[,plotcolor := as.factor(case_when(
    (abs(logFC)>log2(1.25) & fdr<0.1) ~ 1,
    TRUE ~ 2))]

stability <- fread("../ActD Trials/Full_Library_08Apr2026/Stability_RefAlt_LM.txt") %>%
  mutate(logFC_interaction = -(logFC_interaction)) %>%
  .[,plotcolor := as.factor(case_when(
    (abs(logFC_interaction)>log2(1.25) & fdr_interaction<0.1) ~ 1,
    TRUE ~ 2))]

translational <- fread("./Translation_Splicing-Agnostic_RefvsAlt_LM_Annotated.tsv") %>% .[,plotcolor := as.factor(case_when(
    (abs(logFC)>log2(1.25) & fdr<0.1) ~ 1,
    TRUE ~ 2))]

# Define some points we want to label on the plots
# Specifically, intron points that are transcriptional only,
# exon points that are a mix of transcriptional only and stability sig
# all or nearly all the stability points
# all the translational points

intron_elems <- filter(stability, fdr_allele < 0.1, fdr_interaction > 0.1) %>% 
  as.data.table() %>% 
  .[order(abs(fdr_allele))] %>% 
  head(4) %>% pull(Element)

exon_elems <- filter(stability, fdr_interaction < 0.1) %>% 
  as.data.table() %>% 
  .[order(abs(fdr_allele))] %>% 
  head(4) %>% pull(Element)

stability_elems <- filter(stability,fdr_interaction<0.1) %>% as.data.table() %>%
  .[order(abs(fdr_interaction))] %>% head(10) %>% pull(Element)

translational_elems <- filter(translational,fdr<0.1) %>% as.data.table() %>%
  .[order(-abs(logFC))] %>% head(10) %>% pull(Element)

p1 <- ggplot(intron)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor,shape=Pheno),size=2)+
  scale_fill_manual(values=c(cbPalette[1],"lightgray"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  scale_shape_manual(values = c(21,23))+
  labs(fill="Expression Change",
       title="Reference vs. Variant Expression Change\nin Pre-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")+
  theme(legend.position = "none",aspect.ratio = 1)+
  geom_text_repel(data = filter(intron,Element %in% intron_elems),
                  aes(x=logFC,y=-log10(fdr),label=Symbol))
print(p1)

p2 <- ggplot(exon)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor,shape=Pheno),size=2)+
  scale_fill_manual(values=c(cbPalette[1],"lightgray"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  scale_shape_manual(values=c(21,23))+
  labs(fill="Expression Change",
       title="Reference vs. Variant Expression Change\nin Post-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")+
  theme(legend.position = "none",aspect.ratio = 1)+
  geom_text_repel(data = filter(exon,(Element %in% intron_elems | Element %in% exon_elems)),
                  aes(x=logFC,y=-log10(fdr),label=Symbol))
print(p2)

p3 <- ggplot(stability)+
  geom_point(aes(x=logFC_interaction,y=-log10(fdr_interaction),fill=plotcolor,shape=Pheno),size=2)+
  scale_fill_manual(values=c(cbPalette[1],"lightgrey"),
                    labels=c("Increased Stability","Decreased Stability",
                             "Not Significant"))+
  scale_shape_manual(values=c(21,23))+
  labs(fill="Stability Change",
       title="Reference vs. Variant Stability Change",
       x="Log2 Fold Change in Stability",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")+
  theme(legend.position = "none",aspect.ratio = 1)+
  geom_text_repel(data = filter(stability,Element %in% stability_elems),
                  aes(x=logFC_interaction,y=-log10(fdr_interaction),label=Symbol))
print(p3)

p4 <- ggplot(translational)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor,shape = Pheno),size=2)+
  scale_fill_manual(values=c(cbPalette[1],"lightgrey"),
                    labels=c("Increased Translation","Decreased Translation",
                             "Not Significant"))+
  scale_shape_manual(values=c(21,23))+
  labs(fill="Translational Change",
       title="Reference vs. Variant Translational Change",
       x="Log2 Fold Change in Translation",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")+
  guides(fill="none")+
  theme(aspect.ratio = 1)+
  geom_text_repel(data = filter(translational,Element %in% translational_elems),
                  aes(x=logFC,y=-log10(fdr),label=Symbol))+
  labs(shape="Phenotype")
print(p4)

export_graph_start("./Figures for Paper/Figure3/AllRefvsAlt_VolcanoPlot_Patchwork_New.pdf",sideways = TRUE,height = 0.5)
print(p1+theme(legend.position = "none")+labs(title=NULL,subtitle="Pre-Splicing",x="Log2FC Expression")+
        p2+theme(legend.position = "none")+labs(title=NULL,subtitle="Post-Splicing",x="Log2FC Expression")+
        p3+theme(legend.position = "none")+labs(title=NULL,subtitle="Stability",x="Log2FC Stability")+
        p4+theme(legend.position = "none")+labs(title=NULL,subtitle="Translational",x="Log2FC TE")+
        plot_layout(ncol=4)+plot_annotation(title="Reference vs. Variant"))
export_graph_end()

#Test case enrichment for library types
mat <- matrix(
  c(length(filter(intron,fdr<0.1,Pheno=="case")$Element),
    length(filter(intron,fdr<0.1,Pheno=="control")$Element),
    length(filter(intron,fdr>0.1,Pheno=="case")$Element),
    length(filter(intron,fdr>0.1,Pheno=="control")$Element)), 
  nrow = 2,
  dimnames = list(
    Variant = c("Case", "Control"),
    Status = c("Significant", "NonSignificant")
  )
)
fisher.test(mat,alternative = "greater")

mat <- matrix(
  c(length(filter(exon,fdr<0.1,Pheno=="case")$Element),
    length(filter(exon,fdr<0.1,Pheno=="control")$Element),
    length(filter(exon,fdr>0.1,Pheno=="case")$Element),
    length(filter(exon,fdr>0.1,Pheno=="control")$Element)), 
  nrow = 2,
  dimnames = list(
    Variant = c("Case", "Control"),
    Status = c("Significant", "NonSignificant")
  )
)
fisher.test(mat,alternative = "greater")

mat <- matrix(
  c(length(filter(stability,fdr_interaction<0.1,Pheno=="case")$Element),
length(filter(stability,fdr_interaction<0.1,Pheno=="control")$Element),
length(filter(stability,fdr_interaction>0.1,Pheno=="case")$Element),
length(filter(stability,fdr_interaction>0.1,Pheno=="control")$Element)), 
  nrow = 2,
  dimnames = list(
    Variant = c("Case", "Control"),
    Status = c("Significant", "NonSignificant")
  )
)

fisher.test(mat,alternative = "greater")

mat <- matrix(
  c(
    length(filter(stability,fdr_interaction>0.1,fdr_allele<0.1,Pheno=="case")$Element),
    length(filter(stability,fdr_interaction>0.1,fdr_allele<0.1,Pheno=="control")$Element),
    length(filter(stability,fdr_interaction>0.1,fdr_allele>0.1,Pheno=="case")$Element),
    length(filter(stability,fdr_interaction>0.1,fdr_allele>0.1,Pheno=="control")$Element)), 
  nrow = 2,
  dimnames = list(
    Variant = c("Case", "Control"),
    Status = c("Significant", "NonSignificant")
  )
)

fisher.test(mat,alternative = "greater")

#Post is sig, pre is barely sig, stability not sig at all,
#I bet there is a transcriptional bias in these?
#The "transcription only" group shows a significant enrichment too
#Mention it in the discussion

################################################################################
#Figure 4L - Run 4M code first
exon_stability <- left_join(exon_stability,plottable)
exon_stability <- unique.data.frame(exon_stability)
p5_dotplot <- ggplot(exon_stability,aes(x=logFC,y=logFC_interaction,fill=Category))+
  geom_point(shape=21,alpha=0.5,size=3)+
  scale_fill_manual(values = c("Not Significant" = "lightgrey",
                               "True Positive" = "#136d52",
                               "False Positive" = "#21c494",
                               "False Negative" = "#67e5bf"))+
  labs(x="Log Fold Change\n(Post-Splice)",
       y="Log Fold Change\n(Stability)",
       )+
  theme_bw()+
  theme(legend.position = "none",aspect.ratio = 1)
print(p5_dotplot)

export_graph_start("./Figures for Paper/Figure3/logfcpostvsstab_new_refvsalt.pdf")
print(p5_dotplot)
export_graph_end()
################################################################################
#Figure 4M
#False positive analysis
exon_stability <- inner_join(exon,stability,"Element")

#Significant in both and same direction
true_positives <- filter(exon_stability,
                         fdr<0.1,
                         fdr_interaction<0.1,
                         ((logFC>0 & logFC_interaction>0) | (logFC<0 & logFC_interaction<0))
                         )$Element
#Only significant in stability, not post-splicing
false_negatives <- filter(exon_stability,
                         fdr>0.1,
                         fdr_interaction<0.1)$Element
#Either only significant in post-splicing, or significant in both but opposite
#directions
false_positives <- c(
  filter(exon_stability,
         fdr<0.1,
         fdr_interaction>0.1)$Element,
  filter(exon_stability,
         fdr<0.1,
         fdr_interaction<0.1,
         ((logFC>0 & logFC_interaction<0) | (logFC<0 & logFC_interaction>0)))$Element)
  




all_else <- filter(stability_shuf,
                   !Element %in% false_positives,
                   !Element %in% false_negatives,
                   !Element %in% true_positives)$Element

plottable <- rbind(
  data.table("Element"=true_positives,"Category"="True Positive"),
  data.table("Element"=false_positives,"Category"="False Positive"),
  data.table("Element"=false_negatives,"Category"="False Negative"),
  data.table("Element"=all_else,"Category"="Not Significant")
)

plottable[,Category := factor(Category,levels=c("True Positive","False Positive","False Negative","Not Significant"))]

summary_data <- plottable %>%
  group_by(Category) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

# Separate the data for labels
label_data <- summary_data %>%
  filter(Category != "Not Significant") %>%
  mutate(cumulative_percentage = 100 - cumsum(percentage) + percentage/2)

# Create the plot

p5 <- ggplot(summary_data, aes(x = "", y = percentage, fill = Category)) +
  geom_col(width = 0.4, color = "black") +
  scale_fill_manual(values = c("Not Significant" = "lightgrey",
                               "True Positive" = "#136d52",
                               "False Positive" = "#21c494",
                               "False Negative" = "#67e5bf")) +
  labs(y = "Percentage",
       fill = "Category",
       x=NULL) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        legend.position = "right",
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

export_graph_start("./Figures for Paper/Figure3/RefAlt_Accuracy_New.pdf",width=0.75)
print(p5+theme_minimal())
ggplot(summary_data,aes(Category,count,fill=as.factor(count)))+
  geom_col()+
  geom_text(aes(y=count-100,label=paste0(round(percentage,2),"\n",count)))+
  theme(axis.text.x = element_text(angle = 45,hjust=1))
export_graph_end()

################################################################################

#4N,O,P - Example BoxPlots
#Let's write some new code using patchwork to make it a bit cleaner

d <- fread("../../data/formatted/ActD_Trials_08Apr2026/formatted.txt") %>% filter(,!Element %in% badAlt)

#Separate allele names
d <- d %>%
  separate_wider_regex(
    cols = Element,
    patterns = c(
      Element = ".*",        # Match as much as possible (greedy)
      "_",                   # Match the last underscore
      Allele = "[^_]+"       # Match everything remaining that isn't an underscore
    )
  )
d <- as.data.table(d)

#Grab the DNA info
dna <- fread("../../data/formatted/Library1and2BCLess_ISOMPRA_6-30-2022/formatted_Reads_compiled_06-30-2022.tsv") %>% filter(Fraction=="DNA")
dna <- dna %>% group_by(Element,Allele) %>% summarize(CPM=mean(CPM)) %>% as.data.table()
#Put the DNA info in the table, adjoining to correct element
d <- as.data.table(left_join(d,dna,by=c("Element","Allele")))
#Calculate expression
d[, Expression := log2(CPM.x/CPM.y)]
#Get rid of non-finite values
d <- filter(d,is.finite(Expression))
d <- filter(d,Fraction=="0hr",!(Splicing=="Pre" & BiolRep==3))
data_for_concordancy <- d #For later Figure S6
common_y <- c(
  min(filter(d,Element=="MLXIP_wgs", is.finite(Expression))$Expression),
  max(filter(d,Element=="MLXIP_wgs", is.finite(Expression))$Expression)
)

library(ggtext)
#Grab GALNTL5 variant, that was a good one
pre_boxplot <- ggplot(
  filter(d,Element=="MLXIP_wgs",Splicing=="pre") %>%
    mutate(Splicing=factor(Splicing,levels=c("pre","post"),labels=c("Pre","Post"))) %>%
    mutate(Allele=factor(Allele,levels=c("ref","alt"),labels=c("Reference","Variant"))),
  aes(x=Allele,y=Expression,fill=Allele))+
  geom_boxplot()+
  geom_point(shape=21)+
  scale_fill_manual(values = c("#136d52","#67e5bf"))+
  guides(fill="none")+
  theme(plot.title = element_text(hjust=0.5))+
  labs(title="Pre-Splicing",y = "Expression: Log<sub>2</sub> (<sup>RNA</sup>&frasl;<sub>DNA</sub>)") +
  theme(axis.title.y = element_markdown(),aspect.ratio = 2)+
  ylim(common_y)
print(pre_boxplot)

post_boxplot <- ggplot(
  filter(d,Element=="MLXIP_wgs",Splicing=="post") %>%
    mutate(Splicing=factor(Splicing,levels=c("pre","post"),labels=c("Pre","Post"))) %>%
    mutate(Allele=factor(Allele,levels=c("ref","alt"),labels=c("Reference","Variant"))),
  aes(x=Allele,y=Expression,fill=Allele))+
  geom_boxplot()+
  geom_point(shape=21)+
  scale_fill_manual(values = c("#136d52","#67e5bf"))+
  guides(fill="none")+
  theme(plot.title = element_text(hjust=0.5))+
  labs(title="Post-Splicing", y = "Expression: Log<sub>2</sub> (<sup>RNA</sup>&frasl;<sub>DNA</sub>)")+
  theme(axis.title.y = element_markdown())+
  theme(axis.title.y = element_blank(),aspect.ratio = 2)+
  ylim(common_y)
print(post_boxplot)

stab_boxplot <- ggplot(
  filter(d, Element == "MLXIP_wgs")%>%
    mutate(Splicing=factor(Splicing,levels=c("pre","post"),labels=c("Pre","Post"))) %>%
    mutate(Allele=factor(Allele,levels=c("ref","alt"),labels=c("Reference","Variant"))),
  aes(x = Splicing, y = Expression, fill = Allele)
) +
  # 1. Four distinct boxplots
  # We use position_dodge to separate Ref and Alt at each x-axis tick
  geom_boxplot(
    aes(group = interaction(Splicing, Allele)),
    position = position_dodge(width = 0.8),
    outlier.shape = NA
  ) +
  
  # 2. Dots centered over each of the four boxes
  geom_point(
    aes(group = interaction(Splicing, Allele)),
    position = position_dodge(width = 0.8), 
    shape = 21, 
    color = "black"
  ) +
  
  # 3. Lines connecting Pre-Ref to Post-Ref and Pre-Alt to Post-Alt
  # We use group = Allele here so it ignores the Splicing split for the line
  stat_summary(
    fun = median, 
    geom = "line", 
    aes(group = Allele, color = Allele), 
    linewidth = 1,
    position = position_dodge(width = 0.8) # Matches the boxplot shift
  ) +
  
  scale_fill_manual(values = c("#136d52", "#67e5bf")) +
  scale_color_manual(values = c("#136d52", "#67e5bf"))+
  guides(fill="none",color="none")+
  theme(plot.title = element_text(hjust=0.5))+
  labs(title="Stability",y = "Expression: Log<sub>2</sub> (<sup>RNA</sup>&frasl;<sub>DNA</sub>)") +
  theme(axis.title.y = element_markdown())+
  theme(axis.title.y = element_blank(),aspect.ratio = 2)+
  ylim(common_y)
print(stab_boxplot)

export_graph_start("./Figures for Paper/Figure3/New Figure 4N,O,P.pdf")
pre_boxplot+post_boxplot+stab_boxplot
export_graph_end()


```

```{r Figure S2C-H}
#Becca graciously offered to rerun the phyloP/Phastcons/GC/MFE plots, which will
#knock out 8 of these panels, yay
#Need to regenerate the post-splicing minus pre-splicing expression table, averaged on
#BiolRep
#I think for this, keep expression and DNA out of it, and just calculate log2(post CPM)-log2(pre CPM)
#Acutally talking with her and reviewing code, I made the GC and MFE plots, so let's make them again
dir.create("./Figures for Paper/Figure S2/")
d_stab_avg <- d_stab %>% filter(Allele=="ref",is.finite(Stability)) %>% group_by(Element) %>%
  summarize(Mean_Exon_Intron = mean(Stability))
d_stab_avg <- left_join(d_stab_avg,fread("../../code/mpra_tools/fullannotationtable.txt")[,c(2,15)],by="Element")
setnames(d_stab_avg,3,"Sequence")
fwrite(d_stab_avg,"Mean_Exon_Intron_New_for_Becca.txt",sep="\t")


#The reviewers asked for sequence features, so we're going to give them post-
#pre-splicing vs GC content, which is anticorrelated with free folding energy
#(makes sense that it would be) - this is for the new data to replace


GC_table <- fread("./Exon_Minus_Intron_Expression_Table_Average.txt")

free_folding <- readLines("./freefolding.txt")
free_folding <- data.table("Element" = free_folding[seq(1,24291,3)],
                           "Sequence" = free_folding[seq(2,24291,3)],
                           "Energy" = free_folding[seq(3,24291,3)])
free_folding$Element <- str_replace_all(free_folding$Element,"^>","")
free_folding[, DeltaG := as.numeric(str_extract(Energy, "-?\\d+\\.\\d+"))]
free_folding <- free_folding[,c(1,2,4)]

GC_table <- left_join(GC_table,free_folding[,c(1,3)])
#Get rid of alts and non-unique sequence
GC_table <- GC_table[!which(grepl("alt$",GC_table$Element))]
GC_table <- GC_table[!duplicated(GC_table$Sequence)]
GC_table <- GC_table %>%
  separate_wider_regex(
    cols = Element,
    patterns = c(
      Element = ".*",        # Match as much as possible (greedy)
      "_",                   # Match the last underscore
      Allele = "[^_]+"       # Match everything remaining that isn't an underscore
    )
  )
GC_table <- as.data.table(filter(GC_table,Allele=="ref"))
GC_table <- left_join(d_stab_avg,GC_table[,c(1,9,10)])


export_graph_start("./Figures for Paper/Figure S2/GC Content and Free Energy vs Stability Correlation_New.pdf")
ggplot(GC_table,aes(x=GC_content,y=Mean_Exon_Intron))+
  geom_pointdensity()+
  stat_smooth(method = "lm",color="black",linewidth=1)+
  labs(
    x="Element GC Content %",
    y="Stability (Post- Minus Pre-Splicing Expression",
    title="GC Content vs. Stability Correlation")+
  annotate(geom="label",
           x=29,
           y=4,
           label = paste0(
             "Spearman's Rho: ",
             round(
               cor.test(
                 GC_table$GC_content,
                 GC_table$Mean_Exon_Intron,
                 method="spearman")$estimate,2),
             "****"))+
  scale_color_viridis_c()+
  theme(legend.position = "none", aspect.ratio = .75)

ggplot(GC_table,aes(x=DeltaG,y=Mean_Exon_Intron))+
  geom_pointdensity()+
  stat_smooth(method = "lm",color="black",linewidth=1)+
  labs(
    x=expression("Minimum Free Energy (MFE) ("*Delta*"G, kcal/mol)"),
    y="Stability (Post- Minus Pre-Splicing Expression",
    title=expression("Minimum Free Energy (MFE, "*Delta*"G) vs. Stability Correlation"))+
  annotate(
    geom="label",
    x=-65,
    y=4,
    label = paste0(
      "Spearman's Rho: ",
      round(
        cor.test(GC_table$DeltaG,
                 GC_table$Mean_Exon_Intron,
                 method="spearman")$estimate,2),
      "****"))+
  scale_color_viridis_c()+
  theme(legend.position = "none",aspect.ratio = 0.75)
export_graph_end()

#Oh wow, now GC content correlates positively with stability, this is what we
#would expect to see
```




```{r Figure S6}
# Figure S6D,E,F - examples of concordant, discordant (fp and fn)

exon_stability <- left_join(exon_stability,plottable)
exon_stability <- unique.data.frame(exon_stability)

#Order the data table by the two logFCs to find some good points to plot
setorderv(exon_stability,c("logFC.x","logFC.y"))

#Also new 17Jun2025, make some condordant/discordant plots like Din's
#Make new data table of sig in post, not in stability, and vice versa
concord_false_pos <- filter(exon_stability,fdr.x<0.05,fdr.y>0.05)
concord_false_neg <- filter(exon_stability,fdr.x>0.05,fdr.y<0.05)
concord_true_pos <- filter(exon_stability,fdr.x<0.05,fdr.y<0.05)


data_for_concordancy <- left_join(
  filter(data_for_concordancy[,c(1,2,5,6,9)],Splicing=="pre"),
  filter(data_for_concordancy[,c(1,2,5,6,9)],Splicing=="post"),
  by=c("Element","Allele","BiolRep")
)
data_for_concordancy[,Stability := Expression.y-Expression.x]
#We will try to make a huge data table for isolated examples
#False positive selection is chr4:142314758;C|T;Family=12424 (in INPP4B),
#False negative selection is chr19:52291570;A|G;Family=13301 (in ZNF776),
#True positive selection is chr18:45899442;C|T;Family=11456 (in EPG5)
concord_graph_data <- rbind(
  filter(data_for_concordancy,
         Element=="PLEKHG4B_wgs",
         Allele=="ref"),
  filter(data_for_concordancy,
         Element=="chr19:52291570;A|G;Family=13301",
         Allele=="ref"),
  filter(data_for_concordancy,
         Element=="chr18:45899442;C|T;Family=11456",
         Allele=="ref"),
  filter(data_for_concordancy,
         Allele=="shuf") %>% mutate(Element="PLEKHG4B_wgs"),
  filter(data_for_concordancy,
         Allele=="shuf") %>% mutate(Element="chr19:52291570;A|G;Family=13301"),
  filter(data_for_concordancy,
         Allele=="shuf") %>% mutate(Element="chr18:45899442;C|T;Family=11456")
  )
concord_graph_data <- concord_graph_data[,c(1,2,3,7,8)]
#Merge with stability data to get gene name
concord_graph_data <- left_join(concord_graph_data,exon_stability[,c("Element","Symbol.x")])
setnames(concord_graph_data,,c("Element","Allele","BiolRep","Post-Splicing Expression","Stability","Gene"))
test <- as.data.table(pivot_longer(concord_graph_data,4:5,names_to = "Measurement",values_to = "Value"))
test[,Allele := factor(Allele,labels=c("Reference","Shuffled"))]
test[,Gene := factor(Gene,
levels=c("PLEKHG4B",
         "ZNF766",
         "EPG5"),
labels = c("Concordant",
           "Discordant\n(False Positive)",
           "Discordant\n(False Negative)"))]
#Manually set significance values
signif_df <- data.frame(
  Gene = rep(c("Concordant",
                 "Discordant\n(False Positive)",
                 "Discordant\n(False Negative)"), each = 2),
  xnum = rep(c(1, 2), times = 3),  # 1 = Post-Splicing, 2 = Stability
  xmin = rep(c(1 - 0.375), 6),    # Reference left dodge
  xmax = rep(c(1 + 0.375), 6),    # Shuffled right dodge
  y_position = c(4, 5, 4, 5, 4, 5),
  annotations = c("****", "****", "****", "NS", "NS", "****")
)

signif_df$xmin[seq(2,6,2)] <- 2 - 0.375
signif_df$xmax[seq(2,6,2)] <- 2 + 0.375

pconcord <- ggplot(test, aes(x = Measurement, y = Value, fill = Allele)) +
  geom_boxplot(outlier.size = 0.8, position = position_dodge(0.75)) +
  geom_point(shape = 21, position = position_dodge(0.75)) +
  facet_grid(. ~ Gene) +
  geom_signif(
    data = signif_df,
    aes(xmin = xmin, xmax = xmax, annotations = annotations, y_position = y_position),
    manual = TRUE,
    inherit.aes = FALSE,
    tip_length = 0.01,
    textsize = 4
  ) +
  labs(x = NULL, y = "Value")+
  scale_fill_manual(values = c("#136d52","#67e5bf"))+
  theme(aspect.ratio = 1.86)
print(pconcord)

export_graph_start("./Figures for Paper/Figure3/Concordancy_New.pdf",sideways = TRUE)
print(pconcord)
export_graph_end()



```




```{r Figure S7B}
# The dual fluorescence correlation analysis
# We will recycle old code from the dual flourescence folder to make it again

library(data.table)
library(tidyverse)
library(ggrepel)
library(fplot)
library(patchwork)

dt <- fread("./Dual_Fluorescence/ISOMPRA Full Transfection.txt")

# Initialize the result vector
result_vector <- numeric()

# Number of rows to process in each block
block_size <- 6

# Calculate the number of blocks
num_blocks <- ceiling(nrow(dt) / block_size)

# Iterate through each block
for (block in 1:num_blocks) {
  # Calculate the start and end rows for the current block
  start_row <- (block - 1) * block_size + 1
  end_row <- min(block * block_size, nrow(dt))
  
  # Get the current block of rows
  current_block <- dt[start_row:end_row, ]
  
  # Flatten the block column by column and append to the result vector
  for (col in seq_len(ncol(dt))) {
    result_vector <- c(result_vector, current_block[[col]])
  }
}

rm(dt)

d <- data.table("Element"=rep(c("chr9:137013102;G|A;Family=14275",
                                "chr11:17429244;G|A;Family=13057",
                                "chr14:74675817;G|A;Family=11462",
                                "chr4:95156901;A|G;Family=11368",
                              "CD7_wgs",
                              "chr16:28484111;A|G;Family=13368",
                              "FAXC_wgs",
                              "chr7:73340787;G|T;Family=13560",
                              "chr7:4769211;C|T;Family=13709",
                              "chr7:151987279;C|T;Family=13556",
                              "chr15:67373445;T|C;Family=14523",
                              "chr7:16088825;A|C;Family=11382",
                              "chr19:50977229;G|A;Family=22349",
                              "chr7:44384444;G|A;Family=13881",
                              "chr5:177490390;G|A;Family=14124",
                              "PPHLN1_wgs",
                              "chr20:25297973;G|A;Family=12858",
                              "SNX29_wgs",
                              "chr8:133459207;G|A;Family=12656",
                              "chr2:128190451;CTT|C;Family=14623",
                              "chr7:29916016;G|T;Family=12173",
                              "chr8:40531284;A|C;Family=14660"),
                              each=12),
                "Allele"=rep(c("Reference","Variant"),each=6),
                "Ratio"=result_vector[1:264])

d[, Allele := factor(Allele)]
d[,Ratio := log2(Ratio)]
fwrite(d,"ISOMPRA Full Data Table.txt")

d<- fread("./ISOMPRA Full Data Table.txt")
ggplot(d,aes(y=Ratio,x=Element,color=Allele))+geom_boxplot()+
  geom_vline(xintercept = seq(1.5,21.5,1))
ggplot(d,aes(y=Ratio,x=Allele,color=Allele))+geom_boxplot()+facet_wrap(vars(Element))

# Define a function to perform linear regression and extract relevant information
# Define a function to perform linear regression and extract relevant information
lm_summary <- function(dt_group) {
  lm_result <- lm(Ratio ~ Allele, data = dt_group)
  p_value <- summary(lm_result)$coefficients[2, 4]
  if (length(coef(lm_result)) > 1) {
    mean_diff <- coef(lm_result)[2]
  } else {
    mean_diff <- NA
  }
  return(list(Difference_of_Means = mean_diff, p_value = p_value))
}


# Perform linear regression and extract information for each group
result <- d[, lm_summary(.SD), by = Element]

#I need to figure out what ST3GAL1 is, I don't have it on the ISOMPRA list, it's
#not even in the annotated stuff.  I'll need to find the sequence in the counting
#pipeline

#The MPRA results, ST3GAL1 doesn't have an MPRA result.  I think this is from
#an earlier LM iteration that maybe didn't use a cutoff or didn't use EdgeR
#normalization, it is in the annotation table, and the counting pipeline, but
#probably didn't pass QC

#Might be in the new data, which I pass here
pre_df <- fread("../ActD Trials/Full_Library_08Apr2026/PreSplicing_RefAlt_LM.txt")
post_df <- fread("../ActD Trials/Full_Library_08Apr2026/PostSplicing_RefAlt_LM.txt")
agnostic_df <- fread("../Library1and2BCLess_ISOMPRA_6-30-2022/Splicing-Agnostic_RefvsAlt_LM_Annotated.tsv")
stab_df <- fread("../ActD Trials/Full_Library_08Apr2026/Stability_RefAlt_LM.txt") %>% mutate(logFC_interaction = -logFC_interaction)
trap_df <- fread("../Library1and2BCLess_ISOMPRA_6-30-2022/Translation_Splicing-Agnostic_RefvsAlt_LM_Annotated.tsv")

#Old Data for reference, comment out if not needed
pre_df <- fread("../Library1and2BCLess_ISOMPRA_6-30-2022/Pre-Splicing_RefvsAlt_LM_Annotated.tsv")
post_df <- fread("../Library1and2BCLess_ISOMPRA_6-30-2022/Post-Splicing_RefvsAlt_LM_Annotated.tsv")
agnostic_df <- fread("../Library1and2BCLess_ISOMPRA_6-30-2022/Splicing-Agnostic_RefvsAlt_LM_Annotated.tsv")
stab_df <- fread("../Library1and2BCLess_ISOMPRA_6-30-2022/Stability_RefvsAlt_LM_Annotated.tsv") %>% mutate(logFC_interaction = -logFC_interaction)
trap_df <- fread("../Library1and2BCLess_ISOMPRA_6-30-2022/Translation_Splicing-Agnostic_RefvsAlt_LM_Annotated.tsv")

result <- left_join(result,filter(pre_df,Element %in% result$Element)[,c(1,2,4,17)],"Element")
result <- left_join(result,filter(post_df,Element %in% result$Element)[,c(1,2,4,17)],"Element")
result <- left_join(result,filter(agnostic_df,Element %in% result$Element)[,c(1,2,4,17)],"Element")
result <- left_join(result,filter(stab_df,Element %in% result$Element)[,c(1,10,12,25)],"Element")
result <- left_join(result,filter(trap_df,Element %in% result$Element)[,c(1,2,4,17)],"Element")
setnames(result,
         c("logFC.x","logFC.y","logFC.x.x","logFC_interaction","logFC.y.y"),
         c("Pre_logFC","Post_logFC","Agnostic_logFC","Stability_logFC","TE_logFC"))
#result <- result[!duplicated(result$Symbol),]
p1 <- ggplot(result,aes(x=Difference_of_Means,y=Pre_logFC))+geom_point()+stat_smooth(method="lm")+stat_cor(method = "spearman", label.x = -1, label.y = 2)
p2 <- ggplot(result,aes(x=Difference_of_Means,y=Post_logFC))+geom_point()+stat_smooth(method="lm")+stat_cor(method = "spearman", label.x = -1, label.y = 2)
p3 <- ggplot(result,aes(x=Difference_of_Means,y=Agnostic_logFC))+geom_point()+stat_smooth(method="lm")+stat_cor(method = "spearman", label.x = -1, label.y = 2)
p4 <- ggplot(result,aes(x=Difference_of_Means,y=Stability_logFC))+geom_point()+stat_smooth(method="lm")+stat_cor(method = "spearman", label.x = -1, label.y = 2)
p5 <- ggplot(result,aes(x=Difference_of_Means,y=TE_logFC))+geom_point()+stat_smooth(method="lm")+stat_cor(method = "spearman", label.x = -1, label.y = 2)

p1+p2+p3+p4+p5+plot_layout(nrow=1)
#Not really helpful
comparisons <- data.table(
  "Pre"=sign(result$Difference_of_Means)==sign(result$Pre_logFC),
  "Post"=sign(result$Difference_of_Means)==sign(result$Post_logFC),
  "Agnostic"=sign(result$Difference_of_Means)==sign(result$Agnostic_logFC),
  "Stability"=sign(result$Difference_of_Means)==sign(result$Stability_logFC),
  "Translational Efficiency"=sign(result$Difference_of_Means)==sign(result$TE_logFC)
)
result_sig <- filter(result,p_value<0.05)
comparisons_sig <- data.table(
  "Pre"=sign(result_sig$Difference_of_Means)==sign(result_sig$Pre_logFC),
  "Post"=sign(result_sig$Difference_of_Means)==sign(result_sig$Post_logFC),
  "Agnostic"=sign(result_sig$Difference_of_Means)==sign(result_sig$Agnostic_logFC),
  "Stability"=sign(result_sig$Difference_of_Means)==sign(result_sig$Stability_logFC),
  "Translational Efficiency"=sign(result_sig$Difference_of_Means)==sign(result_sig$TE_logFC)
)

comparisons <- pivot_longer(comparisons,1:5,names_to = "Analysis",values_to = "Match")
comparisons_sig <- pivot_longer(comparisons_sig,1:5,names_to = "Analysis",values_to = "Match")

dir.create("../Figures for Paper/Figure S1")
export_graph_start("../Figures for Paper/Figure S1/DOE_Graph.pdf",sideways = TRUE)
ggplot(filter(comparisons,!is.na(Match)),#Analysis %in% c("Stability","Pre","Post")),
       aes(x=factor(Analysis,
                    levels = c("Stability",
                               "Translational Efficiency",
                               "Agnostic",
                               "Pre",
                               "Post")),fill=Match))+
  geom_bar(color="black")+
  scale_y_continuous(breaks = seq(1,22))+
  labs(y="Number",x="Comparison",fill="Same Direction of Effect",
       title="Direction of Effect of Variant from Fluorescence Reporter
       Compared to Direction of Effect from MPRA")+
  scale_fill_manual(values=c("white","darkgrey"))+
  theme(axis.text.x = element_text(angle=45,vjust = 0.75),
        plot.title = element_text(hjust=0.5))
export_graph_end()

ggplot(filter(comparisons_sig,!is.na(Match)),
       aes(x=factor(Analysis,
                    levels = c("Stability",
                               "Translational Efficiency",
                               "Agnostic",
                               "Post",
                               "Pre")),fill=Match))+
  geom_bar(color="black")+
  scale_y_continuous(breaks = seq(1,22))+
  labs(y="Number",x="Comparison",fill="Same Direction of Effect",
       title="Direction of Effect of Variant from Fluorescence Reporter
       Compared to Direction of Effect from MPRA")+
  scale_fill_manual(values=c("white","darkgrey"))+
  theme(axis.text.x = element_text(angle=45,vjust = 0.75))

ggplot(result,aes(x=Stability_logFC,y=TE_logFC))+geom_point()+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  stat_smooth(method="lm")
cor.test(result$Stability_logFC,result$TE_logFC,method = "spearman")$estimate

fwrite(result,"LM Testing.txt")

fwrite(result,"Dual Fluorescence Data.tsv",sep="\t")
```


```{r Figure S8}
# Figure S8 is a volcano plot of main effect of allele only (which shows the
# allele of post-splice, that is how the interaction model is generated), no
# significant interaction - same slopes pre to post
# And some representative boxplots

#####Variants that only changed pre-splicing, from stability#####


#Let's especially label two hits for exon which weren't actually hits for stability
#But had a main effect of allele

stability[,plotcolor2 := as.factor(case_when(
    (abs(logFC_allele)>log2(1.25) & fdr_allele < 0.1 & fdr_interaction > 0.1) ~ 1,
    TRUE ~ 2))]

intersect(filter(exon,fdr<0.1)$Element,
                    filter(stability, fdr_interaction>0.1, fdr_allele<0.1)$Element)

export_graph_start("./Figures for Paper/Figure3/Transcriptional-Only_VolcanoPlot_New.pdf")
p4 <- ggplot(stability)+
  geom_point(aes(x=logFC_allele,y=-log10(fdr_allele),fill=plotcolor2,shape=Pheno),size=2)+
  scale_shape_manual(values=c(21,23))+
  scale_fill_manual(values=c(cbPalette[1],"lightgrey"))+
  geom_text_repel(data=filter(stability,Element %in% c("MORN3","F2RL3")),
                  aes(x=logFC_allele,y=-log10(fdr_allele),label=Element))+
  labs(title="Variant vs. Reference Expression Change\nWith No Stability Effect",
       x="Log2 Fold Change in Main Effect of Allele",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")+
  guides(fill="none")
print(p4)
export_graph_end()

GRSF1_boxplot <- ggplot(
  filter(d, Element == "chr4:70817164;CAAGTTT|C;Family=12175")%>%
    mutate(Splicing=factor(Splicing,levels=c("pre","post"),labels=c("Pre","Post"))) %>%
    mutate(Allele=factor(Allele,levels=c("ref","alt"),labels=c("Reference","Variant"))),
  aes(x = Splicing, y = Expression, fill = Allele)
) +
  # 1. Four distinct boxplots
  # We use position_dodge to separate Ref and Alt at each x-axis tick
  geom_boxplot(
    aes(group = interaction(Splicing, Allele)),
    position = position_dodge(width = 0.8),
    outlier.shape = NA
  ) +
  
  # 2. Dots centered over each of the four boxes
  geom_point(
    aes(group = interaction(Splicing, Allele)),
    position = position_dodge(width = 0.8), 
    shape = 21, 
    color = "black"
  ) +
  
  # 3. Lines connecting Pre-Ref to Post-Ref and Pre-Alt to Post-Alt
  # We use group = Allele here so it ignores the Splicing split for the line
  stat_summary(
    fun = median, 
    geom = "line", 
    aes(group = Allele, color = Allele), 
    linewidth = 1,
    position = position_dodge(width = 0.8) # Matches the boxplot shift
  ) +
  
  scale_fill_manual(values = c("#136d52", "#67e5bf")) +
  scale_color_manual(values = c("#136d52", "#67e5bf"))+
  guides(fill="none",color="none")+
  theme(plot.title = element_text(hjust=0.5))+
  labs(title="GRSF1",y = "Expression: Log<sub>2</sub> (<sup>RNA</sup>&frasl;<sub>DNA</sub>)") +
  theme(axis.title.y = element_markdown())+
  theme(axis.title.y = element_blank())
print(GRFS1_boxplot)

ATXN3_boxplot <- ggplot(
  filter(d, Element == "chr14:92045541;A|C;Family=12892")%>%
    mutate(Splicing=factor(Splicing,levels=c("pre","post"),labels=c("Pre","Post"))) %>%
    mutate(Allele=factor(Allele,levels=c("ref","alt"),labels=c("Reference","Variant"))),
  aes(x = Splicing, y = Expression, fill = Allele)
) +
  # 1. Four distinct boxplots
  # We use position_dodge to separate Ref and Alt at each x-axis tick
  geom_boxplot(
    aes(group = interaction(Splicing, Allele)),
    position = position_dodge(width = 0.8),
    outlier.shape = NA
  ) +
  
  # 2. Dots centered over each of the four boxes
  geom_point(
    aes(group = interaction(Splicing, Allele)),
    position = position_dodge(width = 0.8), 
    shape = 21, 
    color = "black"
  ) +
  
  # 3. Lines connecting Pre-Ref to Post-Ref and Pre-Alt to Post-Alt
  # We use group = Allele here so it ignores the Splicing split for the line
  stat_summary(
    fun = median, 
    geom = "line", 
    aes(group = Allele, color = Allele), 
    linewidth = 1,
    position = position_dodge(width = 0.8) # Matches the boxplot shift
  ) +
  
  scale_fill_manual(values = c("#136d52", "#67e5bf")) +
  scale_color_manual(values = c("#136d52", "#67e5bf"))+
  guides(fill="none",color="none")+
  theme(plot.title = element_text(hjust=0.5))+
  labs(title="ATXN3",y = "Expression: Log<sub>2</sub> (<sup>RNA</sup>&frasl;<sub>DNA</sub>)") +
  theme(axis.title.y = element_markdown())+
  theme(axis.title.y = element_blank())
print(ATXN3_boxplot)

export_graph_start("./Figures for Paper/Figure3/Transcriptional-Only_Boxplots_New.pdf")
ATXN3_boxplot+GRSF1_boxplot
export_graph_end()
```

## Making New FDR Cutoff Plots for Reviewer 2
###Functions
#### Bar Graph Functions

```{r fdr-bar-graph-functions}
# As part of the review process, we were also asked/decided to make additional
# cutoff plots at fdr<0.05, fdr<0.1 and fdr<0.2, so we are going to make those now.

# Create directory ----
dir.create("./Figures for Paper/Reviewer Only Figures")

# Bar Graph Functions ----
## Ref vs Shuf Bar Graphs Function ----

plot_fdr_bargraphs_shuf <- function(exon_df, stability_df, export_path) {
  
  # 1. Inner join the two datasets on 'Element'
  exon_stability <- inner_join(exon_df, stability_df, by = "Element")
  
  # 2. Use a loop to calculate categories across all thresholds dynamically
  thresholds <- c(0.05, 0.1, 0.2)
  plottable_list <- list()
  
  for (t in thresholds) {
    # True Positives: Sig in both, same direction
    tp <- filter(exon_stability,
                 fdr.x < t, fdr.y < t,
                 ((logFC.x > 0 & logFC.y > 0) | (logFC.x < 0 & logFC.y < 0)))$Element
    
    # False Negatives: Sig in stability, not in post-splicing (exon)
    fn <- filter(exon_stability, fdr.x > t, fdr.y < t)$Element
    
    # False Positives: Sig in post-splicing but not stability, OR sig in both but opposite directions
    fp <- c(
      filter(exon_stability, fdr.x < t, fdr.y > t)$Element,
      filter(exon_stability, fdr.x < t, fdr.y < t, ((logFC.x > 0 & logFC.y < 0) | (logFC.x < 0 & logFC.y > 0)))$Element
    )
    
    # Not Significant: Anything left in the stability table
    ns <- filter(stability_df, 
                 !Element %in% tp, 
                 !Element %in% fp, 
                 !Element %in% fn)$Element
    
    # Bind into a data.table for the current threshold tier
    plottable_list[[as.character(t)]] <- rbind(
      data.table("Element" = tp, "Category" = "True Positive",  "Sig" = as.factor(t)),
      data.table("Element" = fp, "Category" = "False Positive", "Sig" = as.factor(t)),
      data.table("Element" = fn, "Category" = "False Negative", "Sig" = as.factor(t)),
      data.table("Element" = ns, "Category" = "Not Significant", "Sig" = as.factor(t))
    )
  }
  
  # Combine all thresholds into one data.table
  plottable <- rbindlist(plottable_list)
  
  # 3. Factor levels for structural ordering
  plottable[, Category := factor(Category, levels = c("True Positive", "False Positive", "False Negative", "Not Significant"))]
  
  # 4. Generate summary stats
  summary_data <- plottable %>%
    group_by(Sig, Category) %>%                      
    summarise(count = n(), .groups = "drop_last") %>% 
    mutate(percentage = count / sum(count) * 100) %>% 
    ungroup()
  
  # 5. Construct the ggplot object
  p_out <- ggplot(summary_data, aes(x = Sig, y = percentage, fill = Category)) +
    geom_col(width = 0.85, color = "black") +
    scale_fill_manual(values = c("Not Significant" = "lightgrey",
                                 "True Positive"    = "#136d52",
                                 "False Positive"   = "#21c494",
                                 "False Negative"   = "#67e5bf")) +
    geom_text(aes(label = paste0(signif(percentage, digits = 3), "%")), 
          position = position_stack(vjust = 0.5), 
          size = 3.5, 
          fontface = "bold")+
    labs(y = "Percentage",
         x = "Significance Level (FDR)",
         fill = "Category",
         title = "False Positive Analysis") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
          legend.position = "right")
  
  # 6. Display plot and handle export commands
  print(p_out)
  
  export_graph_start(export_path)
  print(p_out)
  export_graph_end()
}

## Ref vs Alt Bar Graphs Function ----

plot_fdr_bargraphs_alt <- function(exon_df, stability_df, export_path) {
  
  # 1. Inner join the two datasets on 'Element'
  exon_stability <- inner_join(exon_df, stability_df, by = "Element")
  
  # 2. Use a loop to calculate categories across all thresholds dynamically
  thresholds <- c(0.05, 0.1, 0.2)
  plottable_list <- list()
  
  for (t in thresholds) {
    # True Positives: Sig in both, same direction
    tp <- filter(exon_stability,
                 fdr < t, fdr_interaction < t,
                 ((logFC > 0 & logFC_interaction > 0) | (logFC < 0 & logFC_interaction < 0)))$Element
    
    # False Negatives: Sig in stability, not in post-splicing (exon)
    fn <- filter(exon_stability, fdr > t, fdr_interaction < t)$Element
    
    # False Positives: Sig in post-splicing but not stability, OR sig in both but opposite directions
    fp <- c(
      filter(exon_stability, fdr < t, fdr_interaction > t)$Element,
      filter(exon_stability, fdr < t, fdr_interaction < t, ((logFC > 0 & logFC_interaction < 0) | (logFC < 0 & logFC_interaction > 0)))$Element
    )
    
    # Not Significant: Anything left in the stability table
    ns <- filter(stability_df, 
                 !Element %in% tp, 
                 !Element %in% fp, 
                 !Element %in% fn)$Element
    
    # Bind into a data.table for the current threshold tier
    plottable_list[[as.character(t)]] <- rbind(
      data.table("Element" = tp, "Category" = "True Positive",  "Sig" = as.factor(t)),
      data.table("Element" = fp, "Category" = "False Positive", "Sig" = as.factor(t)),
      data.table("Element" = fn, "Category" = "False Negative", "Sig" = as.factor(t)),
      data.table("Element" = ns, "Category" = "Not Significant", "Sig" = as.factor(t))
    )
  }
  
  # Combine all thresholds into one data.table
  plottable <- rbindlist(plottable_list)
  
  # 3. Factor levels for structural ordering
  plottable[, Category := factor(Category, levels = c("True Positive", "False Positive", "False Negative", "Not Significant"))]
  
  # 4. Generate summary stats
  summary_data <- plottable %>%
    group_by(Sig, Category) %>%                      
    summarise(count = n(), .groups = "drop_last") %>% 
    mutate(percentage = count / sum(count) * 100) %>% 
    ungroup()
  
  # 5. Construct the ggplot object
  p_out <- ggplot(summary_data, aes(x = Sig, y = percentage, fill = Category)) +
    geom_col(width = 0.85, color = "black") +
    scale_fill_manual(values = c("Not Significant" = "lightgrey",
                                 "True Positive"    = "#136d52",
                                 "False Positive"   = "#21c494",
                                 "False Negative"   = "#67e5bf")) +
    geom_text(aes(label = paste0(signif(percentage, digits = 3), "%")), 
          position = position_stack(vjust = 0.5), 
          size = 3.5, 
          fontface = "bold") +
    labs(y = "Percentage",
         x = "Significance Level (FDR)",
         fill = "Category",
         title = "False Positive Analysis") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
          legend.position = "right")
  
  # 6. Display plot and handle export commands
  print(p_out)
  
  export_graph_start(export_path)
  print(p_out)
  export_graph_end()
}
```

####Volcano Plot Functions

```{r}

make_post_stab_volcanoplots_refshuf <- function(exon_df, stability_df,filename){
  exon_df <- exon_df %>% as.data.table()
  stability_df <- stability_df %>% as.data.table()
  exon_df[,plotcolor := factor(plotcolor,levels=c("FDR < 0.05","FDR < 0.1","FDR < 0.2","Not Significant"))]
  stability_df[,plotcolor := factor(plotcolor,levels=c("FDR < 0.05","FDR < 0.1","FDR < 0.2","Not Significant"))]
  
  p1 <- ggplot(exon_df, aes(x = logFC, y = -log10(fdr))) +
    # Points filled by data-level plotcolor
    geom_point(aes(fill = plotcolor), shape = 21, size = 3) + 
    
    # Static black vertical lines for fold change
    geom_vline(xintercept = c(-log2(1.25), log2(1.25)), linetype = "dashed", color = "black") +
    
    # Horizontal lines with mapped color labels
    geom_hline(aes(yintercept = -log10(0.05), color = "FDR = 0.05"), linewidth = 0.8, linetype = "dashed") +
    geom_hline(aes(yintercept = -log10(0.1),  color = "FDR = 0.10"), linewidth = 0.8, linetype = "dashed") +
    geom_hline(aes(yintercept = -log10(0.2),  color = "FDR = 0.20"), linewidth = 0.8, linetype = "dashed") +
    
    # Custom colors for the points (fill)
    scale_fill_manual(name="Significance",
                      values = c("FDR < 0.05" = "#136d52", 
                                 "FDR < 0.1" = "#21c494", 
                                 "FDR < 0.2" = "#67e5bf", 
                                 "Not Significant" = "lightgrey")) +
    
    # Custom colors for the lines (color)
    scale_color_manual(name = "Thresholds",
                       values = c("FDR = 0.05" = "#136d52", 
                                  "FDR = 0.10" = "#21c494", 
                                  "FDR = 0.20" = "#67e5bf")) +
    theme_minimal()+
    guides(color="none",fill="none")+
    labs(x=expression(Log[2] * FC ~ Expression),
         y=expression("-" * log[10] * (FDR)),
         title="Post-Splicing")+
    theme(plot.title = element_text(hjust=0.5), aspect.ratio = 1)
  
  p2 <- ggplot(stability_df, aes(x = logFC, y = -log10(fdr))) +
    # Points filled by data-level plotcolor
    geom_point(aes(fill = plotcolor), shape = 21, size = 3) + 
    
    # Static black vertical lines for fold change
    geom_vline(xintercept = c(-log2(1.25), log2(1.25)), linetype = "dashed", color = "black") +
    
    # Horizontal lines with mapped color labels
    geom_hline(aes(yintercept = -log10(0.05), color = "FDR = 0.05"), linewidth = 0.8, linetype = "dashed") +
    geom_hline(aes(yintercept = -log10(0.1),  color = "FDR = 0.10"), linewidth = 0.8, linetype = "dashed") +
    geom_hline(aes(yintercept = -log10(0.2),  color = "FDR = 0.20"), linewidth = 0.8, linetype = "dashed") +
    
    # Custom colors for the points (fill)
    scale_fill_manual(name="Significance",
                      values = c("FDR < 0.05" = "#136d52", 
                                 "FDR < 0.1" = "#21c494", 
                                 "FDR < 0.2" = "#67e5bf", 
                                 "Not Significant" = "lightgrey")) +
    
    # Custom colors for the lines (color)
    scale_color_manual(name = "Thresholds",
                       values = c("FDR = 0.05" = "#136d52", 
                                  "FDR = 0.10" = "#21c494", 
                                  "FDR = 0.20" = "#67e5bf")) +
    theme_minimal()+
    guides(color="none")+
    labs(x=expression(Log[2] * FC ~ Stability),
         y=expression("-" * log[10] * (FDR)),
         title="Stability")+
    theme(plot.title = element_text(hjust=0.5),aspect.ratio = 1)
  
  export_graph_start(filename,sideways=TRUE,height=0.5)
  print(p1+p2)
  export_graph_end()
  
}

make_post_stab_volcanoplots_refalt <- function(exon_df, stability_df,filename){
  exon_df <- exon_df %>% as.data.table()
  stability_df <- stability_df %>% as.data.table()
  exon_df[,plotcolor := factor(plotcolor,levels=c("FDR < 0.05","FDR < 0.1","FDR < 0.2","Not Significant"))]
  stability_df[,plotcolor := factor(plotcolor,levels=c("FDR < 0.05","FDR < 0.1","FDR < 0.2","Not Significant"))]
  
  p1 <- ggplot(exon_df, aes(x = logFC, y = -log10(fdr))) +
    # Points filled by data-level plotcolor
    geom_point(aes(fill = plotcolor), shape = 21, size = 3) + 
    
    # Static black vertical lines for fold change
    geom_vline(xintercept = c(-log2(1.25), log2(1.25)), linetype = "dashed", color = "black") +
    
    # Horizontal lines with mapped color labels
    geom_hline(aes(yintercept = -log10(0.05), color = "FDR = 0.05"), linewidth = 0.8, linetype = "dashed") +
    geom_hline(aes(yintercept = -log10(0.1),  color = "FDR = 0.10"), linewidth = 0.8, linetype = "dashed") +
    geom_hline(aes(yintercept = -log10(0.2),  color = "FDR = 0.20"), linewidth = 0.8, linetype = "dashed") +
    
    # Custom colors for the points (fill)
    scale_fill_manual(name="Significance",
                      values = c("FDR < 0.05" = "#136d52", 
                                 "FDR < 0.1" = "#21c494", 
                                 "FDR < 0.2" = "#67e5bf", 
                                 "Not Significant" = "lightgrey")) +
    
    # Custom colors for the lines (color)
    scale_color_manual(name = "Thresholds",
                       values = c("FDR = 0.05" = "#136d52", 
                                  "FDR = 0.10" = "#21c494", 
                                  "FDR = 0.20" = "#67e5bf")) +
    theme_minimal()+
    guides(color="none",fill="none")+
    labs(x=expression(Log[2] * FC ~ Expression),
         y=expression("-" * log[10] * (FDR)),
         title="Post-Splicing")+
    theme(plot.title = element_text(hjust=0.5),aspect.ratio = 1)
  
  p2 <- ggplot(stability_df, aes(x = logFC_interaction, y = -log10(fdr_interaction))) +
    # Points filled by data-level plotcolor
    geom_point(aes(fill = plotcolor), shape = 21, size = 3) + 
    
    # Static black vertical lines for fold change
    geom_vline(xintercept = c(-log2(1.25), log2(1.25)), linetype = "dashed", color = "black") +
    
    # Horizontal lines with mapped color labels
    geom_hline(aes(yintercept = -log10(0.05), color = "FDR = 0.05"), linewidth = 0.8, linetype = "dashed") +
    geom_hline(aes(yintercept = -log10(0.1),  color = "FDR = 0.10"), linewidth = 0.8, linetype = "dashed") +
    geom_hline(aes(yintercept = -log10(0.2),  color = "FDR = 0.20"), linewidth = 0.8, linetype = "dashed") +
    
    # Custom colors for the points (fill)
    scale_fill_manual(name="Significance",
                      values = c("FDR < 0.05" = "#136d52", 
                                 "FDR < 0.1" = "#21c494", 
                                 "FDR < 0.2" = "#67e5bf", 
                                 "Not Significant" = "lightgrey")) +
    
    # Custom colors for the lines (color)
    scale_color_manual(name = "Thresholds",
                       values = c("FDR = 0.05" = "#136d52", 
                                  "FDR = 0.10" = "#21c494", 
                                  "FDR = 0.20" = "#67e5bf")) +
    theme_minimal()+
    guides(color="none")+
    labs(x=expression(Log[2] * FC ~ Stability),
         y=expression("-" * log[10] * (FDR)),
         title="Stability")+
    theme(plot.title = element_text(hjust=0.5),aspect.ratio = 1)
  
  export_graph_start(filename,sideways=TRUE,height=0.5)
  print(p1+p2)
  export_graph_end()
  
}

```

### Retrieving Data

```{r}
# Read in refshuf in vitro ----
intron_shuf <- fread("../ActD Trials/Full_Library_08Apr2026/Pre-Splicing_RefvsShuf_Welchs.csv") %>% 
  .[, plotcolor := as.factor(case_when(
    abs(logFC) > log2(1.25) & fdr < 0.05               ~ "FDR < 0.05",
    abs(logFC) > log2(1.25) & (fdr >= 0.05 & fdr < 0.1) ~ "FDR < 0.1",
    abs(logFC) > log2(1.25) & (fdr >= 0.1 & fdr < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

exon_shuf <- fread("../ActD Trials/Full_Library_08Apr2026/Post-Splicing_RefvsShuf_Welchs.tsv") %>% 
  .[, plotcolor := as.factor(case_when(
    abs(logFC) > log2(1.25) & fdr < 0.05               ~ "FDR < 0.05",
    abs(logFC) > log2(1.25) & (fdr >= 0.05 & fdr < 0.1) ~ "FDR < 0.1",
    abs(logFC) > log2(1.25) & (fdr >= 0.1 & fdr < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

stability_shuf <- fread("../ActD Trials/Full_Library_08Apr2026/Stability_RefvsShuf_Welchs.tsv") %>%
 .[, plotcolor := as.factor(case_when(
    abs(logFC) > log2(1.25) & fdr < 0.05               ~ "FDR < 0.05",
    abs(logFC) > log2(1.25) & (fdr >= 0.05 & fdr < 0.1) ~ "FDR < 0.1",
    abs(logFC) > log2(1.25) & (fdr >= 0.1 & fdr < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

translational_shuf <- fread("./Translation_Splicing-Agnostic_RefvsShuf_Welchs.tsv") %>% 
  .[, plotcolor := as.factor(case_when(
    abs(logFC) > log2(1.25) & fdr < 0.05               ~ "FDR < 0.05",
    abs(logFC) > log2(1.25) & (fdr >= 0.05 & fdr < 0.1) ~ "FDR < 0.1",
    abs(logFC) > log2(1.25) & (fdr >= 0.1 & fdr < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

# Read in refalt in vitro ----
intron <- fread("../ActD Trials/Full_Library_08Apr2026/PreSplicing_RefAlt_LM.txt") %>% 
  .[, plotcolor := as.factor(case_when(
    abs(logFC) > log2(1.25) & fdr < 0.05               ~ "FDR < 0.05",
    abs(logFC) > log2(1.25) & (fdr >= 0.05 & fdr < 0.1) ~ "FDR < 0.1",
    abs(logFC) > log2(1.25) & (fdr >= 0.1 & fdr < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

exon <- fread("../ActD Trials/Full_Library_08Apr2026/PostSplicing_RefAlt_LM.txt") %>% .[, plotcolor := as.factor(case_when(
    abs(logFC) > log2(1.25) & fdr < 0.05               ~ "FDR < 0.05",
    abs(logFC) > log2(1.25) & (fdr >= 0.05 & fdr < 0.1) ~ "FDR < 0.1",
    abs(logFC) > log2(1.25) & (fdr >= 0.1 & fdr < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

stability <- fread("../ActD Trials/Full_Library_08Apr2026/Stability_RefAlt_LM.txt") %>%
  .[, plotcolor := as.factor(case_when(
    abs(logFC_interaction) > log2(1.25) & fdr_interaction < 0.05               ~ "FDR < 0.05",
    abs(logFC_interaction) > log2(1.25) & (fdr_interaction >= 0.05 & fdr_interaction < 0.1) ~ "FDR < 0.1",
    abs(logFC_interaction) > log2(1.25) & (fdr_interaction >= 0.1 & fdr_interaction < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

translational <- fread("./Translation_Splicing-Agnostic_RefvsShuf_Welchs.tsv") %>% 
  .[, plotcolor := as.factor(case_when(
    abs(logFC) > log2(1.25) & fdr < 0.05               ~ "FDR < 0.05",
    abs(logFC) > log2(1.25) & (fdr >= 0.05 & fdr < 0.1) ~ "FDR < 0.1",
    abs(logFC) > log2(1.25) & (fdr >= 0.1 & fdr < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

# Read in in vivo whole cortex ref vs shuf ----

intron_shuf_cortex <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Pre-Splicing_Vglut1_CreOFF_RefvsShuf_Welchs_Annotated.tsv") %>%
  .[, plotcolor := as.factor(case_when(
    abs(logFC) > log2(1.25) & fdr < 0.05               ~ "FDR < 0.05",
    abs(logFC) > log2(1.25) & (fdr >= 0.05 & fdr < 0.1) ~ "FDR < 0.1",
    abs(logFC) > log2(1.25) & (fdr >= 0.1 & fdr < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

exon_shuf_cortex <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Post-Splicing_Vglut1_CreOFF_RefvsShuf_Welchs_Annotated.tsv") %>%
  .[, plotcolor := as.factor(case_when(
    abs(logFC) > log2(1.25) & fdr < 0.05               ~ "FDR < 0.05",
    abs(logFC) > log2(1.25) & (fdr >= 0.05 & fdr < 0.1) ~ "FDR < 0.1",
    abs(logFC) > log2(1.25) & (fdr >= 0.1 & fdr < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

stability_shuf_cortex <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreOFF_RefvsShuf_Welchs_Annotated.tsv") %>%
 .[, plotcolor := as.factor(case_when(
    abs(logFC) > log2(1.25) & fdr < 0.05               ~ "FDR < 0.05",
    abs(logFC) > log2(1.25) & (fdr >= 0.05 & fdr < 0.1) ~ "FDR < 0.1",
    abs(logFC) > log2(1.25) & (fdr >= 0.1 & fdr < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

# Read in in vivo whole cortex ref vs alt ----

intron_cortex <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Pre-Splicing_Vglut1_CreOFF_RefvsAlt_LM_Annotated.tsv") %>%
  .[, plotcolor := as.factor(case_when(
    abs(logFC) > log2(1.25) & fdr < 0.05               ~ "FDR < 0.05",
    abs(logFC) > log2(1.25) & (fdr >= 0.05 & fdr < 0.1) ~ "FDR < 0.1",
    abs(logFC) > log2(1.25) & (fdr >= 0.1 & fdr < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

exon_cortex <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Post-Splicing_Vglut1_CreOFF_RefvsAlt_LM_Annotated.tsv") %>%
  .[, plotcolor := as.factor(case_when(
    abs(logFC) > log2(1.25) & fdr < 0.05               ~ "FDR < 0.05",
    abs(logFC) > log2(1.25) & (fdr >= 0.05 & fdr < 0.1) ~ "FDR < 0.1",
    abs(logFC) > log2(1.25) & (fdr >= 0.1 & fdr < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

stability_cortex <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreOFF_RefvsAlt_LM_Annotated.tsv") %>%
 .[, plotcolor := as.factor(case_when(
    abs(logFC_interaction) > log2(1.25) & fdr_interaction < 0.05               ~ "FDR < 0.05",
    abs(logFC_interaction) > log2(1.25) & (fdr_interaction >= 0.05 & fdr_interaction < 0.1) ~ "FDR < 0.1",
    abs(logFC_interaction) > log2(1.25) & (fdr_interaction >= 0.1 & fdr_interaction < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

# Read in in vivo exn ref vs shuf ----
intron_shuf_neuron <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Pre-Splicing_Vglut1_CreON_RefvsShuf_Welchs_Annotated.tsv") %>%
  .[, plotcolor := as.factor(case_when(
    abs(logFC) > log2(1.25) & fdr < 0.05               ~ "FDR < 0.05",
    abs(logFC) > log2(1.25) & (fdr >= 0.05 & fdr < 0.1) ~ "FDR < 0.1",
    abs(logFC) > log2(1.25) & (fdr >= 0.1 & fdr < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

exon_shuf_neuron <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Post-Splicing_Vglut1_CreON_RefvsShuf_Welchs_Annotated.tsv") %>%
  .[, plotcolor := as.factor(case_when(
    abs(logFC) > log2(1.25) & fdr < 0.05               ~ "FDR < 0.05",
    abs(logFC) > log2(1.25) & (fdr >= 0.05 & fdr < 0.1) ~ "FDR < 0.1",
    abs(logFC) > log2(1.25) & (fdr >= 0.1 & fdr < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

stability_shuf_neuron <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreON_RefvsShuf_Welchs_Annotated.tsv") %>%
 .[, plotcolor := as.factor(case_when(
    abs(logFC) > log2(1.25) & fdr < 0.05               ~ "FDR < 0.05",
    abs(logFC) > log2(1.25) & (fdr >= 0.05 & fdr < 0.1) ~ "FDR < 0.1",
    abs(logFC) > log2(1.25) & (fdr >= 0.1 & fdr < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

# Read in in vivo exn ref vs alt ----

intron_neuron <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Pre-Splicing_Vglut1_CreON_RefvsAlt_LM_Annotated.tsv") %>%
  .[, plotcolor := as.factor(case_when(
    abs(logFC) > log2(1.25) & fdr < 0.05               ~ "FDR < 0.05",
    abs(logFC) > log2(1.25) & (fdr >= 0.05 & fdr < 0.1) ~ "FDR < 0.1",
    abs(logFC) > log2(1.25) & (fdr >= 0.1 & fdr < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

exon_neuron <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Post-Splicing_Vglut1_CreON_RefvsAlt_LM_Annotated.tsv") %>%
  .[, plotcolor := as.factor(case_when(
    abs(logFC) > log2(1.25) & fdr < 0.05               ~ "FDR < 0.05",
    abs(logFC) > log2(1.25) & (fdr >= 0.05 & fdr < 0.1) ~ "FDR < 0.1",
    abs(logFC) > log2(1.25) & (fdr >= 0.1 & fdr < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

stability_neuron <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreON_RefvsAlt_LM_Annotated.tsv") %>%
 .[, plotcolor := as.factor(case_when(
    abs(logFC_interaction) > log2(1.25) & fdr_interaction < 0.05               ~ "FDR < 0.05",
    abs(logFC_interaction) > log2(1.25) & (fdr_interaction >= 0.05 & fdr_interaction < 0.1) ~ "FDR < 0.1",
    abs(logFC_interaction) > log2(1.25) & (fdr_interaction >= 0.1 & fdr_interaction < 0.2)  ~ "FDR < 0.2",
    TRUE                                                ~ "Not Significant"
  ))]

```

### Bar Graphs
#### In Vitro Data
##### Ref vs Shuf

```{r in-vitro-fdr-data}

plot_fdr_bargraphs_shuf(exon_shuf,stability_shuf,"./Figures for Paper/Reviewer Only Figures/In Vitro FDR Cutoff Bars RefvsShuf.pdf")
```

##### Ref vs. Alt
```{r ref-vs-alt}

plot_fdr_bargraphs_alt(exon,stability,"./Figures for Paper/Reviewer Only Figures/In Vitro FDR Cutoff Bars RefvsAlt.pdf")

```

#### In Vivo Data
##### In Vivo Whole Cortex Ref vs Shuf

```{r in-vivo-fdr-data}


plot_fdr_bargraphs_shuf(exon_shuf_cortex,stability_shuf_cortex,"./Figures for Paper/Reviewer Only Figures/In Vivo FDR Cutoff Bars RefvsShuf Cortex.pdf")
```

##### In Vivo Whole Cortex Ref vs Alt

```{r in-vivo-fdr-data}


plot_fdr_bargraphs_alt(exon_cortex,stability_cortex,"./Figures for Paper/Reviewer Only Figures/In Vivo FDR Cutoff Bars RefvsAlt Cortex.pdf")
```

##### In Vivo Excitatory Neurons Ref vs Shuf

```{r in-vivo-fdr-data}


plot_fdr_bargraphs_shuf(exon_shuf_neuron,stability_shuf_neuron,"./Figures for Paper/Reviewer Only Figures/In Vivo FDR Cutoff Bars RefvsShuf ExN.pdf")
```

##### In Vivo Excitatory Neurons Ref vs Alt

```{r in-vivo-fdr-data}


plot_fdr_bargraphs_alt(exon_neuron,stability_neuron,"./Figures for Paper/Reviewer Only Figures/In Vivo FDR Cutoff Bars RefvsAlt ExN.pdf")
```

### Volcano Plots
#### In Vitro Data
##### Ref vs Shuf

```{r}

make_post_stab_volcanoplots_refshuf(exon_shuf,stability_shuf,"./Figures for Paper/Reviewer Only Figures/In Vitro Ref vs. Shuf Volcano Plot.pdf")

```

##### Ref vs Alt

```{r}

make_post_stab_volcanoplots_refalt(exon,stability,"./Figures for Paper/Reviewer Only Figures/In Vitro Ref vs. Alt Volcano Plot.pdf")

```

#### In Vivo Data
##### Ref vs Shuf Whole Cortex

```{r}

make_post_stab_volcanoplots_refshuf(exon_shuf_cortex,stability_shuf_cortex,"./Figures for Paper/Reviewer Only Figures/In Vivo Ref vs. Shuf Whole Cortex Volcano Plot.pdf")

```

##### Ref vs Alt Whole Cortex

```{r}

make_post_stab_volcanoplots_refalt(exon_cortex,stability_cortex,"./Figures for Paper/Reviewer Only Figures/In Vivo Ref vs. Alt Whole Cortex Volcano Plot.pdf")

```

##### Ref vs Shuf Excitatory Neurons

```{r}

make_post_stab_volcanoplots_refshuf(exon_shuf_neuron,stability_shuf_neuron,"./Figures for Paper/Reviewer Only Figures/In Vivo Ref vs. Shuf ExN Volcano Plot.pdf")

```

##### Ref vs Alt Excitatory Neurons

```{r}

make_post_stab_volcanoplots_refalt(exon_neuron,stability_neuron,"./Figures for Paper/Reviewer Only Figures/In Vivo Ref vs. Alt ExN Volcano Plot.pdf")

```

### Additions to Supplemental Table 4 for ActinomycinD Assay
#### Read in Data and Calculate Expression from DNA counts from TRAP assay (unchanged Txd DNA)
```{r ActD_Overall_and_Element_Coverage}
# We are going to create some summary stats for TableS4 ----

d <- fread("../../data/formatted/ActD_Trials_08Apr2026/formatted.txt") %>% filter(,!Element %in% badAlt)

#Separate allele names
d <- d %>%
  separate_wider_regex(
    cols = Element,
    patterns = c(
      Element = ".*",        # Match as much as possible (greedy)
      "_",                   # Match the last underscore
      Allele = "[^_]+"       # Match everything remaining that isn't an underscore
    )
  )
d <- as.data.table(d)

#Grab the DNA info
dna <- fread("../../data/formatted/Library1and2BCLess_ISOMPRA_6-30-2022/formatted_Reads_compiled_06-30-2022.tsv") %>% filter(Fraction=="DNA")
dna <- dna %>% group_by(Element,Allele) %>% summarize(CPM=mean(CPM)) %>% as.data.table()
#Put the DNA info in the table, adjoining to correct element
d <- as.data.table(left_join(d,dna,by=c("Element","Allele")))
#Calculate expression
d[, Expression := log2(CPM.x/CPM.y)]
#Get rid of non-finite values
d <- filter(d,is.finite(Expression))
```

#### Coverage (Total Counts/Elements)
```{r}
# 1. Get total unique elements globally
total_elements <- d[, uniqueN(paste(Element, Allele))]

# STEP 1: Summarize per Biological Replicate
d_rep <- d[, .(
  TotalCounts = sum(Counts),
  Items_Above_0 = uniqueN(paste(Element, Allele)[Counts > 0])
), by = .(Fraction, Splicing, BiolRep)]

# Calculate percentage per replicate
d_rep[, PctAbove0 := (Items_Above_0 / total_elements) * 100]

# STEP 2: Average over BiolReps to get your final summary
d_countsummary <- d_rep[, .(
  MeanTotalCounts = mean(TotalCounts),
  sd_TotalCounts  = sd(TotalCounts),
  MeanPctAbove0   = mean(PctAbove0),
  sd_PctAbove0    = sd(PctAbove0),
  Elements        = total_elements
), by = .(Fraction, Splicing)]

# Calculate final coverage
d_countsummary[, Coverage := MeanTotalCounts / Elements]
```

#### CMP and Expression Correlation

```{r CPM_Expression_Correlation}
# 1. Create a temporary copy or just add a modified column for calculation
# Adding a pseudocount of 1 to CPM.x to handle any 0 counts safely
d[, CPM.x_pseudo := CPM.x + 1]

# If your Expression column wasn't already calculated with a pseudocount 
# and has -Inf or NaN from 0 counts, uncomment the line below to recalculate it safely:
d[, Expression_clean := log2(CPM.x_pseudo / (CPM.y + 1))] 

# Step 2: Reshape the data wide so BiolReps become columns (e.g., Rep1, Rep2, Rep3)
# This perfectly pairs up Element/Allele rows across your replicates.
d_wide_cpm  <- dcast(d, Element + Allele + Fraction + Splicing ~ BiolRep, value.var = "CPM.x_pseudo")
d_wide_expr <- dcast(d, Element + Allele + Fraction + Splicing ~ BiolRep, value.var = "Expression_clean")

# Step 3: Identify your replicate column names dynamically (e.g., "Rep1", "Rep2", etc.)
rep_cols <- as.character(unique(d$BiolRep))

# Step 4: Generate all unique pairwise combinations of your replicates (e.g., Rep1 vs Rep2)
rep_pairs <- combn(rep_cols, 2, simplify = FALSE)

# Step 5: Calculate all pairwise correlations within each Fraction/Splicing group
calc_pairwise_corrs <- function(wide_dt) {
  res_list <- list()
  
  for (pair in rep_pairs) {
    repA <- pair[1]
    repB <- pair[2]
    
    # Calculate correlation for this specific pair across Fraction/Splicing groups
    corr_dt <- wide_dt[, .(
      Pair = paste(repA, "vs", repB),
      Corr = cor(.SD[[repA]], .SD[[repB]], method = "spearman", use = "complete.obs")
    ), by = .(Fraction, Splicing)]
    
    res_list[[length(res_list) + 1]] <- corr_dt
  }
  
  return(rbindlist(res_list))
}

# Run the correlation function for both metrics
cpm_corrs  <- calc_pairwise_corrs(d_wide_cpm)
expr_corrs <- calc_pairwise_corrs(d_wide_expr)

# Step 6: Collapse the pairwise correlations to get the final Average and SD
d_cpm_summary <- cpm_corrs[, .(
  Mean_CPM_Correlation = mean(Corr, na.rm = TRUE),
  sd_CPM_Correlation   = sd(Corr, na.rm = TRUE)
), by = .(Fraction, Splicing)]

d_expr_summary <- expr_corrs[, .(
  Mean_Expr_Correlation = mean(Corr, na.rm = TRUE),
  sd_Expr_Correlation   = sd(Corr, na.rm = TRUE)
), by = .(Fraction, Splicing)]

# Step 7: Merge them into one final, clean summary table
d_reproducibility_summary <- merge(d_cpm_summary, d_expr_summary, by = .(Fraction, Splicing))

# Clean up temporary columns from original data table
d[, CPM.x_pseudo := NULL]
```