---
title: "ISOMPRA Paper Figure Making"
author: "Anthony Fischer"
date: "2024-07-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load Tools
```{r}
library(here)
source(here("uISOMPRA/code/mpra_tools","mpra_tools.R"))
library(umap)
library(RColorBrewer)
library(ggforce)
library(fplot)
library(pdftools)
library(gkmSVM)
library(ggbreak)
library(xtable)
library(knitr)
library(readxl)
library(openxlsx)
library(ggrain)
library(tidyverse)
library(limma)
library(dplyr)
cbPalette <- brewer.pal(8,"Dark2")
#A list of elements we do not consider from Library 1.0 as they have bad alts or something
badAlt<-c('ZNF626_wgs','ASIC2_wgs', 'CDKN2B_wgs', 'ELMO1_wgs', 'ENPP5_wgs', 'GDPD4_ssc', 'INSR_wgs', 'LMBRD2_wgs', 'LRP2BP_wgs', 'SCARB2.1_wgs', 'TLK1_wgs', 'TRIM37_wgs', 'ZCCHC11_wgs')
#Set the theme for all ggplots
theme_set(
  theme_classic2(base_family = "sans")+
theme(plot.title=element_text(hjust=0.5, size = 14, face = "bold"),
      plot.subtitle = element_text(hjust=0.5, size = 12, face = "bold"),
  axis.title = element_text(size = 12),
  axis.text = element_text(size = 10),
  legend.title = element_text(size = 14, face = "bold",hjust=0.5),
  legend.text = element_text(size = 10, face = "bold"),
  panel.border = element_rect(color = "black", fill = NA, linewidth = 1)))

dir.create("./Figures")
dir.create("./Tables")
```

##Figure 2
```{r load_data}
#Make the directory
dir.create("./Figures/Figure_2")
d <- fread("../data/formatted/Library1and2BCLess_ISOMPRA_6-30-2022/formatted_Reads_compiled_06-30-2022.tsv") %>% filter(,!Element %in% badAlt)
```

###Figure 2J
```{r figure_2J}
library(tidyverse)

# 1. Harmonize the replicate columns and filter out unwanted comments
d_prepared <- d %>%
  filter(!Comment %in% c("Input Splicing Agnostic mRNA", "TRAP Exonic (mature) mRNA")) %>%
  mutate(BiolRep = na_if(BiolRep, 0)) %>% 
  mutate(UnifiedRep = coalesce(BiolRep, TechRep)) %>%
  select(Element, Allele, Comment, UnifiedRep, CPM)

# 2. Join data against itself, keeping ALL non-self comparisons
d_pairs <- d_prepared %>%
  inner_join(., ., by = c("Element", "Allele", "Comment"), suffix = c("1", "2")) %>%
  # Keeps both Rep1_vs_Rep2 AND Rep2_vs_Rep1, but drops self-comparisons
  filter(UnifiedRep1 != UnifiedRep2)

# 3. Calculate the average correlation for EACH replicate within each Comment
d_cpm_rep_means <- d_pairs %>%
  group_by(Comment, UnifiedRep1, UnifiedRep2) %>%
  summarize(Cor = cor(CPM1, CPM2, method = "spearman"), .groups = "drop") %>%
  
  # Group by UnifiedRep1 to average its correlation against all its peers
  group_by(Comment, UnifiedRep1) %>%
  summarize(MeanCor = mean(Cor), .groups = "drop") %>%
  
  # 4. Apply clean factor labels for plotting
  mutate(Comment = factor(Comment,
    levels = c("Input Intronic (pre) mRNA", "Input Exonic (mature) mRNA", "TRAP Splicing Agnostic mRNA", "Plasmid DNA"),
    labels = c("Pre-Splicing", "Post-Splicing", "TRAP", "Plasmid DNA")
  ))

# 5. Generate the Boxplot (Using MeanCor now)
p7 <- ggplot(d_cpm_rep_means, aes(x = Comment, y = MeanCor, fill = Comment)) +
  geom_boxplot(outliers = FALSE) +
  geom_point(shape = 21, size = 2, position = position_jitter(width = 0.2)) +
  scale_fill_manual(values = c("#7ADCBF", "#21C494", "#178968", "#0A3B2C")) +
  labs(y = "Mean Spearman Correlation per Replicate", title = "CPM Correlation (Replicate Averages)") +
  theme(legend.position = "none", axis.title.x = element_blank()) +
  coord_fixed(ratio = 30)

# 6. Export
export_graph_start("./Figures/Figure_2/cpmcorrelationboxplot_fig2J.pdf")
print(p7)
export_graph_end()
```

###Figure 2K
```{r figure_2K}
library(tidyverse)

# 1. Calculate the average DNA values (Safe against random NAs)
d_dna_avg <- d %>%
  filter(Comment == "Plasmid DNA") %>%
  group_by(Element, Allele) %>%
  summarize(DNA_Mean_CPM = mean(CPM, na.rm = TRUE), .groups = "drop")

# 2. Filter for RNA samples, bring in the DNA baseline, and calculate Expression
d_expression <- d %>%
  filter(Comment %in% c("Input Intronic (pre) mRNA", 
                        "Input Exonic (mature) mRNA", 
                        "TRAP Splicing Agnostic mRNA")) %>%
  left_join(d_dna_avg, by = c("Element", "Allele")) %>%
  
  # Added a small pseudocount (0.01) to protect against log2 of 0 or division by 0
  mutate(Expression = log2((CPM + 0.01) / (DNA_Mean_CPM + 0.01)))

# 3. Harmonize the replicate columns (Done BEFORE factor re-labeling to prevent empty filters)
d_prepared <- d_expression %>%
  mutate(BiolRep = na_if(BiolRep, 0)) %>% 
  mutate(UnifiedRep = coalesce(BiolRep, TechRep)) %>%
  select(Element, Allele, Comment, UnifiedRep, Expression)

# 4. Join data against itself, keeping ALL non-self comparisons
d_pairs <- d_prepared %>%
  inner_join(., ., by = c("Element", "Allele", "Comment"), suffix = c("1", "2")) %>%
  filter(UnifiedRep1 != UnifiedRep2)

# 5. Calculate the average correlation safely
d_exp_rep_means <- d_pairs %>%
  group_by(Comment, UnifiedRep1, UnifiedRep2) %>%
  # use = "complete.obs" ensures it skips rows where Expression is NA
  summarize(Cor = cor(Expression1, Expression2, method = "spearman", use = "complete.obs"), .groups = "drop") %>%
  
  # Group by UnifiedRep1 to average its correlation against all its peers
  group_by(Comment, UnifiedRep1) %>%
  summarize(MeanCor = mean(Cor, na.rm = TRUE), .groups = "drop") %>%
  
  # 6. Apply clean factor labels at the VERY END for safe plotting
  mutate(Comment = factor(Comment,
    levels = c("Input Intronic (pre) mRNA", "Input Exonic (mature) mRNA", "TRAP Splicing Agnostic mRNA"),
    labels = c("Pre-Splicing", "Post-Splicing", "TRAP")
  ))

# 5. Generate the Boxplot (Using MeanCor now)
p_exp <- ggplot(d_exp_rep_means, aes(x = Comment, y = MeanCor, fill = Comment)) +
  geom_boxplot(outliers = FALSE) +
  geom_point(shape = 21, size = 2, position = position_jitter(width = 0.2)) +
  scale_fill_manual(values = c("#7ADCBF", "#21C494", "#178968", "#0A3B2C")) +
  labs(y = "Mean Spearman Correlation per Replicate", title = "Expression Correlation (Replicate Averages)") +
  theme(legend.position = "none", axis.title.x = element_blank()) +
#  coord_fixed(ratio = 30)+
  ylim(min(d_exp_rep_means$MeanCor-0.1),1)

# 6. Export
export_graph_start("./Figures/Figure_2/expcorrelationboxplot_fig2K.pdf")
print(p_exp)
export_graph_end()

```

###Figure 2L
```{r Figure_2L}

umap_CPM <- d %>%
  # 1. Keep only the target fractions
  filter(Comment %in% c("Input Intronic (pre) mRNA",
                        "Input Exonic (mature) mRNA",
                        "Input Splicing Agnostic mRNA",
                        "Plasmid DNA",
                        "TRAP Exonic (mature) mRNA")) %>%
  
  # 2. Convert counts below threshold to NA *before* transforming CPM
  # This stops low-depth noise from influencing the log scale without losing the row structure
  mutate(CPM = if_else(Counts > 49, log2(CPM + 1), NA_real_)) %>% 
  
  # 3. Pivot to wide format cleanly
  pivot_wider(
    id_cols = c(Element, Allele),
    values_from = "CPM",
    names_from = c(Comment, BiolRep, TechRep)
  ) %>% 
  
  # 4. Drop elements that failed the threshold in any single sample fraction
  drop_na()


# 5. Separate data into the paired Input matrix and the standalone matrices
input_metadata <- data.frame(ColName = colnames(umap_CPM[, -c(1, 2)])) %>%
  mutate(
    Fraction = sub("_[0-9]+_[0-9]+$", "", ColName),
    BiolRep  = sub("^.*_([0-9]+)_[0-9]+$", "\\1", ColName)
  )

# Indices for splitting
is_input   <- input_metadata$Fraction %in% c("Input Intronic (pre) mRNA", "Input Exonic (mature) mRNA", "Input Splicing Agnostic mRNA")
is_standalone <- !is_input

# 6. Extract the raw numeric blocks
matrix_all       <- as.matrix(umap_CPM[, -c(1, 2)])
matrix_inputs     <- matrix_all[, is_input]
matrix_standalone <- matrix_all[, is_standalone]

# 7. Correct the Inputs using their true internal pairing structure
meta_inputs <- input_metadata[is_input, ]

# Design protects the differences between Pre, Mature, and Agnostic
design_inputs <- model.matrix(~ 0 + Fraction, data = meta_inputs)
# Batch acts as the pairing factor tracking the 8 RNA pools
batch_inputs  <- meta_inputs$BiolRep

corrected_inputs <- removeBatchEffect(
  matrix_inputs,
  batch = batch_inputs,
  design = design_inputs
)

# 8. Recombine the corrected Inputs with the untouched Plasmid & TRAP baselines
# This preserves the exact baseline variances of controls and target fractions
corrected_combined <- cbind(corrected_inputs, matrix_standalone)

# 9. Restore the original column ordering so your downstream scripts don't flip out
corrected <- corrected_combined[, colnames(matrix_all)]

mds_res_corrected <- limma::plotMDS(corrected)

mds_df_corrected <- data.frame(
  SampleID = colnames(umap_CPM[,-c(1,2)]),
  Dim1 = mds_res_corrected$x,
  Dim2 = mds_res_corrected$y
)

mds_df_corrected <- tidyr::separate(mds_df_corrected,SampleID,c("Library","BiolRep","TechRep"),sep="_")

# Calculate the total variance
total_var <- sum(mds_res_corrected$eigen.values)

# Calculate percentage for Dim 1 and Dim 2
var_pc1 <- (mds_res_corrected$eigen.values[1] / total_var) * 100
var_pc2 <- (mds_res_corrected$eigen.values[2] / total_var) * 100

#Change the Plasmid DNA title
mds_df_corrected$Library[mds_df_corrected$Library=="Plasmid DNA"] <- "Plasmid DNA (Technical Replicates)"

export_graph_start("./Figures/Figure_2/New MDS Plot CPM_Fig2L.pdf")
ggplot(mds_df_corrected,aes(x=Dim1,y=Dim2,fill=Library))+geom_point(shape=21,size=3)+
  geom_polygon(stat = "ellipse",alpha=0.2,color="black")+
  labs(x=paste0("Leading logFC dim 1 (",round(var_pc1,0),"%)"),
       y=paste0("Leading logFC dim 2 (",round(var_pc2,0),"%)"))
export_graph_end()

#Cleanup the environment
rm(list = setdiff(ls(), lsf.str()))
```

###Figure 2N
```{r}
# See "../analysis/Library1_plus_Controls_forPaper_FullRun_31Jul2025/analysis.Rmd"
# for Figure 2N generation
```

##Figure 4
```{r figure_4}
cbPalette <- brewer.pal(8,"Dark2")
dir.create("../figure_making/Figures/Figure_4")
```

###Figure 4A-D
```{r figure_4A-D}
intron_shuf <- fread("../analysis/ActD Trials/Full_Library_08Apr2026/Pre-Splicing_RefvsShuf_Welchs.tsv") %>% .[,plotcolor := as.factor(case_when(
    (abs(logFC)>log2(1.25) & fdr<0.1) ~ 1,
    TRUE ~ 2))]

exon_shuf <- fread("../analysis/ActD Trials/Full_Library_08Apr2026/Post-Splicing_RefvsShuf_Welchs.tsv") %>% .[,plotcolor := as.factor(case_when(
    (abs(logFC)>log2(1.25) & fdr<0.1) ~ 1,
    TRUE ~ 2))]

stability_shuf <- fread("../analysis/ActD Trials/Full_Library_08Apr2026/Stability_RefvsShuf_Welchs.tsv") %>%
  .[,plotcolor := as.factor(case_when(
    (abs(logFC)>log2(1.25) & fdr<0.1) ~ 1,
    TRUE ~ 2))]

translational_shuf <- fread("../analysis/Library1and2BCLess_ISOMPRA_6-30-2022/Translation_Splicing-Agnostic_RefvsShuf_Welchs.tsv") %>% .[,plotcolor := as.factor(case_when(
    (abs(logFC)>log2(1.25) & fdr<0.1) ~ 1,
    TRUE ~ 2))]

p1_shuf <- ggplot(intron_shuf)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor,alpha = plotcolor),shape=21,size=2)+
  scale_fill_manual(values=c(cbPalette[1],"lightgray"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  scale_alpha_manual(values = c(1,1,0.1,0.1))+
  labs(fill="Expression Change",
       title="Reference vs. Shuffled Expression Change\nin Pre-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.05),linetype="dashed")+
  theme(legend.position = "none",aspect.ratio=1)

p2_shuf <- ggplot(exon_shuf)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor),shape=21,size=2)+
  scale_fill_manual(values=c(cbPalette[1],"lightgray"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  labs(fill="Expression Change",
       title="Reference vs. Shuffled Expression Change\nin Post-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.05),linetype="dashed")+
  theme(legend.position = "none",aspect.ratio=1)

p3_shuf <- ggplot(stability_shuf)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor),shape=21,size=2)+
  scale_fill_manual(values=c(cbPalette[1],"lightgrey"),
                    labels=c("Increased Stability","Decreased Stability",
                             "Not Significant"))+
  labs(fill="Stability Change",
       title="Reference vs. Shuffled Stability Change",
       x="Log2 Fold Change in Stability",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.05),linetype="dashed")+
  theme(legend.position = "none",aspect.ratio=1)

p4_shuf <- ggplot(translational_shuf)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor),shape=21,size=2)+
  scale_fill_manual(values=c(cbPalette[1],"lightgrey"),
                    labels=c("Increased Translation","Decreased Translation",
                             "Not Significant"))+
  labs(fill="Translational Change",
       title="Reference vs. Shuffled Translational Change",
       x="Log2 Fold Change in Translation",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.05),linetype="dashed")+
  theme(legend.position = "none",aspect.ratio=1)

export_graph_start("./Figures/Figure_4/AllRefvsShuf_VolcanoPlot_Patchwork_fig4A-D.pdf",sideways = TRUE,height = 0.5)
print(p1_shuf+theme(legend.position = "none")+labs(title=NULL,subtitle="Pre-Splicing",x="Log2FC Expression")+
        p2_shuf+theme(legend.position = "none")+labs(title=NULL,subtitle="Post-Splicing",x="Log2FC Expression")+
        p3_shuf+theme(legend.position = "none")+labs(title=NULL,subtitle="Stability",x="Log2FC Stability")+
        p4_shuf+theme(legend.position = "none")+labs(title=NULL,subtitle="Translational",x="Log2FC TE")+
        plot_layout(ncol=4)+plot_annotation(title="Reference vs. Shuffled"))
export_graph_end()
```

###Figure 4G
```{r figure_4G}
# The code for figure 4G must be run before running the code for 4F
#False positive analysis
exon_stability <- inner_join(exon_shuf,stability_shuf,"Element")

#Significant in both and same direction
true_positives <- filter(exon_stability,
                         fdr.x<0.1,
                         fdr.y<0.1,
                         ((logFC.x>0 & logFC.y>0) | (logFC.x<0 & logFC.y<0))
                         )$Element
#Only significant in stability, not post-splicing
false_negatives <- filter(exon_stability,
                         fdr.x>0.1,
                         fdr.y<0.1)$Element
#Either only significant in post-splicing, or significant in both but opposite
#directions
false_positives <- c(
  filter(exon_stability,
         fdr.x<0.1,
         fdr.y>0.1)$Element,
  filter(exon_stability,
         fdr.x<0.1,
         fdr.y<0.1,
         ((logFC.x>0 & logFC.y<0) | (logFC.x<0 & logFC.y>0)))$Element)
  




all_else <- filter(stability_shuf,
                   !Element %in% false_positives,
                   !Element %in% false_negatives,
                   !Element %in% true_positives)$Element

plottable <- rbind(
  data.table("Element"=true_positives,"Category"="True Positive"),
  data.table("Element"=false_positives,"Category"="False Positive"),
  data.table("Element"=false_negatives,"Category"="False Negative"),
  data.table("Element"=all_else,"Category"="Not Significant")
)

plottable[,Category := factor(Category,levels=c("True Positive","False Positive","False Negative","Not Significant"))]

summary_data <- plottable %>%
  group_by(Category) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

# Separate the data for labels
label_data <- summary_data %>%
  filter(Category != "Not Significant") %>%
  mutate(cumulative_percentage = 100 - cumsum(percentage) + percentage/2)

# Create the plot

p5_shuf <- ggplot(summary_data, aes(x = "", y = percentage, fill = Category)) +
  geom_col(width = 0.4, color = "black") +
  scale_fill_manual(values = c("Not Significant" = "lightgrey",
                               "True Positive" = "#136d52",
                               "False Positive" = "#21c494",
                               "False Negative" = "#67e5bf")) +
  labs(y = "Percentage",
       fill = "Category",
       x=NULL) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        legend.position = "right",
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

export_graph_start("./Figures/Figure_4/RefShuf_Accuracy_fig4G.pdf",width=0.75)
print(p5_shuf+theme_minimal())
ggplot(summary_data,aes(Category,count,fill=as.factor(count)))+
  geom_col()+
  geom_text(aes(y=count-100,label=paste0(round(percentage,2),"\n",count)))+
  theme(axis.text.x = element_text(angle = 45,hjust=1))
export_graph_end()
```

###Figure 4F
```{r figure_4F}
exon_stability <- left_join(exon_stability,plottable)
exon_stability <- unique.data.frame(exon_stability)
p5_shuf_dotplot <- ggplot(exon_stability,aes(x=logFC.x,y=logFC.y,fill=Category))+
  geom_point(shape=21,alpha=0.5,size=3)+
  scale_fill_manual(values = c("Not Significant" = "lightgrey",
                               "True Positive" = "#136d52",
                               "False Positive" = "#21c494",
                               "False Negative" = "#67e5bf"))+
  labs(x="Log Fold Change\n(Post-Splice)",
       y="Log Fold Change\n(Stability)",
       )+
  theme_bw()+
  theme(legend.position = "none",aspect.ratio = 1)
print(p5_shuf_dotplot)

export_graph_start("./Figures/Figure_4/logfcpostvsstab_refshuf_fig4F.pdf")
print(p5_shuf_dotplot + coord_fixed(ratio=1.979))
export_graph_end()
```

###Figure 4H-K
```{figure4_h-k}
intron <- fread("../analysis/ActD Trials/Full_Library_08Apr2026/PreSplicing_RefAlt_LM.txt") %>% .[,plotcolor := as.factor(case_when(
    (abs(logFC)>log2(1.25) & fdr<0.1) ~ 1,
    TRUE ~ 2))]

exon <- fread("../analysis/ActD Trials/Full_Library_08Apr2026/PostSplicing_RefAlt_LM.txt") %>% .[,plotcolor := as.factor(case_when(
    (abs(logFC)>log2(1.25) & fdr<0.1) ~ 1,
    TRUE ~ 2))]

stability <- fread("../analysis/ActD Trials/Full_Library_08Apr2026/Stability_RefAlt_LM.txt") %>%
  mutate(logFC_interaction = -(logFC_interaction)) %>%
  .[,plotcolor := as.factor(case_when(
    (abs(logFC_interaction)>log2(1.25) & fdr_interaction<0.1) ~ 1,
    TRUE ~ 2))]

translational <- fread("../analysis/Library1and2BCLess_ISOMPRA_6-30-2022/Translation_Splicing-Agnostic_RefvsAlt_LM_Annotated.tsv") %>% .[,plotcolor := as.factor(case_when(
    (abs(logFC)>log2(1.25) & fdr<0.1) ~ 1,
    TRUE ~ 2))]

# Define some points we want to label on the plots
# Specifically, intron points that are transcriptional only,
# exon points that are a mix of transcriptional only and stability sig
# all or nearly all the stability points
# all the translational points

intron_elems <- filter(stability, fdr_allele < 0.1, fdr_interaction > 0.1) %>% 
  as.data.table() %>% 
  .[order(abs(fdr_allele))] %>% 
  head(4) %>% pull(Element)

exon_elems <- filter(stability, fdr_interaction < 0.1) %>% 
  as.data.table() %>% 
  .[order(abs(fdr_allele))] %>% 
  head(4) %>% pull(Element)

stability_elems <- filter(stability,fdr_interaction<0.1) %>% as.data.table() %>%
  .[order(abs(fdr_interaction))] %>% head(10) %>% pull(Element)

translational_elems <- filter(translational,fdr<0.1) %>% as.data.table() %>%
  .[order(-abs(logFC))] %>% head(10) %>% pull(Element)

p1 <- ggplot(intron)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor,shape=Pheno),size=2)+
  scale_fill_manual(values=c(cbPalette[1],"lightgray"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  scale_shape_manual(values = c(21,23))+
  labs(fill="Expression Change",
       title="Reference vs. Variant Expression Change\nin Pre-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")+
  theme(legend.position = "none",aspect.ratio = 1)+
  geom_text_repel(data = filter(intron,Element %in% intron_elems),
                  aes(x=logFC,y=-log10(fdr),label=Symbol))
print(p1)

p2 <- ggplot(exon)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor,shape=Pheno),size=2)+
  scale_fill_manual(values=c(cbPalette[1],"lightgray"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  scale_shape_manual(values=c(21,23))+
  labs(fill="Expression Change",
       title="Reference vs. Variant Expression Change\nin Post-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")+
  theme(legend.position = "none",aspect.ratio = 1)+
  geom_text_repel(data = filter(exon,(Element %in% intron_elems | Element %in% exon_elems)),
                  aes(x=logFC,y=-log10(fdr),label=Symbol))
print(p2)

p3 <- ggplot(stability)+
  geom_point(aes(x=logFC_interaction,y=-log10(fdr_interaction),fill=plotcolor,shape=Pheno),size=2)+
  scale_fill_manual(values=c(cbPalette[1],"lightgrey"),
                    labels=c("Increased Stability","Decreased Stability",
                             "Not Significant"))+
  scale_shape_manual(values=c(21,23))+
  labs(fill="Stability Change",
       title="Reference vs. Variant Stability Change",
       x="Log2 Fold Change in Stability",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")+
  theme(legend.position = "none",aspect.ratio = 1)+
  geom_text_repel(data = filter(stability,Element %in% stability_elems),
                  aes(x=logFC_interaction,y=-log10(fdr_interaction),label=Symbol))
print(p3)

p4 <- ggplot(translational)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor,shape = Pheno),size=2)+
  scale_fill_manual(values=c(cbPalette[1],"lightgrey"),
                    labels=c("Increased Translation","Decreased Translation",
                             "Not Significant"))+
  scale_shape_manual(values=c(21,23))+
  labs(fill="Translational Change",
       title="Reference vs. Variant Translational Change",
       x="Log2 Fold Change in Translation",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")+
  guides(fill="none")+
  theme(aspect.ratio = 1)+
  geom_text_repel(data = filter(translational,Element %in% translational_elems),
                  aes(x=logFC,y=-log10(fdr),label=Symbol))+
  labs(shape="Phenotype")
print(p4)

export_graph_start("./Figures/Figure_4/AllRefvsAlt_VolcanoPlot_Patchwork_New_fig_4h-k.pdf",sideways = TRUE,height = 0.5)
print(p1+theme(legend.position = "none")+labs(title=NULL,subtitle="Pre-Splicing",x="Log2FC Expression")+
        p2+theme(legend.position = "none")+labs(title=NULL,subtitle="Post-Splicing",x="Log2FC Expression")+
        p3+theme(legend.position = "none")+labs(title=NULL,subtitle="Stability",x="Log2FC Stability")+
        p4+theme(legend.position = "none")+labs(title=NULL,subtitle="Translational",x="Log2FC TE")+
        plot_layout(ncol=4)+plot_annotation(title="Reference vs. Variant"))
export_graph_end()
```

###Figure 4M
```{figure_4M}
# As before, 4M must be run before 4L
#False positive analysis
exon_stability <- inner_join(exon,stability,"Element")

#Significant in both and same direction
true_positives <- filter(exon_stability,
                         fdr<0.1,
                         fdr_interaction<0.1,
                         ((logFC>0 & logFC_interaction>0) | (logFC<0 & logFC_interaction<0))
                         )$Element
#Only significant in stability, not post-splicing
false_negatives <- filter(exon_stability,
                         fdr>0.1,
                         fdr_interaction<0.1)$Element
#Either only significant in post-splicing, or significant in both but opposite
#directions
false_positives <- c(
  filter(exon_stability,
         fdr<0.1,
         fdr_interaction>0.1)$Element,
  filter(exon_stability,
         fdr<0.1,
         fdr_interaction<0.1,
         ((logFC>0 & logFC_interaction<0) | (logFC<0 & logFC_interaction>0)))$Element)
  




all_else <- filter(stability_shuf,
                   !Element %in% false_positives,
                   !Element %in% false_negatives,
                   !Element %in% true_positives)$Element

plottable <- rbind(
  data.table("Element"=true_positives,"Category"="True Positive"),
  data.table("Element"=false_positives,"Category"="False Positive"),
  data.table("Element"=false_negatives,"Category"="False Negative"),
  data.table("Element"=all_else,"Category"="Not Significant")
)

plottable[,Category := factor(Category,levels=c("True Positive","False Positive","False Negative","Not Significant"))]

summary_data <- plottable %>%
  group_by(Category) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

# Separate the data for labels
label_data <- summary_data %>%
  filter(Category != "Not Significant") %>%
  mutate(cumulative_percentage = 100 - cumsum(percentage) + percentage/2)

# Create the plot

p5 <- ggplot(summary_data, aes(x = "", y = percentage, fill = Category)) +
  geom_col(width = 0.4, color = "black") +
  scale_fill_manual(values = c("Not Significant" = "lightgrey",
                               "True Positive" = "#136d52",
                               "False Positive" = "#21c494",
                               "False Negative" = "#67e5bf")) +
  labs(y = "Percentage",
       fill = "Category",
       x=NULL) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        legend.position = "right",
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

export_graph_start("./Figures/Figure_4/RefAlt_Accuracy_fig4M.pdf",width=0.75)
print(p5+theme_minimal())
ggplot(summary_data,aes(Category,count,fill=as.factor(count)))+
  geom_col()+
  geom_text(aes(y=count-100,label=paste0(round(percentage,2),"\n",count)))+
  theme(axis.text.x = element_text(angle = 45,hjust=1))
export_graph_end()
```

###Figure 4L
```{r figure_4L}
exon_stability <- left_join(exon_stability,plottable)
exon_stability <- unique.data.frame(exon_stability)
p5_dotplot <- ggplot(exon_stability,aes(x=logFC,y=logFC_interaction,fill=Category))+
  geom_point(shape=21,alpha=0.5,size=3)+
  scale_fill_manual(values = c("Not Significant" = "lightgrey",
                               "True Positive" = "#136d52",
                               "False Positive" = "#21c494",
                               "False Negative" = "#67e5bf"))+
  labs(x="Log Fold Change\n(Post-Splice)",
       y="Log Fold Change\n(Stability)",
       )+
  theme_bw()+
  theme(legend.position = "none",aspect.ratio = 1)
print(p5_dotplot)

export_graph_start("./Figures/Figure_4/logfcpostvsstab_refvsalt_4L.pdf")
print(p5_dotplot)
export_graph_end()
```

###Figure 4N-P
```{r figure4_N-P}
badAlt<-c('ZNF626_wgs','ASIC2_wgs', 'CDKN2B_wgs', 'ELMO1_wgs', 'ENPP5_wgs', 'GDPD4_ssc', 'INSR_wgs', 'LMBRD2_wgs', 'LRP2BP_wgs', 'SCARB2.1_wgs', 'TLK1_wgs', 'TRIM37_wgs', 'ZCCHC11_wgs')
d <- fread("../data/formatted/ActD_Trials_08Apr2026/formatted.txt") %>% filter(,!Element %in% badAlt)

#Separate allele names
d <- d %>%
  separate_wider_regex(
    cols = Element,
    patterns = c(
      Element = ".*",        # Match as much as possible (greedy)
      "_",                   # Match the last underscore
      Allele = "[^_]+"       # Match everything remaining that isn't an underscore
    )
  )
d <- as.data.table(d)

#Grab the DNA info
dna <- fread("../data/formatted/Library1and2BCLess_ISOMPRA_6-30-2022/formatted_Reads_compiled_06-30-2022.tsv") %>% filter(Fraction=="DNA")
dna <- dna %>% group_by(Element,Allele) %>% summarize(CPM=mean(CPM)) %>% as.data.table()
#Put the DNA info in the table, adjoining to correct element
d <- as.data.table(left_join(d,dna,by=c("Element","Allele")))
#Calculate expression
d[, Expression := log2(CPM.x/CPM.y)]
#Get rid of non-finite values
d <- filter(d,is.finite(Expression))
d <- filter(d,Fraction=="0hr",!(Splicing=="Pre" & BiolRep==3))
data_for_concordancy <- d #For later Figure S6
common_y <- c(
  min(filter(d,Element=="MLXIP_wgs", is.finite(Expression))$Expression),
  max(filter(d,Element=="MLXIP_wgs", is.finite(Expression))$Expression)
)

library(ggtext)
#Grab GALNTL5 variant, that was a good one
pre_boxplot <- ggplot(
  filter(d,Element=="MLXIP_wgs",Splicing=="pre") %>%
    mutate(Splicing=factor(Splicing,levels=c("pre","post"),labels=c("Pre","Post"))) %>%
    mutate(Allele=factor(Allele,levels=c("ref","alt"),labels=c("Reference","Variant"))),
  aes(x=Allele,y=Expression,fill=Allele))+
  geom_boxplot()+
  geom_point(shape=21)+
  scale_fill_manual(values = c("#136d52","#67e5bf"))+
  guides(fill="none")+
  theme(plot.title = element_text(hjust=0.5))+
  labs(title="Pre-Splicing",y = "Expression: Log<sub>2</sub> (<sup>RNA</sup>&frasl;<sub>DNA</sub>)") +
  theme(axis.title.y = element_markdown(),aspect.ratio = 2)+
  ylim(common_y)
print(pre_boxplot)

post_boxplot <- ggplot(
  filter(d,Element=="MLXIP_wgs",Splicing=="post") %>%
    mutate(Splicing=factor(Splicing,levels=c("pre","post"),labels=c("Pre","Post"))) %>%
    mutate(Allele=factor(Allele,levels=c("ref","alt"),labels=c("Reference","Variant"))),
  aes(x=Allele,y=Expression,fill=Allele))+
  geom_boxplot()+
  geom_point(shape=21)+
  scale_fill_manual(values = c("#136d52","#67e5bf"))+
  guides(fill="none")+
  theme(plot.title = element_text(hjust=0.5))+
  labs(title="Post-Splicing", y = "Expression: Log<sub>2</sub> (<sup>RNA</sup>&frasl;<sub>DNA</sub>)")+
  theme(axis.title.y = element_markdown())+
  theme(axis.title.y = element_blank(),aspect.ratio = 2)+
  ylim(common_y)
print(post_boxplot)

stab_boxplot <- ggplot(
  filter(d, Element == "MLXIP_wgs")%>%
    mutate(Splicing=factor(Splicing,levels=c("pre","post"),labels=c("Pre","Post"))) %>%
    mutate(Allele=factor(Allele,levels=c("ref","alt"),labels=c("Reference","Variant"))),
  aes(x = Splicing, y = Expression, fill = Allele)
) +
  # 1. Four distinct boxplots
  # We use position_dodge to separate Ref and Alt at each x-axis tick
  geom_boxplot(
    aes(group = interaction(Splicing, Allele)),
    position = position_dodge(width = 0.8),
    outlier.shape = NA
  ) +
  
  # 2. Dots centered over each of the four boxes
  geom_point(
    aes(group = interaction(Splicing, Allele)),
    position = position_dodge(width = 0.8), 
    shape = 21, 
    color = "black"
  ) +
  
  # 3. Lines connecting Pre-Ref to Post-Ref and Pre-Alt to Post-Alt
  # We use group = Allele here so it ignores the Splicing split for the line
  stat_summary(
    fun = median, 
    geom = "line", 
    aes(group = Allele, color = Allele), 
    linewidth = 1,
    position = position_dodge(width = 0.8) # Matches the boxplot shift
  ) +
  
  scale_fill_manual(values = c("#136d52", "#67e5bf")) +
  scale_color_manual(values = c("#136d52", "#67e5bf"))+
  guides(fill="none",color="none")+
  theme(plot.title = element_text(hjust=0.5))+
  labs(title="Stability",y = "Expression: Log<sub>2</sub> (<sup>RNA</sup>&frasl;<sub>DNA</sub>)") +
  theme(axis.title.y = element_markdown())+
  theme(axis.title.y = element_blank(),aspect.ratio = 2)+
  ylim(common_y)
print(stab_boxplot)

export_graph_start("./Figures/Figure_4/Figure_4N-P.pdf")
pre_boxplot+post_boxplot+stab_boxplot
export_graph_end()
```

##Case vs Control Bias in Singificant Variants
###Pre-Splicing
```{r case_control_bias_pre}

mat <- matrix(
  c(length(filter(intron,fdr<0.1,Pheno=="case")$Element),
    length(filter(intron,fdr<0.1,Pheno=="control")$Element),
    length(filter(intron,fdr>0.1,Pheno=="case")$Element),
    length(filter(intron,fdr>0.1,Pheno=="control")$Element)), 
  nrow = 2,
  dimnames = list(
    Variant = c("Case", "Control"),
    Status = c("Significant", "NonSignificant")
  )
)
fisher.test(mat,alternative = "greater")
```

###Post-Splicing
```{r case_control_bias_post}
mat <- matrix(
  c(length(filter(exon,fdr<0.1,Pheno=="case")$Element),
    length(filter(exon,fdr<0.1,Pheno=="control")$Element),
    length(filter(exon,fdr>0.1,Pheno=="case")$Element),
    length(filter(exon,fdr>0.1,Pheno=="control")$Element)), 
  nrow = 2,
  dimnames = list(
    Variant = c("Case", "Control"),
    Status = c("Significant", "NonSignificant")
  )
)
fisher.test(mat,alternative = "greater")
```

###Stability
```{r case_control_bias_stability}
mat <- matrix(
  c(length(filter(stability,fdr_interaction<0.1,Pheno=="case")$Element),
length(filter(stability,fdr_interaction<0.1,Pheno=="control")$Element),
length(filter(stability,fdr_interaction>0.1,Pheno=="case")$Element),
length(filter(stability,fdr_interaction>0.1,Pheno=="control")$Element)), 
  nrow = 2,
  dimnames = list(
    Variant = c("Case", "Control"),
    Status = c("Significant", "NonSignificant")
  )
)

fisher.test(mat,alternative = "greater")
```

###Transcriptional-Only
```{r case_control_bias_transcriptional}
mat <- matrix(
  c(
    length(filter(stability,fdr_interaction>0.1,fdr_allele<0.1,Pheno=="case")$Element),
    length(filter(stability,fdr_interaction>0.1,fdr_allele<0.1,Pheno=="control")$Element),
    length(filter(stability,fdr_interaction>0.1,fdr_allele>0.1,Pheno=="case")$Element),
    length(filter(stability,fdr_interaction>0.1,fdr_allele>0.1,Pheno=="control")$Element)), 
  nrow = 2,
  dimnames = list(
    Variant = c("Case", "Control"),
    Status = c("Significant", "NonSignificant")
  )
)

fisher.test(mat,alternative = "greater")
#Cleanup
rm(list = setdiff(ls(), lsf.str()))

```

##Figure 6

```{r create_dir}
dir.create("./Figures/Figure_6")
```

###Figure 6B
```{r figure_6b}
cbPalette <- brewer.pal(8,"Dark2")
d <- fread("../data/formatted/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/formatted_Reads_08Mar2024.tsv") %>% filter(Cre_Type=="VGlut1") %>% mutate(CPM=log2(CPM+1))

d_wide <- pivot_wider(d,
                      id_cols = c("Element","Allele"),
                      names_from = c("Splicing_Pool","Cre_Status","Animal_Hemisphere"),
                      values_from = CPM)

# 1. Isolate the numeric values (skipping 'Element' and 'Allele' columns)
# If column 1 is an empty row name string "", we drop columns 1, 2, and 3
data_matrix <- as.matrix(d_wide[, -c(1, 2)])

# 2. Extract metadata directly from the wide column names
col_metadata <- data.frame(ColName = colnames(data_matrix)) %>%
  mutate(
    # Split the names by the underscores
    Splicing  = sub("^([^_]+)_.*$", "\\1", ColName),     # "Pre" or "Post"
    CreStatus = sub("^[^_]+_([^_]+)_.*$", "\\1", ColName), # "ON" or "OFF"
    BiolRep   = sub("^[^_]+_[^_]+_(.*)$", "\\1", ColName)   # "10L", "10R", "1L", etc.
  ) %>%
  # Combine Splicing and CreStatus into your 4 true biological treatment groups
  mutate(BioGroup = paste(Splicing, CreStatus, sep = "_"))

# 3. Create the protective design matrix
# This shields your 4 biological conditions from being crushed or flattened
design_matrix <- model.matrix(~ 0 + BioGroup, data = col_metadata)

# 4. Define the batch factor using the true physical RNA pool source
# This tells limma to pair columns that share the exact same tissue tube (e.g., "10L")
batch_factor <- col_metadata$BiolRep

# 5. Run the unconfounded batch correction
corrected_matrix <- removeBatchEffect(
  data_matrix,
  batch = batch_factor,
  design = design_matrix
)

mds_invivo <- limma::plotMDS(corrected_matrix)

mds_invivo_df <- data.frame(
  SampleID = colnames(corrected_matrix),
  Dim1 = mds_invivo$x,
  Dim2 = mds_invivo$y
)

mds_invivo_df <- tidyr::separate(mds_invivo_df,SampleID,c("Splicing","Cre","BiolRep"),sep="_")
mds_invivo_df <- tidyr::unite(mds_invivo_df,"Library",c("Splicing","Cre")) %>% as.data.table()
mds_invivo_df[,Library := factor(Library,
                                 levels = c(
                                   "Pre_OFF",
                                   "Post_OFF",
                                   "Pre_ON",
                                   "Post_ON"),
                                 labels = c(
                                   "Cre(-) Cells,\nPre-Splicing",
                                   "Cre(-) Cells,\nPost-Splicing",
                                   "Cre(+) Cells,\nPre-Splicing",
                                   "Cre(+) Cells,\nPost-Splicing"
                                 ))]

# Calculate the total variance
total_var <- sum(mds_invivo$eigen.values)

# Calculate percentage for Dim 1 and Dim 2
var_pc1 <- (mds_invivo$eigen.values[1] / total_var) * 100
var_pc2 <- (mds_invivo$eigen.values[2] / total_var) * 100



export_graph_start("./Figures/Figure_6/MDS Plot log2 CPM Batch Corrected_fig_6b.pdf")
ggplot(mds_invivo_df,aes(x=Dim1,y=Dim2,fill=Library))+geom_point(shape=21,size=3)+
  geom_polygon(stat = "ellipse",alpha=0.2,color="black")+
  labs(x=paste0("Leading logFC dim 1 (",round(var_pc1,0),"%)"),
       y=paste0("Leading logFC dim 2 (",round(var_pc2,0),"%)"))
export_graph_end()
```

###Figure 6C-E
```{r figure_6c-e}

shufdata <- rbind(
  fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Pre-Splicing_Vglut1_CreOFF_RefvsShuf_Welchs.tsv") %>% mutate(Splicing_Pool="Pre",Cre_Status="Off"),
  fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Pre-Splicing_Vglut1_CreON_RefvsShuf_Welchs.tsv") %>% mutate(Splicing_Pool="Pre",Cre_Status="On"),
fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Post-Splicing_Vglut1_CreOFF_RefvsShuf_Welchs.tsv") %>% mutate(Splicing_Pool="Post",Cre_Status="Off"),
fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Post-Splicing_Vglut1_CreON_RefvsShuf_Welchs.tsv") %>% mutate(Splicing_Pool="Post",Cre_Status="On"),
fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreOFF_RefvsShuf_Welchs.tsv") %>% mutate(Splicing_Pool="Stability",Cre_Status="Off"),
fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreON_RefvsShuf_Welchs.tsv") %>% mutate(Splicing_Pool="Stability",Cre_Status="On"))
shufdata <- shufdata %>% .[,plotcolor := as.factor(case_when(
    (logFC>log2(1.25) & fdr<0.1) ~ 1,
    (logFC<log2(0.75) & fdr<0.1) ~ 2,
    TRUE ~ 3))]
shufdata[,Splicing_Pool := factor(Splicing_Pool,levels = c("Pre","Post","Stability"))]
shufdata[,Cre_Status := factor(Cre_Status,levels = c("Off","On"),labels = c("Whole Cortex","Excitatory Neurons"))]

p1_shuf_OFF <- ggplot(filter(shufdata,Splicing_Pool=="Pre",Cre_Status=="Whole Cortex"))+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor),shape=21,size=2)+
  scale_fill_manual(values=c(cbPalette[1],cbPalette[1],"lightgrey"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  labs(fill="Expression Change",
       title="Variant vs. Reference Expression Change\nin Pre-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")
p1_shuf_OFF

p2_shuf_OFF <- ggplot(filter(shufdata,Splicing_Pool=="Post",Cre_Status=="Whole Cortex"))+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor),shape=21,size=2)+
  scale_fill_manual(values=c(cbPalette[1],cbPalette[1],"lightgrey"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  labs(fill="Expression Change",
       title="Variant vs. Reference Expression Change\nin Pre-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")
p2_shuf_OFF

p3_shuf_OFF <- ggplot(filter(shufdata,Splicing_Pool=="Stability",Cre_Status=="Whole Cortex"))+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor),shape=21,size=2)+
  scale_fill_manual(values=c(cbPalette[1],cbPalette[1],"lightgrey"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  labs(fill="Expression Change",
       title="Variant vs. Reference Expression Change\nin Pre-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")
p3_shuf_OFF

export_graph_start("./Figures/Figure_6/RefvsShuf_All_CreOFF_fig6c-e.pdf",sideways=TRUE,height=0.5)
print(p1_shuf_OFF+theme(legend.position = "none")+labs(title=NULL,subtitle="Pre-Splicing",x="Log2FC Expression")+
        p2_shuf_OFF+theme(legend.position = "none")+labs(title=NULL,subtitle="Post-Splicing",x="Log2FC Expression")+
        p3_shuf_OFF+theme(legend.position = "none")+labs(title=NULL,subtitle="Stability",x="Log2FC Stability")+
        plot_layout(ncol=3)+plot_annotation(title="Reference vs. Shuffled"))
export_graph_end()
```

###Figure 6F
```{r figure_6F}
#False positive analysis
exon_stability <- inner_join(
  filter(shufdata,Splicing_Pool=="Post",Cre_Status=="Whole Cortex"),
  filter(shufdata,Splicing_Pool=="Stability",Cre_Status=="Whole Cortex"),
  "Element")


#Significant in both and same direction
true_positives <- filter(exon_stability,
                         fdr.x<0.1,
                         fdr.y<0.1,
                         ((logFC.x>0 & logFC.y>0) | (logFC.x<0 & logFC.y<0))
                         )$Element
#Only significant in stability, not post-splicing
false_negatives <- filter(exon_stability,
                         fdr.x>0.1,
                         fdr.y<0.1)$Element
#Either only significant in post-splicing, or significant in both but opposite
#directions
false_positives <- c(
  filter(exon_stability,
         fdr.x<0.1,
         fdr.y>0.1)$Element,
  filter(exon_stability,
         fdr.x<0.1,
         fdr.y<0.1,
         ((logFC.x>0 & logFC.y<0) | (logFC.x<0 & logFC.y>0)))$Element)
  




all_else <- filter(exon_stability,
                   !Element %in% false_positives,
                   !Element %in% false_negatives,
                   !Element %in% true_positives)$Element

plottable <- rbind(
  data.table("Element"=true_positives,"Category"="True Positive"),
  data.table("Element"=false_positives,"Category"="False Positive"),
  data.table("Element"=false_negatives,"Category"="False Negative"),
  data.table("Element"=all_else,"Category"="Not Significant")
)

plottable[,Category := factor(Category,levels=c("True Positive","False Positive","False Negative","Not Significant"))]

summary_data <- plottable %>%
  group_by(Category) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

# Separate the data for labels
label_data <- summary_data %>%
  filter(Category != "Not Significant") %>%
  mutate(cumulative_percentage = 100 - cumsum(percentage) + percentage/2)

# Create the plot

p5_shuf <- ggplot(summary_data, aes(x = "", y = percentage, fill = Category)) +
  geom_col(width = 0.4, color = "black") +
  scale_fill_manual(values = c("Not Significant" = "lightgrey",
                               "True Positive" = "#136d52",
                               "False Positive" = "#21c494",
                               "False Negative" = "#67e5bf")) +
  labs(y = "Percentage",
       fill = "Category",
       x=NULL) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        legend.position = "right",
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

export_graph_start("./Figures/Figure_6/RefShuf_Accuracy_CreOFF_fig6f.pdf",width=0.75)
print(p5_shuf+theme_minimal())
ggplot(summary_data,aes(Category,count,fill=as.factor(count)))+
  geom_col()+
  geom_text(aes(y=count-100,label=paste0(round(percentage,2),"\n",count)))+
  theme(axis.text.x = element_text(angle = 45,hjust=1))
export_graph_end()

exon_stability <- left_join(exon_stability,plottable)
exon_stability <- unique.data.frame(exon_stability)
cortex_shuf_dotplot <- ggplot(exon_stability,aes(x=logFC.x,y=logFC.y,fill=Category))+
  geom_point(shape=21,alpha=0.5,size=3)+
  scale_fill_manual(values = c("Not Significant" = "lightgrey",
                               "True Positive" = "#136d52",
                               "False Positive" = "#21c494",
                               "False Negative" = "#67e5bf"))+
  labs(x="Log Fold Change\n(Post-Splice)",
       y="Log Fold Change\n(Stability)",
       )+
  theme_bw()+
  theme(legend.position = "none")
print(cortex_shuf_dotplot)
export_graph_start("./Figures/Figure_6/Cortex_RefShuf_PostvsStability_DotPlot_fig6f.pdf")
print(cortex_shuf_dotplot + 
        coord_fixed(ratio = 
    diff(range(exon_stability$logFC.x[is.finite(exon_stability$logFC.x)])) / 
    diff(range(exon_stability$logFC.y[is.finite(exon_stability$logFC.y)]))))
export_graph_end()
```

###Figure 6G-I
```{r figure6_g-i}
#Read in LM results and set a plotting value based on significance and logFC
intron_OFF <- filter(fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Pre-Splicing_Vglut1_CreOFF_RefvsAlt_LM_Annotated.tsv"),is.finite(fdr)) %>%
  .[,plotcolor := as.factor(case_when(
    (logFC>log2(1.25) & fdr<0.1) ~ 1,
    (logFC<-log2(1.25) & fdr<0.1) ~ 2,
    TRUE ~ 3))]

exon_OFF <- filter(fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Post-Splicing_Vglut1_CreOFF_RefvsAlt_LM_Annotated.tsv"),is.finite(fdr))%>%
  .[,plotcolor := as.factor(case_when(
    (logFC>log2(1.25) & fdr<0.1) ~ 1,
    (logFC<-log2(1.25) & fdr<0.1) ~ 2,
    TRUE ~ 3))]

stability_OFF <- filter(fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreOFF_RefvsAlt_LM_Annotated.tsv"),is.finite(fdr_interaction)) %>%
  mutate(logFC_interaction = -(logFC_interaction)) %>%
  .[,plotcolor := as.factor(case_when(
    (logFC_interaction>log2(1.25) & fdr_interaction<0.1) ~ 1,
    (logFC_interaction<-log2(1.25) & fdr_interaction<0.1) ~ 2,
    TRUE ~ 3))]

intron_ON <- filter(fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Pre-Splicing_Vglut1_CreON_RefvsAlt_LM_Annotated.tsv"),is.finite(fdr)) %>%
  .[,plotcolor := as.factor(case_when(
    (logFC>log2(1.25) & fdr<0.1) ~ 1,
    (logFC<-log2(1.25) & fdr<0.1) ~ 2,
    TRUE ~ 3))]

exon_ON <- filter(fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Post-Splicing_Vglut1_CreON_RefvsAlt_LM_Annotated.tsv"),is.finite(fdr))%>%
  .[,plotcolor := as.factor(case_when(
    (logFC>log2(1.25) & fdr<0.1) ~ 1,
    (logFC<-log2(1.25) & fdr<0.1) ~ 2,
    TRUE ~ 3))]

stability_ON <- filter(fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreON_RefvsAlt_LM_Annotated.tsv"),is.finite(fdr_interaction)) %>%
  mutate(logFC_interaction = -(logFC_interaction)) %>%
  .[,plotcolor := as.factor(case_when(
    (logFC_interaction>log2(1.25) & fdr_interaction<0.1) ~ 1,
    (logFC_interaction<-log2(1.25) & fdr_interaction<0.1) ~ 2,
    TRUE ~ 3))]

p1_OFF <- ggplot(intron_OFF)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor,shape=Pheno),size=2)+
  scale_shape_manual(values=c(21,23))+
  scale_fill_manual(values=c(cbPalette[1],cbPalette[1],"lightgrey"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  geom_text_repel(data=filter(intron_OFF,Element %in% c("F2RL1_wgs","chr5:76327253;G|C;Family=11350"),fdr<0.1),
                  aes(x=logFC,y=-log10(fdr),label=Symbol))+
  labs(fill="Expression Change",
       title="Variant vs. Reference Expression Change\nin Pre-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")
print(p1_OFF)

p2_OFF <- ggplot(exon_OFF)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor,shape=Pheno),size=2)+
  scale_shape_manual(values=c(21,23))+
  scale_fill_manual(values=c(cbPalette[1],cbPalette[1],"lightgrey"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  geom_text_repel(data=filter(exon_OFF,Element %in% c("chr5:76327253;G|C;Family=11350","ZNF169_wgs"),fdr<0.1),
                  aes(x=logFC,y=-log10(fdr),label=Symbol))+
  labs(fill="Expression Change",
       title="Variant vs. Reference Expression Change\nin Post-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")
print(p2_OFF)

p3_OFF <- ggplot(stability_OFF)+
  geom_point(aes(x=logFC_interaction,y=-log10(fdr_interaction),fill=plotcolor,shape=Pheno),size=2)+
  scale_shape_manual(values=c(21,23))+
  scale_fill_manual(values=c(cbPalette[1],cbPalette[1],"lightgrey"),
                    labels=c("Increased Stability","Decreased Stability",
                             "Not Significant"))+
  geom_text_repel(data=filter(stability_OFF,Element %in% c("chr7:151987279;C|T;Family=13556","chr16:70687915;C|T;Family=11168"),fdr_interaction<0.1),
                  aes(x=logFC_interaction,y=-log10(fdr_interaction),label=Symbol))+
  labs(fill="Stability Change",
       title="Variant vs. Reference Stability Change",
       x="Log2 Fold Change in Stability",
       y="-log10(FDR)")+
  geom_vline(xintercept = -log2(1.25),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")
print(p3_OFF)

export_graph_start("./Figures/Figure_6/RefvsAlt_All_CreOFF_fig6g-i.pdf",sideways=TRUE,height=0.5)
print(p1_OFF+theme(legend.position = "none")+labs(title=NULL,subtitle="Pre-Splicing",x="Log2FC Expression")+
        p2_OFF+theme(legend.position = "none")+labs(title=NULL,subtitle="Post-Splicing",x="Log2FC Expression")+
        p3_OFF+theme(legend.position = "none")+labs(title=NULL,subtitle="Stability",x="Log2FC Stability")+
        plot_layout(ncol=4)+plot_annotation(title="Reference vs. Variant"))
export_graph_end()
 ```

###Figure 6J
```{r figure_6j}
exon_stability <- inner_join(
  fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Post-Splicing_Vglut1_CreOFF_RefvsAlt_LM.tsv"),
  fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreOFF_RefvsAlt_LM.tsv"),
  "Element")


#Significant in both and same direction
true_positives <- filter(exon_stability,
                         fdr<0.1,
                         fdr_interaction<0.1,
                         ((logFC>0 & logFC_interaction>0) | (logFC & logFC_interaction<0))
                         )$Element
#Only significant in stability, not post-splicing
false_negatives <- filter(exon_stability,
                         fdr>0.1,
                         fdr_interaction<0.1)$Element
#Either only significant in post-splicing, or significant in both but opposite
#directions
false_positives <- c(
  filter(exon_stability,
         fdr<0.1,
         fdr_interaction>0.1)$Element,
  filter(exon_stability,
         fdr<0.1,
         fdr_interaction<0.1,
         ((logFC>0 & logFC_interaction<0) | (logFC<0 & logFC_interaction>0)))$Element)
  




all_else <- filter(fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreOFF_RefvsAlt_LM.tsv"),
                   !Element %in% false_positives,
                   !Element %in% false_negatives,
                   !Element %in% true_positives)$Element

plottable <- rbind(
  data.table("Element"=true_positives,"Category"="True Positive"),
  data.table("Element"=false_positives,"Category"="False Positive"),
  data.table("Element"=false_negatives,"Category"="False Negative"),
  data.table("Element"=all_else,"Category"="Not Significant")
)

plottable[,Category := factor(Category,levels=c("True Positive","False Positive","False Negative","Not Significant"))]

summary_data <- plottable %>%
  group_by(Category) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

# Separate the data for labels
label_data <- summary_data %>%
  filter(Category != "Not Significant") %>%
  mutate(cumulative_percentage = 100 - cumsum(percentage) + percentage/2)

# Calculate the break point (adjust as needed)
break_point <- max(label_data$percentage) * 2  # Double the largest small percentage

# Create the plot

p5 <- ggplot(summary_data, aes(x = "", y = percentage, fill = Category)) +
  geom_col(width = 0.4, color = "black") +
  geom_text(data = subset(summary_data, Category == "Not Significant"),
            aes(label = paste0(round(percentage, 1), "%")),
            position = position_stack(vjust = 0.95),
            size = 5, fontface = "bold", color = "white") +
  geom_label_repel(data = label_data,
                   aes(label = paste0(Category, "\n", round(percentage, 3), "%"),
                       y = cumulative_percentage),
                   size = 3, direction = "y", hjust = 0, nudge_x = 0.6,
                   force = 1, show.legend = FALSE,
                   min.segment.length = 0) +
  scale_fill_manual(values = c("Not Significant" = "lightgrey",
                               "True Positive" = "#136d52",
                               "False Positive" = "#21c494",
                               "False Negative" = "#67e5bf")) +
  labs(title = "Standard MPRA Accuracy",
       y = "Percentage",
       fill = "Category") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        legend.position = "right",
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

export_graph_start("./Figures/Figure_6/RefAlt_Accuracy_CreOFF_fig6J.pdf",width=0.75)
(p5+scale_y_break(c(break_point,85)))+theme_minimal()
ggplot(summary_data,aes(Category,count,fill=as.factor(count)))+
  geom_col()+
  geom_text(aes(y=count-100,label=paste0(round(percentage,2),"\n",count)))+
  theme(axis.text.x = element_text(angle = 45,hjust=1))
export_graph_end()

exon_stability <- left_join(exon_stability,plottable)
exon_stability <- unique.data.frame(exon_stability)
cortex_refalt_dotplot <- ggplot(exon_stability,aes(x=logFC,y=logFC_interaction,fill=Category))+
  geom_point(shape=21,alpha=0.5,size=3)+
  scale_fill_manual(values = c("Not Significant" = "lightgrey",
                               "True Positive" = "#136d52",
                               "False Positive" = "#21c494",
                               "False Negative" = "#67e5bf"))+
  labs(x="Log Fold Change\n(Post-Splice)",
       y="Log Fold Change\n(Stability)",
       )+
  theme_bw()+
  theme(legend.position = "none")
print(cortex_refalt_dotplot)
export_graph_start("./Figures/Figure_6/Cortex_RefAlt_PostvsStability_DotPlot_fig6J.pdf")
print(cortex_refalt_dotplot + 
        coord_fixed(ratio = 
    diff(range(exon_stability$logFC[is.finite(exon_stability$logFC)])) / 
    diff(range(exon_stability$logFC_interaction[is.finite(exon_stability$logFC_interaction)]))))
export_graph_end()
```

###Figure 6K-M
```{r figure6_k-m}

p1_shuf_ON <- ggplot(filter(shufdata,Splicing_Pool=="Pre",Cre_Status=="Excitatory Neurons"))+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor),shape=21,size=2)+
  scale_fill_manual(values=c(cbPalette[1],cbPalette[1],"lightgrey"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  labs(fill="Expression Change",
       title="Variant vs. Reference Expression Change\nin Pre-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = log2(0.75),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")
p1_shuf_ON

p2_shuf_ON <- ggplot(filter(shufdata,Splicing_Pool=="Post",Cre_Status=="Excitatory Neurons"))+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor),shape=21,size=2)+
  scale_fill_manual(values=c(cbPalette[1],cbPalette[1],"lightgrey"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  labs(fill="Expression Change",
       title="Variant vs. Reference Expression Change\nin Pre-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = log2(0.75),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")
p2_shuf_ON

p3_shuf_ON <- ggplot(filter(shufdata,Splicing_Pool=="Stability",Cre_Status=="Excitatory Neurons"))+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor),shape=21,size=2)+
  scale_fill_manual(values=c(cbPalette[1],cbPalette[1],"lightgrey"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  labs(fill="Expression Change",
       title="Variant vs. Reference Expression Change\nin Pre-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = log2(0.75),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")
p3_shuf_ON

export_graph_start("./Figures/Figure_6/RefvsShuf_All_CreON_fig6k-m.pdf",sideways=TRUE,height=0.5)
print(p1_shuf_ON+theme(legend.position = "none")+labs(title=NULL,subtitle="Pre-Splicing",x="Log2FC Expression")+
        p2_shuf_ON+theme(legend.position = "none")+labs(title=NULL,subtitle="Post-Splicing",x="Log2FC Expression")+
        p3_shuf_ON+theme(legend.position = "none")+labs(title=NULL,subtitle="Stability",x="Log2FC Stability")+
        plot_layout(ncol=3)+plot_annotation(title="Reference vs. Shuffled"))
export_graph_end()
```

###Figure 6N
```{r figure_6N}
exon_stability <- inner_join(
  filter(shufdata,Splicing_Pool=="Post",Cre_Status=="Excitatory Neurons"),
  filter(shufdata,Splicing_Pool=="Stability",Cre_Status=="Excitatory Neurons"),
  "Element")


#Significant in both and same direction
true_positives <- filter(exon_stability,
                         fdr.x<0.1,
                         fdr.y<0.1,
                         ((logFC.x>0 & logFC.y>0) | (logFC.x<0 & logFC.y<0))
                         )$Element
#Only significant in stability, not post-splicing
false_negatives <- filter(exon_stability,
                         fdr.x>0.1,
                         fdr.y<0.1)$Element
#Either only significant in post-splicing, or significant in both but opposite
#directions
false_positives <- c(
  filter(exon_stability,
         fdr.x<0.1,
         fdr.y>0.1)$Element,
  filter(exon_stability,
         fdr.x<0.1,
         fdr.y<0.1,
         ((logFC.x>0 & logFC.y<0) | (logFC.x<0 & logFC.y>0)))$Element)
  




all_else <- filter(exon_stability,
                   !Element %in% false_positives,
                   !Element %in% false_negatives,
                   !Element %in% true_positives)$Element

plottable <- rbind(
  data.table("Element"=true_positives,"Category"="True Positive"),
  data.table("Element"=false_positives,"Category"="False Positive"),
  data.table("Element"=false_negatives,"Category"="False Negative"),
  data.table("Element"=all_else,"Category"="Not Significant")
)

plottable[,Category := factor(Category,levels=c("True Positive","False Positive","False Negative","Not Significant"))]

summary_data <- plottable %>%
  group_by(Category) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

# Separate the data for labels
label_data <- summary_data %>%
  filter(Category != "Not Significant") %>%
  mutate(cumulative_percentage = 100 - cumsum(percentage) + percentage/2)

# Create the plot

p5_shuf <- ggplot(summary_data, aes(x = "", y = percentage, fill = Category)) +
  geom_col(width = 0.4, color = "black") +
  scale_fill_manual(values = c("Not Significant" = "lightgrey",
                               "True Positive" = "#136d52",
                               "False Positive" = "#21c494",
                               "False Negative" = "#67e5bf")) +
  labs(y = "Percentage",
       fill = "Category",
       x=NULL) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        legend.position = "right",
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

export_graph_start("./Figures/Figure_6/RefShuf_Accuracy_CreON_fig6N.pdf",width=0.75)
print(p5_shuf+theme_minimal())
ggplot(summary_data,aes(Category,count,fill=as.factor(count)))+
  geom_col()+
  geom_text(aes(y=count-100,label=paste0(round(percentage,2),"\n",count)))+
  theme(axis.text.x = element_text(angle = 45,hjust=1))
export_graph_end()

exon_stability <- left_join(exon_stability,plottable)
exon_stability <- unique.data.frame(exon_stability)
neuron_shuf_dotplot <- ggplot(exon_stability,aes(x=logFC.x,y=logFC.y,fill=Category))+
  geom_point(shape=21,alpha=0.5,size=3)+
  scale_fill_manual(values = c("Not Significant" = "lightgrey",
                               "True Positive" = "#136d52",
                               "False Positive" = "#21c494",
                               "False Negative" = "#67e5bf"))+
  labs(x="Log Fold Change\n(Post-Splice)",
       y="Log Fold Change\n(Stability)",
       )+
  theme_bw()+
  theme(legend.position = "none")
print(neuron_shuf_dotplot)
export_graph_start("./Figures/Figure_6/Neuron_RefShuf_PostvsStability_DotPlot_fig6N.pdf")
print(neuron_shuf_dotplot + 
        coord_fixed(ratio = 
    diff(range(exon_stability$logFC.x[is.finite(exon_stability$logFC.x)])) / 
    diff(range(exon_stability$logFC.y[is.finite(exon_stability$logFC.y)]))))
export_graph_end()
```

###Figure 6O-Q

```{r figure_6o-q}
p1_ON <- ggplot(intron_ON)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor,shape=Pheno),size=2)+
  scale_shape_manual(values=c(21,23))+
  scale_fill_manual(values=c(cbPalette[1],cbPalette[1],"lightgrey"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  geom_text_repel(data=filter(intron_ON,Element %in% c("chr5:76327253;G|C;Family=11350","MAST3_wgs"),fdr<0.1),
                  aes(x=logFC,y=-log10(fdr),label=Symbol))+
  labs(fill="Expression Change",
       title="Variant vs. Reference Expression Change\nin Pre-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = log2(0.75),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")
print(p1_ON)

p2_ON <- ggplot(exon_ON)+
  geom_point(aes(x=logFC,y=-log10(fdr),fill=plotcolor,shape=Pheno),size=2)+
  scale_shape_manual(values=c(21,23))+
  scale_fill_manual(values=c(cbPalette[1],cbPalette[1],"lightgrey"),
                    labels=c("Increased Expression","Decreased Expression",
                             "Not Significant"))+
  geom_text_repel(data=filter(exon_ON,Element %in% c("chr5:76327253;G|C;Family=11350","ZNF169_wgs"),fdr<0.1),
                  aes(x=logFC,y=-log10(fdr),label=Symbol))+
  labs(fill="Expression Change",
       title="Variant vs. Reference Expression Change\nin Post-Splicing Pool",
       x="Log2 Fold Change in Expression",
       y="-log10(FDR)")+
  geom_vline(xintercept = log2(0.75),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")
print(p2_ON)

p3_ON <- ggplot(stability_ON)+
  geom_point(aes(x=logFC_interaction,y=-log10(fdr_interaction),fill=plotcolor,shape=Pheno),size=2)+
  scale_shape_manual(values=c(21,23))+
  scale_fill_manual(values=c(cbPalette[1],cbPalette[1],"lightgrey"),
                    labels=c("Increased Stability","Decreased Stability",
                             "Not Significant"))+
  geom_text_repel(data=filter(stability_ON,Element %in% c("chr10:50306955;T|C;Family=13928","chr4:95156901;A|G;Family=11368"),fdr_interaction<0.1),
                  aes(x=logFC_interaction,y=-log10(fdr_interaction),label=Symbol))+
  labs(fill="Stability Change",
       title="Variant vs. Reference Stability Change",
       x="Log2 Fold Change in Stability",
       y="-log10(FDR)")+
  geom_vline(xintercept = log2(0.75),linetype="dashed")+
  geom_vline(xintercept = log2(1.25),linetype="dashed")+
  geom_hline(yintercept = -log10(0.1),linetype="dashed")
print(p3_ON)

export_graph_start("./Figures/Figure_6/RefvsAlt_All_CreON_fig_6o-q.pdf",sideways=TRUE,height=0.5)
print(p1_ON+theme(legend.position = "none")+labs(title=NULL,subtitle="Pre-Splicing",x="Log2FC Expression")+
        p2_ON+theme(legend.position = "none")+labs(title=NULL,subtitle="Post-Splicing",x="Log2FC Expression")+
        p3_ON+theme(legend.position = "none")+labs(title=NULL,subtitle="Stability",x="Log2FC Stability")+
        plot_layout(ncol=4)+plot_annotation(title="Reference vs. Variant"))
export_graph_end()
```

###Figure 6R
```{r figure_6R}

exon_stability <- inner_join(
  fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Post-Splicing_Vglut1_CreON_RefvsAlt_LM.tsv"),
  fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreON_RefvsAlt_LM.tsv"),
  "Element")


#Significant in both and same direction
true_positives <- filter(exon_stability,
                         fdr<0.1,
                         fdr_interaction<0.1,
                         ((logFC>0 & logFC_interaction>0) | (logFC & logFC_interaction<0))
                         )$Element
#Only significant in stability, not post-splicing
false_negatives <- filter(exon_stability,
                         fdr>0.1,
                         fdr_interaction<0.1)$Element
#Either only significant in post-splicing, or significant in both but opposite
#directions
false_positives <- c(
  filter(exon_stability,
         fdr<0.1,
         fdr_interaction>0.1)$Element,
  filter(exon_stability,
         fdr<0.1,
         fdr_interaction<0.1,
         ((logFC>0 & logFC_interaction<0) | (logFC<0 & logFC_interaction>0)))$Element)
  




all_else <- filter(fread("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreON_RefvsAlt_LM.tsv"),
                   !Element %in% false_positives,
                   !Element %in% false_negatives,
                   !Element %in% true_positives)$Element

plottable <- rbind(
  data.table("Element"=true_positives,"Category"="True Positive"),
  data.table("Element"=false_positives,"Category"="False Positive"),
  data.table("Element"=false_negatives,"Category"="False Negative"),
  data.table("Element"=all_else,"Category"="Not Significant")
)

plottable[,Category := factor(Category,levels=c("True Positive","False Positive","False Negative","Not Significant"))]

summary_data <- plottable %>%
  group_by(Category) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

# Separate the data for labels
label_data <- summary_data %>%
  filter(Category != "Not Significant") %>%
  mutate(cumulative_percentage = 100 - cumsum(percentage) + percentage/2)

# Calculate the break point (adjust as needed)
break_point <- max(label_data$percentage) * 2  # Double the largest small percentage

# Create the plot

p5 <- ggplot(summary_data, aes(x = "", y = percentage, fill = Category)) +
  geom_col(width = 0.4, color = "black") +
  geom_text(data = subset(summary_data, Category == "Not Significant"),
            aes(label = paste0(round(percentage, 1), "%")),
            position = position_stack(vjust = 0.95),
            size = 5, fontface = "bold", color = "white") +
  geom_label_repel(data = label_data,
                   aes(label = paste0(Category, "\n", round(percentage, 3), "%"),
                       y = cumulative_percentage),
                   size = 3, direction = "y", hjust = 0, nudge_x = 0.6,
                   force = 1, show.legend = FALSE,
                   min.segment.length = 0) +
  scale_fill_manual(values = c("Not Significant" = "lightgrey",
                               "True Positive" = "#136d52",
                               "False Positive" = "#21c494",
                               "False Negative" = "#67e5bf")) +
  labs(title = "Standard MPRA Accuracy",
       y = "Percentage",
       fill = "Category") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        legend.position = "right",
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

export_graph_start("./Figures/Figure_6/RefAlt_Accuracy_CreON_fig6r.pdf",width=0.75)
(p5+scale_y_break(c(break_point,95)))+theme_minimal()
ggplot(summary_data,aes(Category,count,fill=as.factor(count)))+
  geom_col()+
  geom_text(aes(y=count-100,label=paste0(round(percentage,2),"\n",count)))+
  theme(axis.text.x = element_text(angle = 45,hjust=1))
export_graph_end()

#New 17Jul2025, make a dotplot

exon_stability <- left_join(exon_stability,plottable)
exon_stability <- unique.data.frame(exon_stability)
neuron_refalt_dotplot <- ggplot(exon_stability,aes(x=logFC,y=logFC_interaction,fill=Category))+
  geom_point(shape=21,alpha=0.5,size=3)+
  scale_fill_manual(values = c("Not Significant" = "lightgrey",
                               "True Positive" = "#136d52",
                               "False Positive" = "#21c494",
                               "False Negative" = "#67e5bf"))+
  labs(x="Log Fold Change\n(Post-Splice)",
       y="Log Fold Change\n(Stability)",
       )+
  theme_bw()+
  theme(legend.position = "none")
print(neuron_refalt_dotplot)
export_graph_start("./Figures/Figure_6//Neuron_RefAlt_PostvsStability_DotPlot_fig6R.pdf")
print(neuron_refalt_dotplot + 
        coord_fixed(ratio = 
    diff(range(exon_stability$logFC[is.finite(exon_stability$logFC)])) / 
    diff(range(exon_stability$logFC_interaction[is.finite(exon_stability$logFC_interaction)]))))
export_graph_end()


```

##Table S3 - 3'UTR Variants
```{r TableS3_3UTRvariants}
fwrite(fread("../code/mpra_tools/fullannotationtable.txt"),"./Tables/Table_S3.txt")
```

##Table S6 - In Cellulo Significance Results
```{r tableS6_singificance_results}

# 1. Define file paths and matching Excel sheet names (Strictly <= 31 chars)
file_info <- data.frame(
  filename = c("../analysis/Library1and2BCLess_ISOMPRA_6-30-2022/Pre-Splicing_RefvsShuf_Welchs.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_6-30-2022/Pre-Splicing_RefvsAlt_LM_Annotated.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_6-30-2022/Post-Splicing_RefvsShuf_Welchs.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_6-30-2022/Post-Splicing_RefvsAlt_LM_Annotated.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_6-30-2022/Stability_RefvsShuf_Welchs.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_6-30-2022/Stability_RefvsAlt_LM_Annotated.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_6-30-2022/Translation_Splicing-Agnostic_RefvsShuf_Welchs.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_6-30-2022/Translation_Splicing-Agnostic_RefvsAlt_LM_Annotated.tsv",
               "../analysis/ActD Trials/Full_Library_08Apr2026/Pre-Splicing_RefvsShuf_Welchs.tsv",
               "../analysis/ActD Trials/Full_Library_08Apr2026/PreSplicing_RefAlt_LM.txt",
               "../analysis/ActD Trials/Full_Library_08Apr2026/Post-Splicing_RefvsShuf_Welchs.tsv",
               "../analysis/ActD Trials/Full_Library_08Apr2026/PostSplicing_RefAlt_LM.txt",
               "../analysis/ActD Trials/Full_Library_08Apr2026/Stability_RefvsShuf_Welchs.tsv",
               "../analysis/ActD Trials/Full_Library_08Apr2026/Stability_RefAlt_LM.txt",
               "../analysis/ActD Trials/Full_Library_08Apr2026/ActD_Zero-to-Five_RefvsAlt_LM.tsv",
               "../analysis/ActD Trials/Full_Library_08Apr2026/ActD_Zero-to-Eleven_RefvsAlt_LM.tsv"),
  sheetname = c("Pre-Splicing RefvsShuf TRAP",
                "Pre-Splicing RefvsAlt TRAP",
                "Post-Splicing RefvsShuf TRAP",
                "Post-Splicing RefvsAlt TRAP",
                "Stability RefvsShuf TRAP",
                "Stability RefvsAlt TRAP",
                "Translation RefvsShuf TRAP",
                "Translation RevvsAlt TRAP",
                "Pre-Splicing RefvsShuf ActD",
                "Pre-Splicing RefvsAlt ActD",
                "Post-Splicing RefvsShuf ActD",
                "Post-Splicing RefvsAlt ActD",
                "Stability RefvsShuf ActD",
                "Stability RefvsAlt ActD",
                "ActD t0 vs. t5 Allelic Ratio",   # Shortened to 31 chars
                "ActD t0 vs t11 Allelic Ratio"),  # Shortened to 31 chars
  stringsAsFactors = FALSE
)

# 2. Define data for the new summary landing page metadata matrix
summary_data <- data.frame(
  SheetName = file_info$sheetname,
  Description = c(
    "HEK233 Cell Element Effects in Pre-Splicing Pool from TRAP Experiment",
    "HEK233 Cell Variant Effects in Pre-Splicing Pool from TRAP Experiment",
    "HEK233 Cell Element Effects in Post-Splicing Pool from TRAP Experiment",
    "HEK233 Cell Variant Effects in Post-Splicing Pool from TRAP Experiment",
    "HEK233 Cell Element Effects on Stability from TRAP Experiment",
    "HEK233 Cell Variant Effects on Stability from TRAP Experiment",
    "HEK233 Cell Element Effects on Translational Efficiency from TRAP Experiment",
    "HEK233 Cell Variant Effects on Translational Efficiency from TRAP Experiment",
    "HEK233 Cell Element Effects in Pre-Splicing Pool from Actinomycin D Assay",
    "HEK233 Cell Variant Effects in Pre-Splicing Pool from Actinomycin D Assay",
    "HEK233 Cell Element Effects in Post-Splicing Pool from Actinomycin D Assay",
    "HEK233 Cell Variant Effects in Post-Splicing Pool from Actinomycin D Assay",
    "HEK233 Cell Element Effects on Stability from Actinomycin D Assay Time Zero",
    "HEK233 Cell Variant Effects on Stability from Actinomycin D Assay Time Zero",
    "HEK233 Cell Variant Effects on Allelic Ratio (Variant/WT) from Actinomycin D Assay Time Zero vs. Time Five Hours",
    "HEK233 Cell Variant Effects on Allelic Ratio (Variant/WT) from Actinomycin D Assay Time Zero vs. Time Eleven Hours"
  ),
  stringsAsFactors = FALSE
)

# Create a new Excel workbook
workbook <- createWorkbook()

# 3. Create and Write the Summary landing sheet at position 1
addWorksheet(workbook, "Summary")

# Write the Table header layout metadata block
writeData(workbook, "Summary", x = "Table S6. HEK293 In Cellulo uISOMPRA Results.", startCol = 1, startRow = 1)
addStyle(workbook, "Summary", style = createStyle(textDecoration = "bold"), rows = 1, cols = 1)

# Write the index dataframe underneath it (row 3, without column names to match image format layout)
writeData(workbook, "Summary", x = summary_data, startCol = 1, startRow = 3, colNames = FALSE)

# 4. Loop through and populate each individual experimental data sheet
for (i in 1:nrow(file_info)) {
  # Read individual source data sets
  data <- fread(file_info$filename[i])
  
  # Append data worksheets sequentially after the summary page
  addWorksheet(workbook, file_info$sheetname[i])
  writeData(workbook, file_info$sheetname[i], data)
}

# 5. Automatically adjust column widths across all sheets for a clean fit
for (sheet in names(workbook)) {
  setColWidths(workbook, sheet, cols = 1:ncol(summary_data), widths = "auto")
}

# Save the workbook file artifact downstream
saveWorkbook(workbook, "./Tables/Table_S6.xlsx", overwrite = TRUE)
```

##Table S8 - In Vivo Significance Results
```{r tableS8_invivo_signficance_results}

# 1. Define file paths and matching Excel sheet names (Strictly <= 31 chars)
file_info <- data.frame(
  filename = c("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Pre-Splicing_Vglut1_CreOFF_RefvsShuf_Welchs_Annotated.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Pre-Splicing_Vglut1_CreOFF_RefvsAlt_LM_Annotated.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Pre-Splicing_Vglut1_CreON_RefvsShuf_Welchs_Annotated.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Pre-Splicing_Vglut1_CreON_RefvsAlt_LM_Annotated.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Post-Splicing_Vglut1_CreOFF_RefvsShuf_Welchs_Annotated.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Post-Splicing_Vglut1_CreOFF_RefvsAlt_LM_Annotated.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Post-Splicing_Vglut1_CreON_RefvsShuf_Welchs_Annotated.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Post-Splicing_Vglut1_CreON_RefvsAlt_LM_Annotated.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreOFF_RefvsShuf_Welchs_Annotated.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreOFF_RefvsAlt_LM_Annotated.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreON_RefvsShuf_Welchs_Annotated.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreON_RefvsAlt_LM_Annotated.tsv"),
  sheetname = c("Pre-Splicing Cortex RefvsShuf",
                "Pre-Splicing Cortex RefvsAlt",
                "Pre-Splicing ExN RefvsShuf",
                "Pre-Splicing ExN RefvsAlt",
                "Post-Splicing Cortex RefvsShuf", # Shortened to 30 chars
                "Post-Splicing Cortex RefvsAlt",
                "Post-Splicing ExN RefvsShuf",
                "Post-Splicing ExN RefvsAlt",
                "Stability Cortex RefvsShuf",
                "Stability Cortex RefvsAlt",
                "Stability ExN RefvsShuf",
                "Stability ExN RefvsAlt"),
  stringsAsFactors = FALSE
)

# 2. Define data for the Summary landing page text
summary_data <- data.frame(
  SheetName = file_info$sheetname,
  Description = c(
    "Whole Cortex Element Effects in Pre-Splicing Pool",
    "Whole Cortex Variant Effects in Pre-Splicing Pool",
    "Excitatory Neuron Element Effects in Pre-Splicing Pool",
    "Excitatory Neuron Variant Effects in Pre-Splicing Pool",
    "Whole Cortex Element Effects in Post-Splicing Pool",
    "Whole Cortex Variant Effects in Post-Splicing Pool",
    "Excitatory Neuron Element Effects in Post-Splicing Pool",
    "Excitatory Neuron Variant Effects in Post-Splicing Pool",
    "Whole Cortex Element Effects on Stability",
    "Whole Cortex Variant Effects on Stability",
    "Excitatory Neuron Element Effects on Stability",
    "Excitatory Neuron Variant Effects on Stability"
  ),
  stringsAsFactors = FALSE
)

# Create a new Excel workbook
workbook <- createWorkbook()

# 3. Create and Write the Summary landing sheet at position 1
addWorksheet(workbook, "Summary")

# Write the title block at row 1, column 1
writeData(workbook, "Summary", x = "Table S8. uISOMPRA In Vivo Results", startCol = 1, startRow = 1)
addStyle(workbook, "Summary", style = createStyle(textDecoration = "bold"), rows = 1, cols = 1)

# Write the summary index block below it (starting at row 3) without headers
writeData(workbook, "Summary", x = summary_data, startCol = 1, startRow = 3, colNames = FALSE)

# Set clean sizing for the Summary tab columns so descriptions look perfect
setColWidths(workbook, "Summary", cols = 1, widths = 35)
setColWidths(workbook, "Summary", cols = 2, widths = 65)

# 4. Loop through and populate each individual experimental data sheet
for (i in 1:nrow(file_info)) {
  data <- fread(file_info$filename[i])
  
  addWorksheet(workbook, file_info$sheetname[i])
  writeData(workbook, file_info$sheetname[i], data)
  
  # Automatically adjust data sheet columns so formatting is clean
  setColWidths(workbook, file_info$sheetname[i], cols = 1:ncol(data), widths = "auto")
}

# 5. Save the final openxlsx object to the required folder path
saveWorkbook(workbook, "./Tables/Table_S8.xlsx", overwrite = TRUE)
```

##Table S9 - In Vivo Stability:Sex Effects
```{r tableS9_invivo_stability_sex_effects}
#And separately an xlsx for the Allele:SpliceStatus:Sex, just two workbooks
file_info <- data.frame(
  filename = c("../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreOFF_AllelevsSplicevsSex_LM_Annotated.tsv",
               "../analysis/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreON_AllelevsSplicevsSex_LM_Annotated.tsv"),
  sheetname = c("Cortex AllelevsSplicevsSex",
                "ExN AllelevsSplicevsSex"), # Ensure names are ≤ 31 characters
  stringsAsFactors = FALSE
)

# 2. Define data for the Summary landing page text
summary_data <- data.frame(
  SheetName = file_info$sheetname,
  Description = c(
    "uISOMPRA Allele*SpliceStatus*Sex Significance Testing in Whole Cortex
",
    "uISOMPRA Allele*SpliceStatus*Sex Significance Testing in Excitatory Neurons
"),
  stringsAsFactors = FALSE
)

# Create a new Excel workbook
workbook <- createWorkbook()

# 3. Create and Write the Summary landing sheet at position 1
addWorksheet(workbook, "Summary")

# Write the title block at row 1, column 1
writeData(workbook, "Summary", x = "Table S9. uISOMPRA In Vivo Sex Effects
", startCol = 1, startRow = 1)
addStyle(workbook, "Summary", style = createStyle(textDecoration = "bold"), rows = 1, cols = 1)

# Write the summary index block below it (starting at row 3) without headers
writeData(workbook, "Summary", x = summary_data, startCol = 1, startRow = 3, colNames = FALSE)

# Set clean sizing for the Summary tab columns so descriptions look perfect
setColWidths(workbook, "Summary", cols = 1, widths = 35)
setColWidths(workbook, "Summary", cols = 2, widths = 65)

# 4. Loop through and populate each individual experimental data sheet
for (i in 1:nrow(file_info)) {
  data <- fread(file_info$filename[i])
  
  addWorksheet(workbook, file_info$sheetname[i])
  writeData(workbook, file_info$sheetname[i], data)
  
  # Automatically adjust data sheet columns so formatting is clean
  setColWidths(workbook, file_info$sheetname[i], cols = 1:ncol(data), widths = "auto")
}

# 5. Save the final openxlsx object to the required folder path
saveWorkbook(workbook, "./Tables/Table_S9.xlsx", overwrite = TRUE)
```


```{r}
#From latest talk before end of 2024, make a chart with the Allele:Sex sigs from
#the above two data tables, boxplots with lines, following the code from earlier
#to make the boxplot data table, only cre-off yielded something

d_off <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreON_AllelevsSplicevsSex_LM_Annotated.tsv")
d_on <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Stability_Vglut1_CreOFF_AllelevsSplicevsSex_LM_Annotated.tsv")
d <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/Vglut1-Cre Expression Table.csv")
#give the data a symbol (gene name) for plotting
d <- left_join(d,select(fread("../../code/mpra_tools/fullannotationtable.txt"),c("Element","Symbol")))

make_sex_lineplot <- function(data){
  elements <- unique(data$Element)
  plots <- lapply(elements,function(element) {
   sub_data <- filter(data,Element==element,(Allele=="ref" | Allele=="alt"))
   sub_data$Allele <- factor(sub_data$Allele,levels = c("ref","alt"),labels = c("Reference","Variant"))
   sub_data$Splicing_Pool <- factor(sub_data$Splicing_Pool,levels = c("Pre","Post"))
   sub_data$Sex <- factor(sub_data$Sex,levels = c("m","f"),labels = c("Male","Female"))
   genetitle <- sub_data$Symbol
    ggplot(filter(sub_data,Element==element),aes(x=Sex,y=Expression,fill=Splicing_Pool,group=interaction(Splicing_Pool,Sex)))+
    geom_boxplot(color="black",outliers = TRUE,position = position_dodge(width = 0),linewidth=0.1)+
    geom_point(shape=21,color="black",position = position_dodge(width = 0))+
      stat_smooth(method="lm",aes(group=Splicing_Pool,color=Splicing_Pool),show.legend = FALSE,linewidth=0.25)+
      labs(title=unique(sub_data$Symbol))+
      scale_fill_manual(values=c("#136d52","#67e5bf"))+
      scale_color_manual(values=c("#136d52","#67e5bf"))
  })
  return(plots)
}

export_graph_start("VGlut1_CreOFF_SplicingPoolvsSex.pdf")  
make_sex_lineplot(filter(d,Cre_Status=="OFF",
                         Element %in% filter(
                           d_off,`fdr_Splicing_Pool:Sex`<0.1)$Element
                         ))
export_graph_end()
```

```{r}
#Data from the TRAP in vivo, looking at Allele*Sex interaction
#No Ref vs alt sigs from whole cortex, but some from ExN
refalt_on <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_6-1-2023/Translation_RefvsAlt_CreON_LM_Annotated.tsv")
allelesex_off <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_6-1-2023/Translation_AllelevsSex_CreOFF_LM_Annotated.tsv")
allelesex_on <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_6-1-2023/Translation_AllelevsSex_CreON_LM_Annotated.tsv")
d_te_off <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_6-1-2023/Cre-OFF_Translation_Values.tsv")
d_te_on <- fread("../Library1and2BCLess_ISOMPRA_In_Vivo_6-1-2023/Cre-ON_Translation_Values.tsv")

d_te_off <- left_join(d_te_off,select(fread("../../code/mpra_tools/fullannotationtable.txt"),c("Element","Symbol")))

make_te_sex_lineplot <- function(data){
  elements <- unique(data$Element)
  plots <- lapply(elements,function(element) {
   sub_data <- filter(data,Element==element,(Allele=="ref" | Allele=="alt"))
   sub_data$Allele <- factor(sub_data$Allele,levels = c("ref","alt"),labels = c("Reference","Variant"))
   sub_data$Sex <- factor(sub_data$Sex,levels = c("m","f"),labels = c("Male","Female"))
   genetitle <- sub_data$Symbol
    ggplot(filter(sub_data,Element==element),aes(x=Sex,y=Expression,fill=Allele,group=interaction(Sex,Allele)))+
    geom_boxplot(color="black",outliers = TRUE,position = position_dodge(width = 0),linewidth=0.1)+
    geom_point(shape=21,color="black",position = position_dodge(width = 0))+
      stat_smooth(method="lm",aes(group=Allele,color=Allele),show.legend = FALSE,linewidth=0.25)+
      labs(title=unique(sub_data$Symbol),y="Translational Efficiency")+
      scale_fill_manual(values=c("#136d52","#67e5bf"))+
      scale_color_manual(values=c("#136d52","#67e5bf"))
  })
  return(plots)
}

export_graph_start("Whole_Cortex_SexMainEffect_Lineplots.pdf")
#Get a list of the top values from Cre-OFF
sigs <- filter(allelesex_off,fdr_secondterm<0.1)
setorderv(sigs,c("fdr_secondterm","logFC_secondterm"))
make_te_sex_lineplot(filter(d_te_off,Element %in% unique(sigs$Element)))
export_graph_end()

#Clean up
rm(list = setdiff(ls(), lsf.str()))
```


##Figure S2
```{r New Supplemental Figures 09Apr2025}
badAlt<-c('ZNF626_wgs','ASIC2_wgs', 'CDKN2B_wgs', 'ELMO1_wgs', 'ENPP5_wgs', 'GDPD4_ssc', 'INSR_wgs', 'LMBRD2_wgs', 'LRP2BP_wgs', 'SCARB2.1_wgs', 'TLK1_wgs', 'TRIM37_wgs', 'ZCCHC11_wgs')
GC_table <- fread("../data/formatted/ActD_Trials_08Apr2026/formatted.txt") %>% 
  filter(,!Element %in% badAlt,Fraction=="0hr",grepl("_ref$",Element)) %>%
  mutate(CPM = CPM + 1) %>% #Add a pseudocount
  pivot_wider(names_from = "Splicing",id_cols = c("Element","BiolRep"), values_from = "CPM") %>%
  mutate(Stability = log2(post)-log2(pre)) %>% group_by(Element) %>%
  summarize(Mean_Exon_Intron = mean(Stability)) %>% filter(is.finite(Mean_Exon_Intron))
#The reviewers asked for sequence features, so we're going to give them post-
#pre-splicing vs GC content, which is anticorrelated with free folding energy
#(makes sense that it would be)
free_folding <- readLines("../code/mpra_tools/freefolding.txt")
free_folding <- data.table("Element" = free_folding[seq(1,24291,3)],
                           "Sequence" = free_folding[seq(2,24291,3)],
                           "Energy" = free_folding[seq(3,24291,3)])
free_folding$Element <- str_replace_all(free_folding$Element,"^>","")
free_folding[, DeltaG := as.numeric(str_extract(Energy, "-?\\d+\\.\\d+"))]
free_folding <- free_folding[,c(1,2,4)]
free_folding <- free_folding %>%
  mutate(
    # Count G and C case-insensitively, then divide by total length
    GC_content = (str_count(Sequence, "(?i)[GC]") / str_length(Sequence))*100
  )
GC_table <- right_join(free_folding,GC_table)


export_graph_start("./Figures/Figure_S2/GC Content and Free Energy vs Stability Correlation_figS2c-d.pdf")
ggplot(GC_table,aes(x=GC_content,y=Mean_Exon_Intron))+
  geom_pointdensity()+
  stat_smooth(method = "lm",color="black",linewidth=1)+
  labs(
    x="Element GC Content %",
    y="Stability (Post- Minus Pre-Splicing Expression",
    title="GC Content vs. Stability Correlation")+
  geom_text(aes(x=30,
                y=2.5,
                label = paste0(
                  "Spearman's Rho: ",
                  round(
                    cor.test(GC_table$GC_content,
                             GC_table$Mean_Exon_Intron,
                             method="spearman")$estimate,2),
                  "****"
                  )))+
  scale_color_viridis_c()+
  theme(legend.position = "none")

ggplot(GC_table,aes(x=DeltaG,y=Mean_Exon_Intron))+
  geom_pointdensity()+
  stat_smooth(method = "lm",color="black",linewidth=1)+
  labs(
    x=expression("Minimum Free Energy (MFE) ("*Delta*"G, kcal/mol)"),
    y="Stability (Post- Minus Pre-Splicing Expression",
    title=expression("Minimum Free Energy (MFE, "*Delta*"G) vs. Stability Correlation"))+
  geom_text(aes(x=-65,
                y=2.5,
                label = paste0(
                  "Spearman's Rho: ",
                  round(
                    cor.test(GC_table$DeltaG,
                             GC_table$Mean_Exon_Intron,
                             method="spearman")$estimate,2),
                  "****"
                  )))+
  scale_color_viridis_c()+
  theme(legend.position = "none")
export_graph_end()



 
#New Raincloud plot for stability sequences
d <- fread("../../data/formatted/Library1and2BCLess_ISOMPRA_6-30-2022/formatted_Reads_compiled_06-30-2022.tsv") %>% filter(,!Element %in% badAlt)
d <- left_join(filter(d,Comment=="Input Intronic (pre) mRNA"),filter(d,Comment=="Input Exonic (mature) mRNA"),by=c("Element","Allele","BiolRep")) %>% mutate(Stability=log2(CPM.y)-log2(CPM.x))


p2 <- ggplot(
  rbind(
    filter(d,Allele=="dnvctl",Element!="QREx4",is.finite(Stability)),
    filter(d,Allele=="shuf",is.finite(Stability))%>%mutate(Element="Shuffled Elements"))
  %>%mutate(Element=factor(Element,levels=c("Shuffled Elements","PREx4","SREx4","WPRE-alpha"),labels=c("Shuffled Elements","Pumilio Response Element","Smaug Response Element","Woodchuck HV Posttranscriptional Response Element"))),
    aes(x=Element,y=Stability,fill=Element))+
  geom_rain(point.args = list(shape=21, alpha = 0.2),
            violin.args.pos = list(scale="width",width=1.5,position=position_nudge(x=0.15)),
            rain.side = "r")+
  geom_signif(comparisons = list(
    c("Pumilio Response Element","Shuffled Elements"),
    c("Smaug Response Element","Shuffled Elements"),
    c("Woodchuck HV Posttranscriptional Response Element","Shuffled Elements")),map_signif_level = TRUE,test = "t.test",y_position = c(1.5,1.8,2.1),tip_length = 0)+
  labs(title="Stability of Control Sequences",y="Stability\n(Post-Splicing Expression\nMinus Pre-Splicing Expression)")+
  theme(plot.title = element_text(hjust=0.5),
        legend.position = "none",
        axis.title.x = element_blank())+
  scale_y_continuous(limits = c(-2.5,2.7))+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 13))+
  scale_fill_manual(values=c("#454545","#808080","#bababa","#f5f5f5"))
print(p2)

p2 <- ggplot(
  rbind(
    filter(d,Allele=="dnvctl",Element!="QREx4",is.finite(Stability)),
    filter(d,Allele=="shuf",is.finite(Stability))%>%mutate(Element="Shuffled Elements"))
  %>%mutate(Element=factor(Element,levels=c("Shuffled Elements","PREx4","SREx4","WPRE-alpha"),labels=c("Shuffled","PRE","SRE","WPRE"))),
    aes(x=Element,y=Stability,fill=Element))+
  geom_rain(point.args = list(shape=21, alpha = 0.2))+
  geom_signif(comparisons = list(
    c("PRE","Shuffled"),
    c("SRE","Shuffled"),
    c("WPRE","Shuffled")),map_signif_level = TRUE,test = "t.test",y_position = c(1.5,1.8,2.1),tip_length = 0)+
  labs(title="Stability of Control Sequences",y="Stability\n(Post-Splicing Expression\nMinus Pre-Splicing Expression)")+
  theme(plot.title = element_text(hjust=0.5),legend.position = "none")+
  scale_y_continuous(limits = c(-2.5,2.7))+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 13))+
  scale_fill_manual(values=c("#454545","#808080","#bababa","#f5f5f5"))

print(p2)

export_graph_start("./Figures/Figure1/Control Stability - RainCloud.pdf")
print(p2+coord_fixed(ratio=0.45))
export_graph_end()


```

