---
title: "Variant Analysis of In Vivo ISOMPRA No TRAP"
author: "Anthony Fischer"
date: "2023-03-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
sessionInfo()
```

## Load Tools
```{r load_tools, include=FALSE}
source("../../code/mpra_tools/mpra_tools.R")
library(umap)
library(RColorBrewer)
library(ggforce)
library(fplot)
library(pdftools)
library(ggrepel)
cbPalette <- brewer.pal(8,"Dark2")
#A list of elements we do not consider from Library 1.0 as they have bad variant sequences
badAlt<-c('ZNF626_wgs','ASIC2_wgs', 'CDKN2B_wgs', 'ELMO1_wgs', 'ENPP5_wgs', 'GDPD4_ssc', 'INSR_wgs', 'LMBRD2_wgs', 'LRP2BP_wgs', 'SCARB2.1_wgs', 'TLK1_wgs', 'TRIM37_wgs', 'ZCCHC11_wgs')
#Set the theme for all ggplots
theme_set(
  theme_classic2(base_family = "sans")+
theme(
  plot.title=element_text(hjust=0.5, size = 14, face = "bold"),
  plot.subtitle = element_text(size = 12, hjust=0.5),
  axis.title = element_text(size = 12),
  axis.text = element_text(size = 10),
  legend.title = element_text(size = 14, face = "bold",hjust=0.5),
  legend.text = element_text(size = 10, face = "bold"),
  panel.border = element_rect(color = "black", fill = NA, linewidth = 1)))
```

## Define Welch's Reference vs. Shuffled T-Test Function
```{r welchs_refvsshuf_function}
welch_shuftest <- function(data_table,element,allele,null_values){
  d <- filter(data_table,Allele==allele,Element==element)
  if (nrow(d) < 3) {
    return(c(element, NA, NA))
  } else {
  res <- t.test(d$Expression,null_values, var.equal=FALSE)
  dif <- res$estimate[1] - res$estimate[2]
  pval <- res$p.value
  return(c(element,dif,pval))
  }
}
```

## Custom Function for Non-Parametric Sex Specific Element Effects
```{r sex-specific_elements}

library(data.table)
library(progress)  # for progress bar

analyze_expression_data <- function(dt, n_permutations = 10000) {
  # Convert to data.table once and subset to only needed columns
  if(!is.data.table(dt)) {
    dt <- as.data.table(dt)[, .(Element, Expression, Allele, Sex)]
  }
  
  # Pre-filter to remove NA/Inf/-Inf
  dt <- dt[is.finite(Expression)]
  setkey(dt, Element)
  
  # Function to analyze one element (optimized)
  analyze_one_element <- function(element_data) {
    # Get sample counts using fast data.table operations
    counts <- element_data[, .N, by = .(Sex, Allele)]
    setkey(counts, Sex, Allele)
    
    n_male_ref <- counts[.("m", "ref"), N]
    n_female_ref <- counts[.("f", "ref"), N]
    n_male_shuf <- counts[.("m", "shuf"), N]
    n_female_shuf <- counts[.("f", "shuf"), N]
    
    # Early return if not enough samples
    if(length(n_male_ref) == 0 || length(n_female_ref) == 0 || 
       n_male_ref < 3 || n_female_ref < 3) {
      return(c(
        n_male_ref = ifelse(length(n_male_ref) == 0, 0, n_male_ref),
        n_female_ref = ifelse(length(n_female_ref) == 0, 0, n_female_ref),
        n_male_shuf = ifelse(length(n_male_shuf) == 0, 0, n_male_shuf),
        n_female_shuf = ifelse(length(n_female_shuf) == 0, 0, n_female_shuf),
        male_effect = NA_real_,
        female_effect = NA_real_,
        male_p = NA_real_,
        female_p = NA_real_,
        interaction_effect = NA_real_,
        interaction_p = NA_real_
      ))
    }
    
    # Pre-calculate means
    means <- element_data[, .(mean_expr = mean(Expression)), by = .(Sex, Allele)]
    setkey(means, Sex, Allele)
    
    male_effect <- means[.("m", "ref"), mean_expr] - means[.("m", "shuf"), mean_expr]
    female_effect <- means[.("f", "ref"), mean_expr] - means[.("f", "shuf"), mean_expr]
    interaction_effect <- male_effect - female_effect
    
    perm_data <- element_data[, .(Expression, Sex, is_ref = (Allele == "ref"))]
    perm_diffs <- numeric(n_permutations)
    
    # Create mini progress bar for permutations (updates every 1000 iterations)
    for(i in 1:n_permutations) {
      perm_data[Sex == "m", is_ref_perm := sample(is_ref[Sex == "m"])]
      perm_data[Sex == "f", is_ref_perm := sample(is_ref[Sex == "f"])]
      
      perm_means <- perm_data[, .(mean_expr = mean(Expression)), by = .(Sex, is_ref_perm)]
      setkey(perm_means, Sex, is_ref_perm)
      
      male_perm <- perm_means[.("m", TRUE), mean_expr] - perm_means[.("m", FALSE), mean_expr]
      female_perm <- perm_means[.("f", TRUE), mean_expr] - perm_means[.("f", FALSE), mean_expr]
      perm_diffs[i] <- male_perm - female_perm
    }
    
    male_p <- mean(abs(perm_diffs) >= abs(male_effect))
    female_p <- mean(abs(perm_diffs) >= abs(female_effect))
    interaction_p <- mean(abs(perm_diffs) >= abs(interaction_effect))
    
    return(c(
      n_male_ref = n_male_ref,
      n_female_ref = n_female_ref,
      n_male_shuf = n_male_shuf,
      n_female_shuf = n_female_shuf,
      male_effect = male_effect,
      female_effect = female_effect,
      male_p = male_p,
      female_p = female_p,
      interaction_effect = interaction_effect,
      interaction_p = interaction_p
    ))
  }
  
  elements <- dt[, unique(Element)]
  n_elements <- length(elements)
  
  # Create progress bar for elements
  pb <- progress_bar$new(
    format = "Analyzing elements [:bar] :percent | Elapsed: :elapsed | ETA: :eta",
    total = n_elements,
    clear = FALSE,
    width = 80
  )
  
  # Process elements with progress bar
  results <- rbindlist(lapply(seq_along(elements), function(i) {
    elem <- elements[i]
    res <- analyze_one_element(dt[Element == elem])
    pb$tick()
    if(i %% 100 == 0) {
      cat(sprintf("\nProcessed %d/%d elements\n", i, n_elements))
    }
    as.list(c(Element = elem, res))
  }))
  
  # Add adjusted p-values
  cat("\nCalculating adjusted p-values...\n")
  p_cols <- c("male_p", "female_p", "interaction_p")
  for(col in p_cols) {
    results[, paste0(col, "_fdr") := p.adjust(get(col), method = "BH")]
    results[, paste0(col, "_bonf") := p.adjust(get(col), method = "bonferroni")]
  }
  
  setkey(results, Element)
  cat("Analysis complete!\n")
  return(results)
}

```

## Load Data
```{r}

d <- fread("../../data/formatted/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/formatted_Reads_08Mar2024.tsv")
d$Fraction <- "RNA"
dna <- fread("../../data/formatted/Library1and2BCLess_ISOMPRA_In_Vivo_NOTRAP_06Mar2024/formatted_viral_reads.tsv")
dna <- aggregate.counts(dna,"Fraction","Plasmid DNA","DNA.AGG",c("BiolRep","TechRep"),c(1,1),c("Fraction","BiolRep")) %>% filter(Fraction=="DNA.AGG")
dna <- tidyr::unite(dna,Element,c("Element","Library"))
dna$Element <- str_replace_all(dna$Element,"_$","")
dna <- dna[,1:9]
setnames(dna,
         c("BiolRep","CellType","SpliceStatus"),
         c("Animal_Hemisphere","Cre_Type","Splicing_Pool"))
d <- rbind(d,dna,fill=TRUE)

d_temp <- d
d_temp <- d_temp %>%
  mutate(Fraction = if_else(Fraction == "RNA",
                            paste(Cre_Type, Cre_Status, Splicing_Pool, sep = "_"),
                            Fraction))

d_temp <- d_temp[,c(1,2,4,7,8,9)]
d_temp$BC <- 1
#Let's filter for elements that are well represented
#We need to filter separately for each of the types of libraries, but the Vgat
#samples sequenced horribly, unacceptable correlations, so we'll only be looking
#at excitatory neurons (Vglut1) and whole cortex

Vglut1.OFF.Pre.to.test <- filter_elements(d_temp,
                                          "Fraction",
                                          "VGlut1_OFF_Pre",
                                          "DNA.AGG",
                                          min.bc = 1,
                                          replicate.field = "Animal_Hemisphere")
Vglut1.OFF.Post.to.test <- filter_elements(d_temp,
                                          "Fraction",
                                          "VGlut1_OFF_Post",
                                          "DNA.AGG",
                                          min.bc = 1,
                                          replicate.field = "Animal_Hemisphere")
Vglut1.ON.Pre.to.test <- filter_elements(d_temp,
                                          "Fraction",
                                          "VGlut1_ON_Pre",
                                          "DNA.AGG",
                                          min.bc = 1,
                                          replicate.field = "Animal_Hemisphere")
Vglut1.ON.Post.to.test <- filter_elements(d_temp,
                                          "Fraction",
                                          "VGlut1_ON_Post",
                                          "DNA.AGG",
                                          min.bc = 1,
                                          replicate.field = "Animal_Hemisphere")

#It took forever to generate these lists of testable elements that passed QC,
#save them so I don't have to regenerate
fwrite(Vglut1.OFF.Post.to.test,"Vglut1_Cre-OFF_Post_TestingElements.txt",sep="\t")
fwrite(Vglut1.ON.Post.to.test,"Vglut1_Cre-ON_Post_TestingElements.txt",sep="\t")
fwrite(Vglut1.OFF.Pre.to.test,"Vglut1_Cre-OFF_Pre_TestingElements.txt",sep="\t")
fwrite(Vglut1.ON.Pre.to.test,"Vglut1_Cre-ON_Pre_TestingElements.txt",sep="\t")

#We'll use edgeR to normalize the counts, keep it separate
#for each type of library and combine at the end

d_norm <- rbind(
  normalize.counts.edgeR(
    filter(d,Cre_Type=="VGlut1",Cre_Status=="OFF",Splicing_Pool=="Pre"),
    sample.fields = c("Cre_Type","Animal_Hemisphere","Cre_Status","Splicing_Pool"),
    element.fields = c("Element","Allele")),
  normalize.counts.edgeR(
    filter(d,Cre_Type=="VGlut1",Cre_Status=="OFF",Splicing_Pool=="Post"),
    sample.fields = c("Cre_Type","Animal_Hemisphere","Cre_Status","Splicing_Pool"),
    element.fields = c("Element","Allele")),
  normalize.counts.edgeR(
    filter(d,Cre_Type=="VGlut1",Cre_Status=="ON",Splicing_Pool=="Pre"),
    sample.fields = c("Cre_Type","Animal_Hemisphere","Cre_Status","Splicing_Pool"),
    element.fields = c("Element","Allele")),
  normalize.counts.edgeR(
    filter(d,Cre_Type=="VGlut1",Cre_Status=="ON",Splicing_Pool=="Post"),
    sample.fields = c("Cre_Type","Animal_Hemisphere","Cre_Status","Splicing_Pool"),
    element.fields = c("Element","Allele"))
)


d_expr <- left_join(d_norm,dna[,c(1,2,4)],c("Element","Allele")) %>%
  mutate(Expression=log2(CPM.x/CPM.y))
d_expr <- as.data.table(d_expr)

fwrite(d_expr[,c(1,2,3,4,5,6,9)],"Vglut1-Cre Expression Table.csv")

#Note here, this wasn't available in the original data table, but any biolrep (L or R)
#1-5 was male, 6-10 was female.  If running Vgat (not advised), 1-4 was male, 5-8 was female
d_expr[, Sex := {
    # Extract only numeric characters and convert to number
    num <- as.numeric(gsub("[^0-9]", "", Animal_Hemisphere))
    # Assign sex based on number range
    ifelse(num >= 1 & num <= 5, "m",
           ifelse(num >= 6 & num <= 10, "f", NA_character_))
  }]
d_expr <- as.data.table(d_expr)

fwrite(d_expr,"Vglut1-Cre Expression Table.csv")
```

##Reference vs. Variant
```{r Linear Modeling}
#We will run these tests separately for cell type
# Refer to the figure_making folder for an MDS plot of biological replicates,
# based on that analysis we will be dropping the following from the modeling
# as they correlated poorly:
# 2R and 5R from Cre+ pre-splicing
# 2R from Cre- pre-splicing
# 2R and 4R from Cre+ post-splicing
# 3R from Cre- post-splicing
d_expr <- fread("./Vglut1-Cre Expression Table.csv")
Vglut1.OFF.Pre.to.test <- fread("./Vglut1_Cre-OFF_Pre_TestingElements.txt", sep = "\t")
Vglut1.ON.Pre.to.test <- fread("./Vglut1_Cre-ON_Pre_TestingElements.txt", sep = "\t")
Vglut1.OFF.Post.to.test <- fread("./Vglut1_Cre-OFF_Post_TestingElements.txt", sep = "\t")
Vglut1.ON.Post.to.test <- fread("./Vglut1_Cre-ON_Post_TestingElements.txt", sep = "\t")
```

###Pre-Splicing Whole Cortex
```{r pre_cortex_refvsalt}
#Pre-Splicing
res <- mpratools.lm(filter(d_expr,
                      Element %in% Vglut1.OFF.Pre.to.test$Element,
                      is.finite(Expression),
                      Cre_Status=="OFF",
                      Splicing_Pool=="Pre",
                      Animal_Hemisphere!="2R",
                      is.finite(Expression)))
fwrite(res,"Pre-Splicing_Vglut1_CreOFF_RefvsAlt_LM.tsv", sep = "\t")

fwrite(
  full_join(
    fread("./Pre-Splicing_Vglut1_CreOFF_RefvsAlt_LM.tsv"),
    fread("../../code/mpra_tools/fullannotationtable.txt"))
  ,"Pre-Splicing_Vglut1_CreOFF_RefvsAlt_LM_Annotated.tsv", sep = "\t")
```

###Pre-Splicing Excitatory Neurons
```{r pre_exn_refvsalt}
res <- mpratools.lm(filter(d_expr,
                      Element %in% Vglut1.ON.Pre.to.test$Element,
                      is.finite(Expression),
                      Cre_Status=="ON",
                      Animal_Hemisphere!="2R",
                      Animal_Hemisphere!="5R",
                      Splicing_Pool=="Pre",
                      is.finite(Expression)))
fwrite(res,"Pre-Splicing_Vglut1_CreON_RefvsAlt_LM.tsv",sep = "\t")
fwrite(
  full_join(
    fread("./Pre-Splicing_Vglut1_CreON_RefvsAlt_LM.tsv"),
    fread("../../code/mpra_tools/fullannotationtable.txt"))
  ,"Pre-Splicing_Vglut1_CreON_RefvsAlt_LM_Annotated.tsv", sep = "\t")
```

###Post-Splicing Whole Cortex
```{r post_cortex_refvsalt}
res <- mpratools.lm(filter(d_expr,
                      Element %in% Vglut1.OFF.Post.to.test$Element,
                      is.finite(Expression),
                      Cre_Status=="OFF",
                      Animal_Hemisphere!="3R",
                      Splicing_Pool=="Post",
                      is.finite(Expression)))
fwrite(res,"Post-Splicing_Vglut1_CreOFF_RefvsAlt_LM.tsv", sep = "\t")
fwrite(
  full_join(
    fread("./Post-Splicing_Vglut1_CreOFF_RefvsAlt_LM.tsv"),
    fread("../../code/mpra_tools/fullannotationtable.txt"))
  ,"Post-Splicing_Vglut1_CreOFF_RefvsAlt_LM_Annotated.tsv", sep = "\t")
```

###Post-Splicing Excitatory Neurons
```{r post_exn_refvsalt}
res <- mpratools.lm(filter(d_expr,
                      Element %in% Vglut1.ON.Post.to.test$Element,
                      is.finite(Expression),
                      Cre_Status=="ON",
                      Animal_Hemisphere!="2R",
                      Animal_Hemisphere!="4R",
                      Splicing_Pool=="Post",
                      is.finite(Expression)))
fwrite(res,"Post-Splicing_Vglut1_CreON_RefvsAlt_LM.tsv", sep = "\t")
fwrite(
  full_join(
    fread("./Post-Splicing_Vglut1_CreON_RefvsAlt_LM.tsv"),
    fread("../../code/mpra_tools/fullannotationtable.txt"))
  ,"Post-Splicing_Vglut1_CreON_RefvsAlt_LM_Annotated.tsv", sep = "\t")
```

### Stability Whole Cortex
```{r stability_cortex}
res <- mpratools.lm.int(d = filter(d_expr,Element %in%
                                               intersect(Vglut1.OFF.Pre.to.test$Element,
                                                         Vglut1.OFF.Post.to.test$Element),
                                             is.finite(Expression),
                                             Cre_Status=="OFF",
                                   !(Animal_Hemisphere=="3R" & Splicing_Pool=="Post"),
                                   !(Animal_Hemisphere=="2R" & Splicing_Pool=="Pre")),
                                  model = Expression ~ Allele * Splicing_Pool)
fwrite(res,"Stability_Vglut1_CreOFF_RefvsAlt_LM.tsv", sep="\t")
fwrite(
  full_join(
    fread("./Stability_Vglut1_CreOFF_RefvsAlt_LM.tsv"),
    fread("../../code/mpra_tools/fullannotationtable.txt"))
  ,"Stability_Vglut1_CreOFF_RefvsAlt_LM_Annotated.tsv", sep = "\t")
```

###Stability Excitatory Neurons
```{r stability_exn}
res <- mpratools.lm.int(d = filter(d_expr,
                                Element %in% intersect(Vglut1.ON.Pre.to.test$Element,
                                                       Vglut1.ON.Post.to.test$Element),
                                is.finite(Expression),
                                Cre_Status=="ON",
                                Animal_Hemisphere!="2R",
                                !(Animal_Hemisphere=="4R" & Splicing_Pool=="Post"),
                                !(Animal_Hemisphere=="5R" & Splicing_Pool=="Pre")),
                     model = Expression ~ Allele * Splicing_Pool)
fwrite(res,"Stability_Vglut1_CreON_RefvsAlt_LM.tsv", sep = "\t")
fwrite(
  full_join(
    fread("./Stability_Vglut1_CreON_RefvsAlt_LM.tsv"),
    fread("../../code/mpra_tools/fullannotationtable.txt"))
  ,"Stability_Vglut1_CreON_RefvsAlt_LM_Annotated.tsv", sep = "\t") 

```

##Reference vs. Shuffled
###Pre-Splicing Whole Cortex
```{r pre_cortex_refvsshuf}

nulls <- as.data.table(filter(d_expr,
                              Allele=="shuf",
                              Cre_Status=="OFF",
                              Splicing_Pool=="Pre",
                              is.finite(Expression)) %>% 
                         group_by(Element) %>% 
                         summarize(Expression=mean(Expression)))
pre_OFF_shuf <- lapply(Vglut1.OFF.Pre.to.test$Element, function(element) {
  welch_shuftest(data_table = filter(d_expr,
                              Cre_Status=="OFF",
                              Splicing_Pool=="Pre",
                              is.finite(Expression)),
                 element = element,
                 allele = "ref",
                 null_values=nulls$Expression)
})
pre_OFF_shuf <- as.data.table(do.call(rbind,pre_OFF_shuf))
setnames(pre_OFF_shuf,,c("Element","logFC","p.val"))
pre_OFF_shuf$fdr <- p.adjust(pre_OFF_shuf$p.val,method = "fdr")
fwrite(pre_OFF_shuf,"Pre-Splicing_Vglut1_CreOFF_RefvsShuf_Welchs.tsv", sep="\t")
fwrite(full_join(pre_OFF_shuf,fread("../../code/mpra_tools/fullannotationtable.txt"))[,!c("Pos","Ref","Alt","Type","Pheno","library_seq_alt","library_seq_shuf")],"Pre-Splicing_Vglut1_CreOFF_RefvsShuf_Welchs_Annotated.tsv", sep="\t")
```

###Pre-Splicing Excitatory Neurons
```{r pre_exn_refvsshuf}
#Pre-Splicing Cre-ON
nulls <- as.data.table(filter(d_expr,
                              Allele=="shuf",
                              Cre_Status=="ON",
                              Splicing_Pool=="Pre",
                              is.finite(Expression)) %>% 
                         group_by(Element) %>% 
                         summarize(Expression=mean(Expression)))
pre_ON_shuf <- lapply(Vglut1.ON.Pre.to.test$Element, function(element) {
  welch_shuftest(data_table = filter(d_expr,
                              Cre_Status=="ON",
                              Splicing_Pool=="Pre",
                              is.finite(Expression)),
                 element = element,
                 allele = "ref",
                 null_values=nulls$Expression)
})
pre_ON_shuf <- as.data.table(do.call(rbind,pre_ON_shuf))
setnames(pre_ON_shuf,,c("Element","logFC","p.val"))
pre_ON_shuf$fdr <- p.adjust(pre_ON_shuf$p.val,method = "fdr")
fwrite(pre_ON_shuf,"Pre-Splicing_Vglut1_CreON_RefvsShuf_Welchs.tsv", sep="\t")
fwrite(full_join(pre_ON_shuf,fread("../../code/mpra_tools/fullannotationtable.txt"))[,!c("Pos","Ref","Alt","Type","Pheno","library_seq_alt","library_seq_shuf")],"Pre-Splicing_Vglut1_CreON_RefvsShuf_Welchs_Annotated.tsv", sep="\t")
```

###Post-Splicing Whole Cortex
```{r post_cortex_refvsshuf}
#Post-Splicing Cre-OFF
nulls <- as.data.table(filter(d_expr,
                              Allele=="shuf",
                              Cre_Status=="OFF",
                              Splicing_Pool=="Post",
                              is.finite(Expression)) %>% 
                         group_by(Element) %>% 
                         summarize(Expression=mean(Expression)))
post_OFF_shuf <- lapply(Vglut1.OFF.Post.to.test$Element, function(element) {
  welch_shuftest(data_table = filter(d_expr,
                              Cre_Status=="OFF",
                              Splicing_Pool=="Post",
                              is.finite(Expression)),
                 element = element,
                 allele = "ref",
                 null_values=nulls$Expression)
})
post_OFF_shuf <- as.data.table(do.call(rbind,post_OFF_shuf))
setnames(post_OFF_shuf,,c("Element","logFC","p.val"))
post_OFF_shuf$fdr <- p.adjust(post_OFF_shuf$p.val,method = "fdr")
fwrite(post_OFF_shuf,"Post-Splicing_Vglut1_CreOFF_RefvsShuf_Welchs.tsv", sep="\t")
fwrite(full_join(post_OFF_shuf,fread("../../code/mpra_tools/fullannotationtable.txt"))[,!c("Pos","Ref","Alt","Type","Pheno","library_seq_alt","library_seq_shuf")],"Post-Splicing_Vglut1_CreOFF_RefvsShuf_Welchs_Annotated.tsv", sep="\t")
```

###Post-Splicing Excitatory Neurons
```{r post_exn_refvsshuf}
#Post-Splicing Cre-ON
nulls <- as.data.table(filter(d_expr,
                              Allele=="shuf",
                              Cre_Status=="ON",
                              Splicing_Pool=="Post",
                              is.finite(Expression)) %>% 
                         group_by(Element) %>% 
                         summarize(Expression=mean(Expression)))
post_ON_shuf <- lapply(Vglut1.ON.Post.to.test$Element, function(element) {
  welch_shuftest(data_table = filter(d_expr,
                              Cre_Status=="ON",
                              Splicing_Pool=="Post",
                              is.finite(Expression)),
                 element = element,
                 allele = "ref",
                 null_values=nulls$Expression)
})
post_ON_shuf <- as.data.table(do.call(rbind,post_ON_shuf))
setnames(post_ON_shuf,,c("Element","logFC","p.val"))
post_ON_shuf$fdr <- p.adjust(post_ON_shuf$p.val,method = "fdr")
fwrite(post_ON_shuf,"Post-Splicing_Vglut1_CreON_RefvsShuf_Welchs.tsv", sep="\t")
fwrite(full_join(post_ON_shuf,fread("../../code/mpra_tools/fullannotationtable.txt"))[,!c("Pos","Ref","Alt","Type","Pheno","library_seq_alt","library_seq_shuf")],"Post-Splicing_Vglut1_CreON_RefvsShuf_Welchs_Annotated.tsv", sep="\t")
```

###Stability Whole Cortex
```{r stability_cortex_refvsshuf}
#We cannot do an interaction model for non-parametric (well, we can, but it's
#complicated), so we are going to subtract pre from post expression to get "stability"
stability_OFF <- inner_join(
  filter(d_expr,Cre_Status=="OFF",Splicing_Pool=="Pre"),
  filter(d_expr,Cre_Status=="OFF",Splicing_Pool=="Post"),
  by=c("Element","Allele","Animal_Hemisphere"))
stability_OFF[,Expression := Expression.y-Expression.x]

elems <- unique(stability_OFF$Element)
nulls <- as.data.table(filter(stability_OFF,Allele=="shuf") %>% group_by(Element) %>%
    summarize(Expression=mean(Expression)))

stab_shuf_OFF <- lapply(elems, function(element) {
  welch_shuftest(data_table = filter(stability_OFF,is.finite(Expression)), element = element, allele = "ref",null_values=is.finite(nulls$Expression))
})
stab_shuf_OFF <- as.data.table(do.call(rbind,stab_shuf_OFF))
setnames(stab_shuf_OFF,,c("Element","logFC","p.val"))
stab_shuf_OFF$fdr <- p.adjust(stab_shuf_OFF$p.val,method = "fdr")
fwrite(stab_shuf_OFF,"Stability_Vglut1_CreOFF_RefvsShuf_Welchs.tsv")
fwrite(full_join(stab_shuf_OFF,fread("../../code/mpra_tools/fullannotationtable.txt"))[,!c("Pos","Ref","Alt","Type","Pheno","library_seq_alt","library_seq_shuf")],"Stability_Vglut1_CreOFF_RefvsShuf_Welchs_Annotated.tsv", sep="\t")
```

###Stability Excitatory Neurons
```{r stability_exn_refvsshuf}

stability_ON <- inner_join(
  filter(d_expr,Cre_Status=="ON",Splicing_Pool=="Pre"),
  filter(d_expr,Cre_Status=="ON",Splicing_Pool=="Post"),
  by=c("Element","Allele","Animal_Hemisphere"))
stability_ON[,Expression := Expression.y-Expression.x]

elems <- unique(stability_ON$Element)
nulls <- as.data.table(filter(stability_ON,Allele=="shuf") %>% group_by(Element) %>%
    summarize(Expression=mean(Expression)))

stab_shuf_ON <- lapply(elems, function(element) {
  welch_shuftest(data_table = filter(stability_ON,is.finite(Expression)), element = element, allele = "ref",null_values=is.finite(nulls$Expression))
})

stab_shuf_ON <- as.data.table(do.call(rbind,stab_shuf_ON))
setnames(stab_shuf_ON,,c("Element","logFC","p.val"))
stab_shuf_ON$fdr <- p.adjust(stab_shuf_ON$p.val,method = "fdr")
fwrite(stab_shuf_ON,"Stability_Vglut1_CreON_RefvsShuf_Welchs.tsv")
fwrite(full_join(stab_shuf_ON,fread("../../code/mpra_tools/fullannotationtable.txt"))[,!c("Pos","Ref","Alt","Type","Pheno","library_seq_alt","library_seq_shuf")],"Stability_Vglut1_CreON_RefvsShuf_Welchs_Annotated.tsv", sep="\t")
```


##Sex Specific Effects
###Define Three-Term Interaction Model
```{r threeway_int_function}
library(emmeans)
d_expr <- fread("./Vglut1-Cre Expression Table.csv")

mpratools.lm.int.threeway <- function(dt, # The data table in question
                          response = "Expression", # The value you want to examine, usually Expression
                          factor1 = "Allele", # The first predictor, usually Allele
                          factor2 = "Splicing_Pool", # The second predictor, usually Splicing_Pool
                          factor3 = "Sex", # The third predictor, usually Sex
                          levels_factor1 = c("ref", "alt"), # Order of predictor1, usually c("ref", "alt")
                          levels_factor2 = c("Pre", "Post"), # Order of predictor2, usually c("Pre", "Post")
                          levels_factor3 = c("m", "f"), # Order of predictor3, usually c("m", "f")
                          separator = "Element") { # What to separate on, usually Element
  # Ensure data is in data.table format
  setDT(dt)
  
  # Set factor levels if provided
  dt[, (factor1) := factor(get(factor1), levels = levels_factor1)]
  dt[, (factor2) := factor(get(factor2), levels = levels_factor2)]
  dt[, (factor3) := factor(get(factor3), levels = levels_factor3)]
  
  # Construct the model formula
  formula <- as.formula(paste(response, "~", factor1, "*", factor2, "*", factor3))
  
  # Define the result columns dynamically
  coef_cols <- c(
    paste0("logFC_", c(factor1, factor2, factor3,
                       paste0(factor1, ":", factor2),
                       paste0(factor1, ":", factor3),
                       paste0(factor2, ":", factor3),
                       paste0(factor1, ":", factor2, ":", factor3))),
    paste0("pval_", c(factor1, factor2, factor3,
                      paste0(factor1, ":", factor2),
                      paste0(factor1, ":", factor3),
                      paste0(factor2, ":", factor3),
                      paste0(factor1, ":", factor2, ":", factor3)))
  )
  
  # Initialize an empty results table with proper column types
  n_elements <- length(unique(dt[[separator]]))
  res_table <- data.table(
    separator_col = character(n_elements)
  )
  setnames(res_table, "separator_col", separator)
  
  # Add numeric columns for logFC and p-values
  for(col in coef_cols) {
    res_table[, (col) := rep(NA_real_, n_elements)]
  }
  
  # Assign separator values
  res_table[[separator]] <- unique(dt[[separator]])
  
  # Loop over each unique separator value
  for (i in seq_len(nrow(res_table))) {
    sep_value <- res_table[[separator]][i]
    # Subset the data for the current separator
    dt_subset <- dt[get(separator) == sep_value]
    
    # Check if there are at least 3 measurements per combination
    valid_combinations <- dt_subset[, .N, by = .(get(factor1), get(factor2), get(factor3))][N >= 3]
    if (nrow(valid_combinations) < 8) {
      next  # Skip this separator if fewer than 8 valid combinations
    }
    
    # Try fitting the model
    tryCatch({
      model <- lm(formula, data = dt_subset)
      coef_values <- as.data.table(coef(summary(model)))
      
      # Get the number of coefficients (excluding intercept)
      n_coef <- nrow(coef_values) - 1
      
      # Extract logFC values (coefficients)
      logfc_cols <- grep("^logFC_", names(res_table), value = TRUE)
      for (col_idx in seq_along(logfc_cols)) {
        res_table[i, (logfc_cols[col_idx]) := coef_values[col_idx + 1, 1]]
      }
      
      # Extract p-values
      pval_cols <- grep("^pval_", names(res_table), value = TRUE)
      for (col_idx in seq_along(pval_cols)) {
        res_table[i, (pval_cols[col_idx]) := coef_values[col_idx + 1, 4]]
      }
      
    }, error = function(e) {
      # On error, fill all coefficient columns with NA
      coef_cols <- names(res_table)[-1]  # All columns except separator
      res_table[i, (coef_cols) := NA_real_]
    })
  }
  
  # Calculate FDR and Bonferroni corrections for each p-value column
  pval_cols <- grep("^pval_", names(res_table), value = TRUE)
  for(pval_col in pval_cols) {
    # Create column names for corrections
    fdr_col <- gsub("^pval_", "fdr_", pval_col)
    bonf_col <- gsub("^pval_", "bonf_", pval_col)
    
    # Calculate corrections
    res_table[, (fdr_col) := p.adjust(get(pval_col), method = "fdr")]
    res_table[, (bonf_col) := p.adjust(get(pval_col), method = "bonferroni")]
  }
  
  return(res_table)
}
```

###Sex-Specific Whole Cortex Variants
```{r run_threeway_cortex}
res <- mpratools.lm.int.threeway(d = filter(d_expr,Element %in%
                                               intersect(Vglut1.OFF.Pre.to.test$Element,
                                                         Vglut1.OFF.Post.to.test$Element),
                                             is.finite(Expression),
                                             Cre_Status=="OFF",
                                   !(Animal_Hemisphere=="3R" & Splicing_Pool=="Post"),
                                   !(Animal_Hemisphere=="2R" & Splicing_Pool=="Pre")))
fwrite(res,"Stability_Vglut1_CreOFF_AllelevsSplicevsSex_LM.tsv", sep="\t")
fwrite(
  full_join(
    fread("./Stability_Vglut1_CreOFF_AllelevsSplicevsSex_LM.tsv"),
    fread("../../code/mpra_tools/fullannotationtable.txt"))
  ,"Stability_Vglut1_CreOFF_AllelevsSplicevsSex_LM_Annotated.tsv", sep = "\t")
```

###Sex-Specific Excitatory Neurons Variants
```{r run_threeway_exn}
res <- mpratools.lm.int.threeway(d = filter(d_expr,
                                Element %in% intersect(Vglut1.ON.Pre.to.test$Element,
                                                       Vglut1.ON.Post.to.test$Element),
                                is.finite(Expression),
                                Cre_Status=="ON",
                                Animal_Hemisphere!="2R",
                                !(Animal_Hemisphere=="4R" & Splicing_Pool=="Post"),
                                !(Animal_Hemisphere=="5R" & Splicing_Pool=="Pre")))

fwrite(res,"Stability_Vglut1_CreON_AllelevsSplicevsSex_LM.tsv", sep = "\t")
fwrite(
  full_join(
    fread("./Stability_Vglut1_CreON_AllelevsSplicevsSex_LM.tsv"),
    fread("../../code/mpra_tools/fullannotationtable.txt"))
  ,"Stability_Vglut1_CreON_AllelevsSplicevsSex_LM_Annotated.tsv", sep = "\t") 
```

###Element-Level Expression Sex Differences

```{r}
#Further delving into sex-specific effects, we will use mpratools.lm to
#investigate element by element expression differences, but use tidyr::unite
#to join element and allele
#so we can feed it the whole data table, then separate later
d_expr <- fread("./Vglut1-Cre Expression Table.csv")

Vglut1.OFF.Pre.to.test <- fread("./Vglut1_Cre-OFF_Pre_TestingElements.txt", sep = "\t")
Vglut1.ON.Pre.to.test <- fread("./Vglut1_Cre-ON_Pre_TestingElements.txt", sep = "\t")
Vglut1.OFF.Post.to.test <- fread("./Vglut1_Cre-OFF_Post_TestingElements.txt", sep = "\t")
Vglut1.ON.Post.to.test <- fread("./Vglut1_Cre-ON_Post_TestingElements.txt", sep = "\t")
```

####Male vs. Female Pre-Splicing Whole Cortex
```{r mvsf_pre_cortex}
res <- mpratools.lm(filter(d_expr,
                      Element %in% Vglut1.OFF.Pre.to.test$Element,
                      is.finite(Expression),
                      Cre_Status=="OFF",
                      Animal_Hemisphere!="2R",
                      Splicing_Pool=="Pre",
                      is.finite(Expression)) %>%
                      tidyr::unite(.,Element,c("Element",
                                        "Allele"),sep="@"),
                    Expression ~ Sex,,
                    "Sex",
                    "m",
                    "f")
res <- tidyr::separate(res,Element,c("Element","Allele"),sep="@")
fwrite(res,"Pre-Splicing_Vglut1_CreOFF_MvsF_LM.tsv", sep = "\t")

fwrite(
  full_join(
    fread("./Pre-Splicing_Vglut1_CreOFF_MvsF_LM.tsv"),
    fread("../../code/mpra_tools/fullannotationtable.txt"))
  ,"Pre-Splicing_Vglut1_CreOFF_MvsF_LM_Annotated.tsv", sep = "\t")
```

####Male vs. Female Pre-Splicing Excitatory Neurons
```{r mvsf_pre_exn}
res <- mpratools.lm(filter(d_expr,
                      Element %in% Vglut1.ON.Pre.to.test$Element,
                      is.finite(Expression),
                      Cre_Status=="ON",
                      Splicing_Pool=="Pre",
                      Animal_Hemisphere!="2R",
                      Animal_Hemisphere!="5R",
                      is.finite(Expression)) %>%
                      tidyr::unite(.,Element,c("Element",
                                        "Allele"),sep="@"),
                    Expression ~ Sex,,
                    "Sex",
                    "m",
                    "f")
res <- tidyr::separate(res,Element,c("Element","Allele"),sep="@")
fwrite(res,"Pre-Splicing_Vglut1_CreON_MvsF_LM.tsv",sep = "\t")
fwrite(
  full_join(
    fread("./Pre-Splicing_Vglut1_CreON_MvsF_LM.tsv"),
    fread("../../code/mpra_tools/fullannotationtable.txt"))
  ,"Pre-Splicing_Vglut1_CreON_MvsF_LM_Annotated.tsv", sep = "\t")
```

####Male vs. Female Post-Splicing Whole Cortex
```{r mvsf_post_cortex}
res <- mpratools.lm(filter(d_expr,
                      Element %in% Vglut1.OFF.Post.to.test$Element,
                      is.finite(Expression),
                      Cre_Status=="OFF",
                      Splicing_Pool=="Post",
                      Animal_Hemisphere!="3R",
                      is.finite(Expression)) %>%
                      tidyr::unite(.,Element,c("Element",
                                        "Allele"),sep="@"),
                    Expression ~ Sex,,
                    "Sex",
                    "m",
                    "f")
res <- tidyr::separate(res,Element,c("Element","Allele"),sep="@")
fwrite(res,"Post-Splicing_Vglut1_CreOFF_MvsF_LM.tsv", sep = "\t")
fwrite(
  full_join(
    fread("./Post-Splicing_Vglut1_CreOFF_MvsF_LM.tsv"),
    fread("../../code/mpra_tools/fullannotationtable.txt"))
  ,"Post-Splicing_Vglut1_CreOFF_MvsF_LM_Annotated.tsv", sep = "\t")
```

####Male vs. Female Post-Splicing Excitatory Neurons
```{r mvsf_post_exn}
res <- mpratools.lm(filter(d_expr,
                      Element %in% Vglut1.ON.Post.to.test$Element,
                      is.finite(Expression),
                      Cre_Status=="ON",
                      Splicing_Pool=="Post",
                      Animal_Hemisphere!="2R",
                      Animal_Hemisphere!="4R",
                      is.finite(Expression)) %>%
                      tidyr::unite(.,Element,c("Element",
                                        "Allele"),sep="@"),
                    Expression ~ Sex,,
                    "Sex",
                    "m",
                    "f")
res <- tidyr::separate(res,Element,c("Element","Allele"),sep="@")
fwrite(res,"Post-Splicing_Vglut1_CreON_MvsF_LM.tsv", sep = "\t")
fwrite(
  full_join(
    fread("./Post-Splicing_Vglut1_CreON_MvsF_LM.tsv"),
    fread("../../code/mpra_tools/fullannotationtable.txt"))
  ,"Post-Splicing_Vglut1_CreON_MvsF_LM_Annotated.tsv", sep = "\t")
```


###Interaction Modeling of Allele and Sex on Expression

```{read_data}
################################################################################
#Now we can do the allele*sex interaction
d_expr <- fread("./Vglut1-Cre Expression Table.csv")
#Set the sex as a factor for the interaction modeling
d_expr$Sex <- factor(d_expr$Sex,levels = c("m","f"))
Vglut1.OFF.Pre.to.test <- fread("./Vglut1_Cre-OFF_Pre_TestingElements.txt")
Vglut1.ON.Pre.to.test <- read.csv("./Vglut1_Cre-ON_Pre_TestingElements.txt")
Vglut1.OFF.Post.to.test <- fread("./Vglut1_Cre-OFF_Post_TestingElements.txt")
Vglut1.ON.Post.to.test <- fread("./Vglut1_Cre-ON_Post_TestingElements.txt")
```

####Pre-Splicing Whole Cortex
```{r pre_allele-sex_cortex}
res <- mpratools.lm.int(filter(d_expr,
                      Element %in% Vglut1.OFF.Pre.to.test$Element,
                      is.finite(Expression),
                      Cre_Status=="OFF",
                      Animal_Hemisphere!="2R",
                      Splicing_Pool=="Pre",
                      is.finite(Expression)),
                    Expression ~ Allele*Sex,,
                    "Allele",
                    "ref",
                    "alt")
fwrite(res,"Pre-Splicing_Vglut1_CreOFF_AllelevsSex_LM.tsv", sep = "\t")
fwrite(
  full_join(
    fread("./Pre-Splicing_Vglut1_CreOFF_AllelevsSex_LM.tsv"),
    fread("../../code/mpra_tools/fullannotationtable.txt"))
  ,"Pre-Splicing_Vglut1_CreOFF_AllelevsSex_LM_Annotated.tsv", sep = "\t")
```

####Pre-Splicing Excitatory Neurons
```{r pre_allele-sex_exn}
res <- mpratools.lm.int(filter(d_expr,
                      Element %in% Vglut1.ON.Pre.to.test$Element,
                      is.finite(Expression),
                      Cre_Status=="ON",
                      Splicing_Pool=="Pre",
                      Animal_Hemisphere!="2R",
                      Animal_Hemisphere!="5R",
                      is.finite(Expression)),
                    Expression ~ Allele*Sex,,
                    "Allele",
                    "ref",
                    "alt")
fwrite(res,"Pre-Splicing_Vglut1_CreON_AllelevsSex_LM.tsv",sep = "\t")
fwrite(
  full_join(
    fread("./Pre-Splicing_Vglut1_CreON_AllelevsSex_LM.tsv"),
    fread("../../code/mpra_tools/fullannotationtable.txt"))
  ,"Pre-Splicing_Vglut1_CreON_AllelevsSex_LM_Annotated.tsv", sep = "\t")
```

####Post-Splicing Whole Cortex
```{r post_allele-sex_cortex}
res <- mpratools.lm.int(filter(d_expr,
                      Element %in% Vglut1.OFF.Post.to.test$Element,
                      is.finite(Expression),
                      Cre_Status=="OFF",
                      Splicing_Pool=="Post",
                      Animal_Hemisphere!="3R",
                      is.finite(Expression)),
                    Expression ~ Allele*Sex,,
                    "Allele",
                    "ref",
                    "alt")
fwrite(res,"Post-Splicing_Vglut1_CreOFF_AllelevsSex_LM.tsv", sep = "\t")
fwrite(
  full_join(
    fread("./Post-Splicing_Vglut1_CreOFF_AllelevsSex_LM.tsv"),
    fread("../../code/mpra_tools/fullannotationtable.txt"))
  ,"Post-Splicing_Vglut1_CreOFF_AllelevsSex_LM_Annotated.tsv", sep = "\t")
```

####Post-Splicing Excitatory Neurons
```{r post_allele-sex_exn}
res <- mpratools.lm.int(filter(d_expr,
                      Element %in% Vglut1.ON.Post.to.test$Element,
                      is.finite(Expression),
                      Cre_Status=="ON",
                      Splicing_Pool=="Post",
                      Animal_Hemisphere!="2R",
                      Animal_Hemisphere!="4R",
                      is.finite(Expression)),
                    Expression ~ Allele*Sex,,
                    "Allele",
                    "ref",
                    "alt")
fwrite(res,"Post-Splicing_Vglut1_CreON_AllelevsSex_LM.tsv", sep = "\t")
fwrite(
  full_join(
    fread("./Post-Splicing_Vglut1_CreON_AllelevsSex_LM.tsv"),
    fread("../../code/mpra_tools/fullannotationtable.txt"))
  ,"Post-Splicing_Vglut1_CreON_AllelevsSex_LM_Annotated.tsv", sep = "\t")
```

###Excitatory Neurons vs. Whole Cortex Variant vs. Reference Stability Differences

```{r read_data}
#ILook at ExN vs. Cortex
#Run three way, see if there are any effects
d_expr <- fread("./Vglut1-Cre Expression Table.csv")
d_expr$Sex <- factor(d_expr$Sex,levels = c("m","f"))
Vglut1.OFF.Pre.to.test <- fread("./Vglut1_Cre-OFF_Pre_TestingElements.txt")
Vglut1.ON.Pre.to.test <- read.csv("./Vglut1_Cre-ON_Pre_TestingElements.txt")
Vglut1.OFF.Post.to.test <- fread("./Vglut1_Cre-OFF_Post_TestingElements.txt")
Vglut1.ON.Post.to.test <- fread("./Vglut1_Cre-ON_Post_TestingElements.txt")
```

####Excitatory Neuron vs. Whole Cortex
```{r }
res <- mpratools.lm.int.threeway(d = filter(d_expr,
                                Element %in% 
                                  intersect(
                                    intersect(Vglut1.ON.Pre.to.test$Element,
                                              Vglut1.ON.Post.to.test$Element),
                                    intersect(Vglut1.OFF.Pre.to.test$Element,
                                              Vglut1.OFF.Post.to.test$Element)
                                    ),
                                is.finite(Expression),
                                !(Animal_Hemisphere=="3R" & Splicing_Pool=="Post"),
                                !(Animal_Hemisphere=="2R" & Splicing_Pool=="Pre"),
                                Animal_Hemisphere!="2R",
                                !(Animal_Hemisphere=="4R" & Splicing_Pool=="Post"),
                                !(Animal_Hemisphere=="5R" & Splicing_Pool=="Pre")
                                ),
                                factor1 = "Allele",
                                factor2 = "Splicing_Pool",
                                factor3 = "Cre_Status",
                                levels_factor3 = c("OFF","ON")
)


fwrite(res, "Vglut1_AllelevsSplicevsCreStatus_LM.tsv", sep="\t")
fwrite(
  full_join(
    fread("./Vglut1_AllelevsSplicevsCreStatus_LM.tsv"),
    fread("../../code/mpra_tools/fullannotationtable.txt")),
  "./Vglut1_AllelevsSplicevsCreStatus_LM_Annotated.tsv", sep="\t")
``` 