---
title: "Library QC and Organizing"
author: "Din Selmanovic"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
---

<!--
Library QC and Organizing
--------------------------
This document is the quality-control (QC) and data-organizing step of the
in vivo / in vitro MPRA pipeline. It takes the raw per-sample alignment count
tables produced by the upstream alignment/counting bash scripts and:

  1. Reads in and harmonizes the ISO-MPRA (splicing-aware) count tables and the
     eMPRA (barcode-level) count tables.
  2. Computes per-sample alignment / counting QC metrics.
  3. Computes replicate-to-replicate correlations as a sanity check.
  4. Reshapes everything into the long 'UTR' format and writes the .rds / .csv
     files that the downstream analysis notebook consumes.

Expected inputs (in the working directory):
  - ms*.txt              : ISO-MPRA per-sample count tables (ID + count)
  - *_foundBCs.txt       : eMPRA found-barcode lists, one per sample
  - trevino.names.csv    : barcode <-> rsID map for the eMPRA library
  - nreads.txt           : total read counts per ISO-MPRA sample
  - nreads.empra.txt     : total read counts per eMPRA sample
  - corr.values.trevino*.csv : pre-computed correlation summaries for plotting
  - meta.file3.trev.rds  : library metadata (barcode_allele -> ref/alt)

Outputs:
  - reads.table.trevino.isompra.csv / reads.table.trevino.empra.csv (QC tables)
  - invivo.ps1.utr.format.*.rds/csv  (ISO-MPRA long format)
  - empratrev.utr.format.*.rds/csv   (eMPRA long format)
  - corr.*.csv (replicate correlation matrices)

NOTE: a couple of chunks read a library-metadata file from an absolute Windows
path (C:/Users/dinse/...). Update that path before running on another machine.
-->

```{r setup, include=FALSE}
# Global knitr options: show code, but hide package start-up warnings/messages.
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Setup

## Packages and helper functions

Load the packages used throughout the QC pipeline and define a few small helper
functions for reshaping correlation matrices, cleaning element IDs, and
converting raw counts to counts-per-million (CPM).

```{r load-packages-and-helpers, include=FALSE}
library(Biostrings)
library(tidyr)
library(ggplot2)
library(data.table)
library(dplyr)
library(readr)
library(stringr)
library(mgsub)
library(psych)
library(Hmisc)
library(limma)
library(data.table)
library(ggpmisc)

flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
#To get this to work with downstream stuff, I need to remove the "=*" portions
extract_and_replace <- function(x) {
  rs_numbers <- str_extract_all(x, "rs\\d+")
  if (length(rs_numbers) > 0) {
    str_replace_all(x, ".*?(rs\\d+).*", "\\1")
  } else {
    x
  }
}

apply_operation <- function(data, column_names) {
  for (col_name in column_names) {
    # Calculate the sum for the specified column
    col_sum <- sum(data[[col_name]])
    # Apply the operation to the column
    data[, paste0(col_name, ".cpm") := (data[[col_name]] / col_sum) * 1e6]
  }
  return(data)
}
```

# Reading and assembling count tables

## ISO-MPRA count tables

Read every `ms*.txt` ISO-MPRA count table, standardize the element IDs (strip
alignment prefixes, tag the scrambled/control elements), sum the technical
re-sequencing runs (e.g. `ms1` + `ms1.1`) into a single post-splice count, then
merge all samples into a single wide `merged_data` table keyed by `rsID`.

```{r assemble-isompra-counts, include=FALSE}
#The following is now going to bring in the ISOMPRA format files which are already in a table as the bash script aligns and then counts for us.
files <- list.files(pattern = "ms.*.txt")
files <- files[!grepl("foundBCS\\.txt",files)]
# Loop through each file
for (file in files) {
  # Read the data from the file (adjust read.table parameters as needed)
  data <- fread(file, header = TRUE, fill = TRUE)
  # Apply the specified transformations to the 'ID' column
  data$ID <- sub(".*\\|", "", data$ID)
  data$ID <- ifelse(grepl("^ X", data$ID), paste0("trevino.ctrl", data$ID), data$ID)
  data$ID <- sub("trevino.ctrl\\s*X", "trevino.ctrlX", data$ID)
  data$ID <- gsub("\\s+", "", data$ID)
  data$ID <- as.character(data$ID)
  data$ID <- ifelse(grepl("rs\\d+", data$ID) & !grepl("=", data$ID),
                    paste0("scrambled", data$ID), data$ID)
  # Save the modified data to the environment with a unique name
  assign(paste0("modified_", tools::file_path_sans_ext(file)), data)
}

modified_ms1.post$count<-modified_ms1.post$count+modified_ms1.1.post$count
modified_ms2.post$count <- modified_ms2.post$count + modified_ms2.1.post$count
modified_ms3.post$count <- modified_ms3.post$count + modified_ms3.1.post$count
modified_ms4.post$count <- modified_ms4.post$count + modified_ms4.1.post$count
modified_ms5.post$count <- modified_ms5.post$count + modified_ms5.1.post$count
modified_ms8.post$count <- modified_ms8.post$count + modified_ms8.1.post$count
modified_ms12.post$count <- modified_ms12.post$count + modified_ms12.1.post$count
modified_ms13.post$count <- modified_ms13.post$count + modified_ms13.1.post$count
modified_ms14.post$count <- modified_ms14.post$count + modified_ms14.1.post$count
modified_ms15.post$count <- modified_ms15.post$count + modified_ms15.1.post$count
modified_ms17.post$count <- modified_ms17.post$count + modified_ms17.1.post$count
modified_ms18.post$count <- modified_ms18.post$count + modified_ms18.1.post$count
modified_ms19.post$count <- modified_ms19.post$count + modified_ms19.1.post$count
modified_ms20.post$count <- modified_ms20.post$count + modified_ms20.1.post$count
modified_ms21.post$count <- modified_ms21.post$count + modified_ms21.1.post$count
modified_ms22.post$count <- modified_ms22.post$count + modified_ms22.1.post$count

rm(modified_ms1.1.post, modified_ms2.1.post, modified_ms3.1.post, modified_ms4.1.post, 
   modified_ms5.1.post, modified_ms8.1.post, modified_ms12.1.post, modified_ms13.1.post, 
   modified_ms14.1.post, modified_ms15.1.post, modified_ms17.1.post, modified_ms18.1.post, 
   modified_ms19.1.post, modified_ms20.1.post, modified_ms21.1.post, modified_ms22.1.post)

objects <- ls(pattern = "modified_")

# Initialize an empty list to store data.tables
data_tables <- list()

# Loop through each object with "modified" in its name
for (object in objects) {
  # Convert the object to a data.table
  dt <- as.data.table(get(object))
  
  # Add the data.table to the list
  data_tables[[object]] <- dt
}

# Merge data.tables by 'ID'
merged_data <- data_tables[[1]]  # Initialize with the first data.table
for (i in seq_along(data_tables)[-1]) {
  # Merge each subsequent data.table
  merged_data <- merge(merged_data, data_tables[[i]], by = "ID", all = TRUE, suffixes = c("", paste0("_", objects[i])))
}

#Renaming the merged file
names(merged_data)<-c("rsID","ms1","ms12","ms13","ms14","ms15","ms17","ms18","ms19","ms2","ms20","ms21","ms22","ms3","ms4","ms5","ms6","ms7","ms8","ms9","aav","plasmid","ms1.post","ms12.post","ms13.post","ms14.post","ms15.post","ms17.post","ms18.post","ms19.post","ms2.post","ms20.post","ms21.post","ms22.post","ms3.post","ms4.post","ms5.post","ms8.post")

#Bringing in the total reads so that we can align and count how many percent aligned
totalreads<-fread("nreads.txt")

rm(dt,data_tables,data,curcounts,curfinds,modified_ms.1,modified_ms.12,modified_ms.13,modified_ms.14,modified_ms.15,modified_ms.17,modified_ms.18,modified_ms.19,modified_ms.2,modified_ms.20,modified_ms.21,modified_ms.22,modified_ms.3,modified_ms.4,modified_ms.5,modified_ms.6,modified_ms.7,modified_ms.8,modified_ms.9,modified_ms.aav,modified_ms.plasmid)

rm(modified_ms1.post, modified_ms2.post, modified_ms3.post, modified_ms4.post, 
   modified_ms5.post, modified_ms8.post, modified_ms12.post, modified_ms13.post, 
   modified_ms14.post, modified_ms15.post, modified_ms17.post, modified_ms18.post, 
   modified_ms19.post, modified_ms20.post, modified_ms21.post, modified_ms22.post)
```

## eMPRA barcode count tables

Read the per-sample found-barcode lists, reverse-complement the barcodes so they
match the library orientation, associate each barcode with its element via
`trevino.names.csv`, and sum technical re-sequencing runs into one count per
sample. The result is `SIC_countstrevino`, a wide barcode-level count table.

```{r assemble-empra-counts, include=FALSE}
#These lines of code read in the found BCs and then rev comp them to match the library so we can associate the BC with the correct Element
SIC_countstrevino<-fread("trevino.names.csv",header=F,fill=T)
colnames(SIC_countstrevino)<-c("revcomp_barcode","rsID")
SIC_countstrevino$revcomp_barcode<- sapply(SIC_countstrevino$revcomp_barcode, function(x) as.character(reverseComplement(DNAString(x))))
bcfinds <- list.files(pattern='BCS.txt',full.names = TRUE)
i<-1
for (i in c(1:length(bcfinds))){  
  curfinds <- fread(paste0(bcfinds[i]),header=FALSE)  
  curcounts <- as.data.table(table(curfinds))  
  setnames(curcounts,c(1:2),c("revcomp_barcode", gsub(bcfinds[i],pattern="^(.*)_foundBCs.txt",replacement="\\1")))  
  SIC_countstrevino <- merge.data.table(curcounts,SIC_countstrevino,by="revcomp_barcode",all.y=TRUE)
}

rm(curcounts,curfinds)

SIC_counts2<-SIC_countstrevino

SIC_counts2$aav <- SIC_counts2$'./aav.fastq_foundBCS.txt' + SIC_counts2$'./aav.fastq_foundBCS.txt'
SIC_counts2$ms1 <- SIC_counts2$'./ms1.fastq_foundBCS.txt' + SIC_counts2$'./ms1.1.fastq_foundBCS.txt'
SIC_counts2$ms2 <- SIC_counts2$'./ms2.fastq_foundBCS.txt' + SIC_counts2$'./ms2.1.fastq_foundBCS.txt'
SIC_counts2$ms3 <- SIC_counts2$'./ms3.fastq_foundBCS.txt' + SIC_counts2$'./ms3.1.fastq_foundBCS.txt'
SIC_counts2$ms4 <- SIC_counts2$'./ms4.fastq_foundBCS.txt' + SIC_counts2$'./ms4.1.fastq_foundBCS.txt'
SIC_counts2$ms5 <- SIC_counts2$'./ms5.fastq_foundBCS.txt' + SIC_counts2$'./ms5.1.fastq_foundBCS.txt'
SIC_counts2$ms6 <- SIC_counts2$'./ms6.fastq_foundBCS.txt' + SIC_counts2$'./ms6.1.fastq_foundBCS.txt'
SIC_counts2$ms7 <- SIC_counts2$'./ms7.fastq_foundBCS.txt' + SIC_counts2$'./ms7.1.fastq_foundBCS.txt'
SIC_counts2$ms8 <- SIC_counts2$'./ms8.fastq_foundBCS.txt' + SIC_counts2$'./ms8.1.fastq_foundBCS.txt'
SIC_counts2$ms8.2 <- SIC_counts2$'./ms8.2.fastq_foundBCS.txt' + SIC_counts2$'./ms8.2.1.fastq_foundBCS.txt'
SIC_counts2$ms9 <- SIC_counts2$'./ms9.fastq_foundBCS.txt' + SIC_counts2$'./ms9.1.fastq_foundBCS.txt'
SIC_counts2$ms10 <- SIC_counts2$'./ms10.fastq_foundBCS.txt' + SIC_counts2$'./ms10.1.fastq_foundBCS.txt'
SIC_counts2$ms11 <- SIC_counts2$'./ms11.fastq_foundBCS.txt' + SIC_counts2$'./ms11.1.fastq_foundBCS.txt'
SIC_counts2$ms12 <- SIC_counts2$'./ms12.fastq_foundBCS.txt' + SIC_counts2$'./ms12.1.fastq_foundBCS.txt'
SIC_counts2$ms13 <- SIC_counts2$'./ms13.fastq_foundBCS.txt' + SIC_counts2$'./ms13.1.fastq_foundBCS.txt'
SIC_counts2$ms14 <- SIC_counts2$'./ms14.fastq_foundBCS.txt' + SIC_counts2$'./ms14.1.fastq_foundBCS.txt'
SIC_counts2$ms15 <- SIC_counts2$'./ms15.fastq_foundBCS.txt' + SIC_counts2$'./ms15.1.fastq_foundBCS.txt'
SIC_counts2$ms16 <- SIC_counts2$'./ms16.fastq_foundBCS.txt' + SIC_counts2$'./ms16.1.fastq_foundBCS.txt'
SIC_counts2$ms17 <- SIC_counts2$'./ms17.fastq_foundBCS.txt' + SIC_counts2$'./ms17.1.fastq_foundBCS.txt'
SIC_counts2$ms18 <- SIC_counts2$'./ms18.fastq_foundBCS.txt' + SIC_counts2$'./ms18.1.fastq_foundBCS.txt'
SIC_counts2$ms19 <- SIC_counts2$'./ms19.fastq_foundBCS.txt' + SIC_counts2$'./ms19.1.fastq_foundBCS.txt'
SIC_counts2$ms20 <- SIC_counts2$'./ms20.fastq_foundBCS.txt' + SIC_counts2$'./ms20.1.fastq_foundBCS.txt'
SIC_counts2$ms21 <- SIC_counts2$'./ms21.fastq_foundBCS.txt' + SIC_counts2$'./ms21.1.fastq_foundBCS.txt'
SIC_counts2$ms22 <- SIC_counts2$'./ms22.fastq_foundBCS.txt' + SIC_counts2$'./ms22.1.fastq_foundBCS.txt'
SIC_counts2$plasmid <- SIC_counts2$'./plasmid.fastq_foundBCS.txt' + SIC_counts2$'./plasmid.1.fastq_foundBCS.txt'


SIC_countstrevino<-SIC_counts2[,c(1,54:61,63:77,53,52)]

totalreads.empra<-fread("nreads.empra.txt")

rm(SIC_counts2)
```

# Alignment QC

## ISO-MPRA percent alignment

For each ISO-MPRA sample, compute reads aligned, percent aligned (aligned /
total reads), and the fraction of the 466 library elements that were detected.

```{r qc-percent-alignment-isompra, include=TRUE}
trevino.meta.file<-read_csv("trevino.names.csv",col_names=F)

trevino.meta.file<-trevino.meta.file[,c(2)]
trevino.meta.file<-distinct(trevino.meta.file)
names(trevino.meta.file)<-c("rsID")
trevino<-merged_data
trevino[trevino == 0] <- NA
trevino<-inner_join(trevino.meta.file,trevino,by="rsID")

mappedcounts<-as.data.table(colSums(trevino[,c(2:38)],na.rm=T),keep.rownames=T)
names(mappedcounts)<-c("sample","V2")
totalreadsout<-merge.data.table(totalreads,mappedcounts,by="sample")

setnames(totalreadsout,3,"reads_aligned")
totalreadsout[,pctalign:=(reads_aligned/counts)*100]

i<-0
for (i in c(0:nrow(totalreadsout))){  
  cursample <- names(trevino)[i+1]  
  counted <- sum(!is.na(trevino[,(i+1),with=FALSE])) 
  totalreadsout[sample==cursample, BCs_counted:=counted]
  totalreadsout[sample == cursample, pctElems_Counted := counted / 466 * 100]
}

totalreadsout.trevino.isompra<-totalreadsout
print(totalreadsout.trevino.isompra)
write.csv(totalreadsout.trevino.isompra,"reads.table.trevino.isompra.csv")

rm(mappedcounts,totalreadsout)
```

## eMPRA percent alignment

Same QC metrics for the eMPRA samples (here the denominator is the ~9,400
barcodes in the library), plus a sanity-check boxplot of the pre-computed
replicate correlation values.

```{r qc-percent-alignment-empra, include=TRUE}
mappedcounts<-as.data.table(colSums(SIC_countstrevino[,c(2:25)],na.rm=T),keep.rownames=T)
names(mappedcounts)<-c("sample","V2")
totalreadsout<-merge.data.table(totalreads.empra,mappedcounts,by="sample")

setnames(totalreadsout,3,"reads_aligned")
totalreadsout[,pctalign:=(reads_aligned/counts)*100]

i<-1
for (i in c(1:nrow(totalreadsout))){  
  cursample <- names(SIC_countstrevino)[i+1]  
  counted <- sum(!is.na(SIC_countstrevino[,(i+1),with=FALSE]))  
  totalreadsout[sample==cursample, BCs_counted:=counted]
  totalreadsout[sample == cursample, pctBCs_Counted := counted / 9400 * 100]
}

totalreadsout.trevino<-totalreadsout
print(totalreadsout.trevino)
write.csv(totalreadsout.trevino,"reads.table.trevino.empra.csv")

corr.values.trevino.sanity <- read_csv("corr.values.trevino.sanity.csv")

ggplot(corr.values.trevino.sanity,aes(x=factor(MPRA),y=Value,fill=factor(MPRA)))+geom_boxplot()+geom_dotplot(binaxis='y',stackdir='center',position=position_dodge(0.75),dotsize=0.5)+facet_wrap(.~Type,scale="free_y")+scale_fill_manual(values=c("#E1BE6A","#8e3636","#8E3636")) + theme(legend.position = "bottom")
```

# Replicate correlations

## ISO-MPRA replicate correlations

Drop elements with any zero counts, convert to CPM, and look at replicate
agreement both on the CPM scale and on the log2 expression scale
(log2(RNA CPM / AAV DNA CPM)). Correlation matrices are written to disk and an
MDS plot summarizes overall replicate structure.

```{r correlations-isompra, include=TRUE}
trevino<-merged_data
trevino<-trevino[,c(1:15,17:19,21,23:38)]

trevino <- trevino[rowSums(trevino[,2:35] == 0) == 0, ]
columns_list <- colnames(trevino)[-which(colnames(trevino) == "rsID")]
trevino<-apply_operation(trevino, columns_list)

pairs.panels(trevino[,c(36:69)], 
             method = "spearman", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE,
             lm=T,
             pch="."
             )

cor_matrix <- cor(trevino[,c(36:69)], use = "pairwise.complete.obs",method="spearman")

# Convert to a data frame
cor_df <- as.data.frame(cor_matrix)

# Save as CSV
write.csv(cor_df, file = "corr.isompra.trevino.cpm.csv", row.names = TRUE)

cpm_columns <- grep("cpm", colnames(trevino), value = TRUE)
# Loop through cpm columns and apply the operation
for (col_name in cpm_columns) {
  # Extract the corresponding column name from 'merged_data'
  expr_column <- gsub("cpm", "expr", col_name)
  
  # Apply the operation to the column
    trevino[, paste0(col_name, ".expr") := (log2(trevino[[col_name]] / trevino$aav.cpm))]
}


pairs.panels(trevino[,c(70:86,88:103)], 
             method = "spearman", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE,
             lm=T,
             pch="."
             )

cor_matrix <- cor(trevino[,c(70:86,88:103)], use = "pairwise.complete.obs",method="spearman")

# Convert to a data frame
cor_df <- as.data.frame(cor_matrix)

# Save as CSV
write.csv(cor_df, file = "corr.isompra.trevino.expr.csv", row.names = TRUE)

data_for_mds.expr<-trevino[,c(36:69)]
mds_result.expr<-plotMDS(data_for_mds.expr)
```

## eMPRA replicate correlations

The eMPRA equivalent: filter to fully-observed barcodes, convert to CPM, compute
log2 expression relative to the AAV plasmid DNA, and summarize replicate
agreement with `pairs.panels`, correlation matrices, an MDS plot, and the
cross-MPRA correlation summary plots.

```{r correlations-empra, include=TRUE}
SIC_counts2<-SIC_countstrevino[rowSums(is.na(SIC_countstrevino[,2:25]))==0,]
tmp<-SIC_countstrevino[,c(1:11,13:14,16:19,23:26)]
tmp<-tmp[rowSums(is.na(tmp[,c(2:20)]))==0]


setDT(SIC_counts2)
setDT(tmp)
cols<-c("ms1","ms2","ms3","ms4","ms5","ms6","ms7","ms8","ms9","ms10","ms11","ms12","ms13","ms14","ms15","ms16","ms17","ms18","ms19","ms20","ms21","ms22","plasmid","aav")


for (col in cols) {
  SIC_counts2[, paste0(col, "_cpm") := (.SD[[col]]/sum(.SD[[col]])) * 1e6, .SDcols = col]
}

cols<-c("ms1","ms2","ms3","ms4","ms5","ms6","ms7","ms8","ms9","ms10","ms12","ms13","ms15","ms16","ms17","ms18","ms22","plasmid","aav")

for (col in cols) {
  tmp[, paste0(col, "_cpm") := (.SD[[col]]/sum(.SD[[col]])) * 1e6, .SDcols = col]
}


pairs.panels(SIC_counts2[,c(27:50)], 
             method = "spearman", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE,
             lm=T,
             pch="."
             )

pairs.panels(tmp[,c(22:40)], 
             method = "spearman", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE,
             lm=T,
             pch="."
             )

cor_matrix <- cor(tmp[,c(22:40)], use = "pairwise.complete.obs",method="spearman")

# Convert to a data frame
cor_df <- as.data.frame(cor_matrix)

# Save as CSV
write.csv(cor_df, file = "corr.empra.trevino.cpm.csv", row.names = TRUE)


#Calculating Raw Expression from Plasmid DNA

cols <- c("ms1_cpm","ms2_cpm","ms3_cpm","ms4_cpm","ms5_cpm","ms6_cpm","ms7_cpm","ms8_cpm","ms9_cpm","ms10_cpm","ms11_cpm","ms12_cpm","ms13_cpm","ms14_cpm","ms15_cpm","ms16_cpm","ms17_cpm","ms18_cpm","ms19_cpm","ms20_cpm","ms21_cpm","ms22_cpm")


for (col in cols) {
  col_log <- log2(SIC_counts2[, .SD[[col]]/SIC_counts2[, "aav_cpm"]])
  col_name <- paste0(col, ".expr.aav")
  SIC_counts2[, (col_name) := col_log]
}

cols <- c("ms1_cpm","ms2_cpm","ms3_cpm","ms4_cpm","ms5_cpm","ms6_cpm","ms7_cpm","ms8_cpm","ms9_cpm","ms10_cpm","ms12_cpm","ms13_cpm","ms15_cpm","ms16_cpm","ms17_cpm","ms18_cpm","ms22_cpm")


for (col in cols) {
  col_log <- log2(tmp[, .SD[[col]]/tmp[, "aav_cpm"]])
  col_name <- paste0(col, ".expr.aav")
  tmp[, (col_name) := col_log]
}

pairs.panels(SIC_counts2[,c(51:72)], 
             method = "spearman", # correlation method
             hist.col = "seagreen",
             density = TRUE,  # show density plots
             ellipses = TRUE,
             lm=T,
             pch="."# show correlation ellipses
             )

pairs.panels(tmp[,c(41:57)], 
             method = "spearman", # correlation method
             hist.col = "seagreen",
             density = TRUE,  # show density plots
             ellipses = TRUE,
             lm=T,
             pch="."# show correlation ellipses
             )

cor_matrix <- cor(tmp[,c(41:57)], use = "pairwise.complete.obs",method="spearman")

# Convert to a data frame
cor_df <- as.data.frame(cor_matrix)

# Save as CSV
write.csv(cor_df, file = "corr.empra.trevino.expr.csv", row.names = TRUE)


data_for_mds.expr<-tmp[,c(2:20)]
mds_result.expr<-plotMDS(data_for_mds.expr)

corr.values.trevino <- read_csv("corr.values.trevino.csv")

ggplot(corr.values.trevino,aes(x=factor(CorrType),y=Value,fill=factor(Sample)))+geom_jitter(aes(shape=factor(MPRA),color=factor(Sample),size=4),width = 0.1)

ggplot(corr.values.trevino,aes(x=factor(MPRA),y=Value,fill=factor(CorrType)))+geom_boxplot()+geom_dotplot(binaxis='y',stackdir='center',position=position_dodge(0.75),dotsize=0.5)+scale_fill_manual(values=c("#E1BE6A","#40B0A6"))

ggplot(corr.values.trevino,aes(x=factor(MPRA),y=Value,fill=factor(MPRA)))+geom_boxplot()+geom_dotplot(binaxis='y',stackdir='center',position=position_dodge(0.75),dotsize=0.5)+scale_fill_manual(values=c("#8E3636","#E1BE6A","#E1BE6A"))+facet_wrap(.~CorrType,scale="free_y")
```

# Reshaping to long 'UTR' format

These final chunks convert the wide count tables into the long format expected by
the downstream analysis notebook (one row per element x replicate, with raw
counts and CPM), attach the ref/alt allele annotation from the library metadata,
and write the `.rds`/`.csv` files.

## ISO-MPRA long format

```{r utr-format-isompra, include=FALSE}
# NOTE: meta.file3.trev.rds is read from an absolute path; update for your machine.
meta.file3.trev <- readRDS("C:/Users/dinse/Desktop/ISOMPRA/ps1.isompra.mgi/meta.file3.trev.rds")
metatab.reformatted<-meta.file3.trev[,c(6,10)]

pivotedsic<-trevino[,c(1:2,10,14:18,3:9,11:13,20,28,32:35,21:27,29:31,19)]

pivoted_long <- pivot_longer(pivotedsic, cols = c(2:35), names_to = "Replicate", values_to = "Value")
pivoted_long <- pivoted_long %>%
  dplyr::group_by(rsID) %>%
  dplyr::arrange(rsID)

pivoted_long <- pivoted_long %>%
  dplyr::mutate(Replicate = ifelse(Replicate %in% c("ms1","ms2","ms3","ms4","ms6","ms7","ms8","ms12","ms13","ms14","ms15","ms17","ms18","ms19","ms20","ms21","ms22","ms1.post","ms2.post","ms3.post","ms4.post","ms5.post","ms8.post","ms12.post","ms13.post","ms14.post","ms15.post","ms17.post","ms18.post","ms19.post","ms20.post","ms21.post","ms22.post"), match(Replicate, c("ms1","ms2","ms3","ms4","ms6","ms7","ms8","ms12","ms13","ms14","ms15","ms17","ms18","ms19","ms20","ms21","ms22","ms1.post","ms2.post","ms3.post","ms4.post","ms5.post","ms8.post","ms12.post","ms13.post","ms14.post","ms15.post","ms17.post","ms18.post","ms19.post","ms20.post","ms21.post","ms22.post")), Replicate))

pivoted_long <- pivoted_long %>%
  dplyr::mutate(
    Replicate = ifelse(Replicate %in% c("ms1","ms2","ms3","ms4","ms6","ms7","ms8","ms12","ms13","ms14","ms15","ms17","ms18","ms19","ms20","ms21","ms22","ms1.post","ms2.post","ms3.post","ms4.post","ms5.post","ms8.post","ms12.post","ms13.post","ms14.post","ms15.post","ms17.post","ms18.post","ms19.post","ms20.post","ms21.post","ms22.post"), match(Replicate, c("ms1","ms2","ms3","ms4","ms6","ms7","ms8","ms12","ms13","ms14","ms15","ms17","ms18","ms19","ms20","ms21","ms22","ms1.post","ms2.post","ms3.post","ms4.post","ms5.post","ms8.post","ms12.post","ms13.post","ms14.post","ms15.post","ms17.post","ms18.post","ms19.post","ms20.post","ms21.post","ms22.post")), Replicate),
    BiolRep = ifelse(Replicate %in% c("1", "2", "3","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18","19","20","21","22","23","24","25","26","27","28","29","30","31","32","33"), "rna.total", 
                      ifelse(Replicate == "aav", "dna.total", NA_character_))
  )

pivotedsic2<-trevino[,c(1,36,44,48:52,37:43,45:47,54,62,66:69,55:61,63:65,53)]

pivoted_long2 <- pivot_longer(pivotedsic2, cols = c(2:35), names_to = "Replicate", values_to = "Value")
pivoted_long2 <- pivoted_long2 %>%
  dplyr::group_by(rsID) %>%
  dplyr::arrange(rsID)

pivoted_long2 <- pivoted_long2 %>%
  mutate(Replicate = ifelse(Replicate %in% c("ms1.cpm","ms2.cpm","ms3.cpm","ms4.cpm","ms6.cpm","ms7.cpm","ms8.cpm","ms12.cpm","ms13.cpm","ms14.cpm","ms15.cpm","ms17.cpm","ms18.cpm","ms19.cpm","ms20.cpm","ms21.cpm","ms22.cpm","ms1.post.cpm","ms2.post.cpm","ms3.post.cpm","ms4.post.cpm","ms5.post.cpm","ms8.post.cpm","ms12.post.cpm","ms13.post.cpm","ms14.post.cpm","ms15.post.cpm","ms17.post.cpm","ms18.post.cpm","ms19.post.cpm","ms20.post.cpm","ms21.post.cpm","ms22.post.cpm"
), match(Replicate, c("ms1.cpm","ms2.cpm","ms3.cpm","ms4.cpm","ms6.cpm","ms7.cpm","ms8.cpm","ms12.cpm","ms13.cpm","ms14.cpm","ms15.cpm","ms17.cpm","ms18.cpm","ms19.cpm","ms20.cpm","ms21.cpm","ms22.cpm","ms1.post.cpm","ms2.post.cpm","ms3.post.cpm","ms4.post.cpm","ms5.post.cpm","ms8.post.cpm","ms12.post.cpm","ms13.post.cpm","ms14.post.cpm","ms15.post.cpm","ms17.post.cpm","ms18.post.cpm","ms19.post.cpm","ms20.post.cpm","ms21.post.cpm","ms22.post.cpm"
)), Replicate))

pivoted_long2 <- pivoted_long2 %>%
  mutate(
    Replicate = ifelse(Replicate %in% c("ms1.cpm","ms2.cpm","ms3.cpm","ms4.cpm","ms6.cpm","ms7.cpm","ms8.cpm","ms12.cpm","ms13.cpm","ms14.cpm","ms15.cpm","ms17.cpm","ms18.cpm","ms19.cpm","ms20.cpm","ms21.cpm","ms22.cpm","ms1.post.cpm","ms2.post.cpm","ms3.post.cpm","ms4.post.cpm","ms5.post.cpm","ms8.post.cpm","ms12.post.cpm","ms13.post.cpm","ms14.post.cpm","ms15.post.cpm","ms17.post.cpm","ms18.post.cpm","ms19.post.cpm","ms20.post.cpm","ms21.post.cpm","ms22.post.cpm"
), match(Replicate, c("ms1.cpm","ms2.cpm","ms3.cpm","ms4.cpm","ms6.cpm","ms7.cpm","ms8.cpm","ms12.cpm","ms13.cpm","ms14.cpm","ms15.cpm","ms17.cpm","ms18.cpm","ms19.cpm","ms20.cpm","ms21.cpm","ms22.cpm","ms1.post.cpm","ms2.post.cpm","ms3.post.cpm","ms4.post.cpm","ms5.post.cpm","ms8.post.cpm","ms12.post.cpm","ms13.post.cpm","ms14.post.cpm","ms15.post.cpm","ms17.post.cpm","ms18.post.cpm","ms19.post.cpm","ms20.post.cpm","ms21.post.cpm","ms22.post.cpm"
)), Replicate),
    BiolRep = ifelse(Replicate %in% c("1", "2", "3","4","5","6","7","8","9","10","11","12","13","14","15","16","17","18","19","20","21","22","23","24","25","26","27","28","29","30","31","32","33"), "rna.total", 
                      ifelse(Replicate == "aav.cpm", "dna.total", NA_character_))
  )

pivoted_long2$Replicate[pivoted_long2$Replicate == "aav.cpm"] <- "aav"
merged_df <- inner_join(pivoted_long,pivoted_long2,by=c("rsID","Replicate"))
merged_df<-merged_df[,c(1:3,5,4)]
names(merged_df)<-c("Element","BiolRep","Counts","CPM","Fraction")
metatab.reformatted$barcode_allele <- gsub("_", "=", metatab.reformatted$barcode_allele)
names(metatab.reformatted)<-c("ref,alt","Element")
merged_df$Element <- gsub("scrambled*", "scrambled=", merged_df$Element)
merged_df<-right_join(metatab.reformatted,merged_df,by="Element")
merged_df<-distinct(merged_df)
merged_df<-merged_df[order(merged_df$Element),]

merged_df$TechRep<-1
merged_df<-mutate(merged_df,BiolRep = ifelse(BiolRep == "aav", 1, BiolRep))
merged_df<-mutate(merged_df,`ref,alt` = ifelse(str_detect(Element, "scrambled="), "shuf", `ref,alt`))
merged_df$Experiment<-"PS1.invivo.ISOMPRA.10.8.2024"
merged_df<-dplyr::rename(merged_df,Allele = `ref,alt`)

merged_df<-merged_df[,c(8,2,1,3,7,6,4:5)]
write.csv(merged_df,"invivo.ps1.utr.format.10.8.24.csv")
write_rds(merged_df,"invivo.ps1.utr.format.10.8.24.rds")
```

## eMPRA long format

```{r utr-format-empra, include=FALSE}
# NOTE: meta.file3.trev.rds is read from an absolute path; update for your machine.
meta.file3.trev <- readRDS("C:/Users/dinse/Desktop/ISOMPRA/ps1.isompra.mgi/meta.file3.trev.rds")
metatab.reformatted<-meta.file3.trev[,c(6,10)]

pivotedsic<-tmp[,c(1:18,20,21)]

pivoted_long <- pivot_longer(pivotedsic, cols = c(2:19), names_to = "Replicate", values_to = "Value")
pivoted_long <- pivoted_long %>%
  dplyr::group_by(rsID) %>%
  dplyr::arrange(rsID, revcomp_barcode) %>%
  dplyr::mutate(BC = pmin(dense_rank(revcomp_barcode), 20))

pivoted_long <- pivoted_long %>%
  dplyr::mutate(Replicate = ifelse(Replicate %in% c("ms1","ms2","ms3","ms4","ms5","ms6","ms7","ms8","ms9","ms10","ms12","ms13","ms15","ms16","ms17","ms18","ms22"), match(Replicate, c("ms1","ms2","ms3","ms4","ms5","ms6","ms7","ms8","ms9","ms10","ms12","ms13","ms15","ms16","ms17","ms18","ms22")), Replicate))

pivoted_long <- pivoted_long %>%
  dplyr::mutate(
    Replicate = ifelse(Replicate %in% c("ms1","ms2","ms3","ms4","ms5","ms6","ms7","ms8","ms9","ms10","ms12","ms13","ms15","ms16","ms17","ms18","ms22"), match(Replicate, c("ms1","ms2","ms3","ms4","ms5","ms6","ms7","ms8","ms9","ms10","ms12","ms13","ms15","ms16","ms17","ms18","ms22")), Replicate),
    BiolRep = ifelse(Replicate %in% c("1", "2", "3","4","5","6","7","8","9","10","11","12","13","14","15","16","17"), "rna.total", 
                      ifelse(Replicate == "aav", "dna.total", NA_character_))
  )

pivotedsic2<-tmp[,c(1,22:38,40,21)]

pivoted_long2 <- pivot_longer(pivotedsic2, cols = c(2:19), names_to = "Replicate", values_to = "Value")
pivoted_long2 <- pivoted_long2 %>%
  dplyr::group_by(rsID) %>%
  dplyr::arrange(rsID, revcomp_barcode) %>%
  dplyr::mutate(BC = pmin(dense_rank(revcomp_barcode), 20))

pivoted_long2 <- pivoted_long2 %>%
  mutate(Replicate = ifelse(Replicate %in% c("ms1_cpm","ms2_cpm","ms3_cpm","ms4_cpm","ms5_cpm","ms6_cpm","ms7_cpm","ms8_cpm","ms9_cpm","ms10_cpm","ms12_cpm","ms13_cpm","ms15_cpm","ms16_cpm","ms17_cpm","ms18_cpm","ms22_cpm"), match(Replicate, c("ms1_cpm","ms2_cpm","ms3_cpm","ms4_cpm","ms5_cpm","ms6_cpm","ms7_cpm","ms8_cpm","ms9_cpm","ms10_cpm","ms12_cpm","ms13_cpm","ms15_cpm","ms16_cpm","ms17_cpm","ms18_cpm","ms22_cpm")), Replicate))

pivoted_long2 <- pivoted_long2 %>%
  mutate(
    Replicate = ifelse(Replicate %in% c("ms1_cpm","ms2_cpm","ms3_cpm","ms4_cpm","ms5_cpm","ms6_cpm","ms7_cpm","ms8_cpm","ms9_cpm","ms10_cpm","ms12_cpm","ms13_cpm","ms15_cpm","ms16_cpm","ms17_cpm","ms18_cpm","ms22_cpm"
), match(Replicate, c("ms1_cpm","ms2_cpm","ms3_cpm","ms4_cpm","ms5_cpm","ms6_cpm","ms7_cpm","ms8_cpm","ms9_cpm","ms10_cpm","ms12_cpm","ms13_cpm","ms15_cpm","ms16_cpm","ms17_cpm","ms18_cpm","ms22_cpm")), Replicate),
    BiolRep = ifelse(Replicate %in% c("1", "2", "3","4","5","6","7","8","9","10","11","12","13","14","15","16","17"), "rna.total", 
                      ifelse(Replicate == "aav_cpm", "dna.total", NA_character_))
  )

pivoted_long2$Replicate[pivoted_long2$Replicate == "aav_cpm"] <- "aav"
merged_df <- inner_join(pivoted_long,pivoted_long2,by=c("revcomp_barcode","Replicate"))
merged_df<-merged_df[,c(1:4,8,5:6)]
names(merged_df)<-c("revcomp_barcode","Element","BiolRep","Counts","CPM","BC","Fraction")
metatab.reformatted$barcode_allele <- gsub("_", "=", metatab.reformatted$barcode_allele)
names(metatab.reformatted)<-c("ref,alt","Element")
merged_df$Element <- gsub("scrambled*", "scrambled=", merged_df$Element)
merged_df$Element <- gsub("Basal*", "Basal=", merged_df$Element)
merged_df<-right_join(metatab.reformatted,merged_df,by="Element")
merged_df<-distinct(merged_df)
merged_df<-merged_df[order(merged_df$Element),]

merged_df$TechRep<-1
merged_df<-mutate(merged_df,BiolRep = ifelse(BiolRep == "aav", 1, BiolRep))
merged_df<-mutate(merged_df,`ref,alt` = ifelse(str_detect(Element, "scrambled="), "shuf", `ref,alt`))
merged_df<-mutate(merged_df,`ref,alt` = ifelse(str_detect(Element, "Basal="), "ref", `ref,alt`))
merged_df$Experiment<-"PS1.HEK.eMPRA.9.20.24"
merged_df<-dplyr::rename(merged_df,Allele = `ref,alt`)

merged_df<-merged_df[,c(10,3,2,1,7,4,9,8,5:6)]
write.csv(merged_df,"empratrev.utr.format.9.20.24.csv")
write_rds(merged_df,"empratrev.utr.format.9.20.24.rds")


rm(col_log,cor_df,cor_matrix,corr.values.trevino,corr.values.trevino.sanity,data_for_mds.expr,mappedcounts,mds_result.expr,merged_df,pivoted_long,pivoted_long2,pivotedsic,pivotedsic2,totalreads,totalreads.empra,totalreadsout,totalreadsout.trevino,totalreadsout.trevino.isompra)
```
