---
title: "Untitled"
author: "Anthony Fischer"
date: "2026-04-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load in tools
```{r include=FALSE}
source("../../../code/mpra_tools/mpra_tools.R")
library(fplot)
```

##Read in Data
```{r Reading_In_Data, include=FALSE}
d <- fread("../../../data/formatted/ActD_Trials_08Apr2026/formatted.txt")
#Separate allele names
d <- d %>%
  separate_wider_regex(
    cols = Element,
    patterns = c(
      Element = ".*",        # Match as much as possible (greedy)
      "_",                   # Match the last underscore
      Allele = "[^_]+"       # Match everything remaining that isn't an underscore
    )
  )
d <- as.data.table(d)

#Grab the DNA info
dna <- fread("../../../data/formatted/Library1and2BCLess_ISOMPRA_6-30-2022/formatted_Reads_compiled_06-30-2022.tsv") %>% filter(Fraction=="DNA")
dna <- dna %>% group_by(Element,Allele) %>% summarize(CPM=mean(CPM)) %>% as.data.table()
```

##Calculate Expression
```{r expression_calculation}
#Put the DNA info in the table, adjoining to correct element
d <- as.data.table(left_join(d,dna,by=c("Element","Allele")))
#Calculate expression
d[, Expression := log2(CPM.x/CPM.y)]
#Get rid of non-finite values
d <- filter(d,is.finite(Expression))
```

##Calculate Library Coverage
```{r Coverage}
#Total coverage for each library
# 1. Create a new summary table
total_counts <- d[, .(TotalCounts = sum(Counts)), 
                      by = .(Fraction, BiolRep, Splicing)]
total_counts[,Coverage := TotalCounts/8690]
``` 

##Examine CPM and Expression Correlations
```{r Correlation_Plots_NewData}
#Let's look at correlation plots
CPM_corrs <- d %>% 
  pivot_wider(id_cols = c(Element,Allele),
              names_from = c(Fraction,Splicing,BiolRep),
              values_from = CPM.x) %>%
  select(3:26) %>% cor(use = "pairwise.complete.obs") %>%
  as.data.table(,keep.rownames = TRUE) %>% pivot_longer(2:25) %>% filter(rn<name)
Exp_corrs <- d %>% 
  pivot_wider(id_cols = c(Element,Allele),
              names_from = c(Fraction,Splicing,BiolRep),
              values_from = Expression) %>%
  select(3:26) %>% cor(use = "pairwise.complete.obs") %>% 
  as.data.table(,keep.rownames = TRUE) %>% pivot_longer(2:25) %>% filter(rn<name)
ggplot(CPM_corrs,aes(x=rn,y=name,fill=value))+
  geom_tile()+
  geom_text(aes(label = round(value,2)))+
  scale_fill_gradient2(low = "blue", 
                       high = "red", 
                       mid = "white", 
                       midpoint = 0, 
                       limit = c(-1, 1), 
                       name = "Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_fixed()+
  labs(title="CPM Correlation")
ggplot(Exp_corrs,aes(x=rn,y=name,fill=value))+
  geom_tile()+
  geom_text(aes(label = round(value,2)))+
  scale_fill_gradient2(low = "blue", 
                       high = "red", 
                       mid = "white", 
                       midpoint = 0, 
                       limit = c(-1, 1), 
                       name = "Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_fixed()+
  labs(title="Expression Correlation")

#And look at the old data
Exp_corrs <- filter(d_old,!grepl("TRAP",Comment)) %>% 
  pivot_wider(id_cols = c(Element,Allele),
              names_from = c(Comment,BiolRep,TechRep),
              values_from = Expression) %>%
  select(3:26) %>% cor(use = "pairwise.complete.obs") %>% 
  as.data.table(,keep.rownames = TRUE) %>% pivot_longer(2:25) %>% filter(rn<name)
ggplot(Exp_corrs,aes(x=rn,y=name,fill=value))+
  geom_tile()+
  geom_text(aes(label = round(value,2)))+
  scale_fill_gradient2(low = "blue", 
                       high = "red", 
                       mid = "white", 
                       midpoint = 0, 
                       limit = c(-1, 1), 
                       name = "Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_fixed()+
  labs(title="Exp Correlation")


################################################################################
#CPM PCA
# 1. Transform: the data must be transformed first. Raw CPMs are often
# "heteroscedastic" (the variance changes with the mean), which can trick PCA
# into only looking at your highest-expressed elements. You learn something new
#every day...
pca_input <- d %>% 
  pivot_wider(id_cols = c(Element,Allele),
              names_from = c(Fraction,Splicing,BiolRep),
              values_from = CPM.x) %>% tidyr::unite(.,Element,c("Element","Allele")) %>%
  column_to_rownames("Element") %>%
  select(1:24) %>%
  mutate(across(everything(), ~log2(.x + 1))) %>%
  t() # PCA expects samples in rows, elements in columns
#Fill NAs with 0
pca_input[is.na(pca_input)] <- 0

# 2. Run PCA
pca_res <- prcomp(pca_input, scale. = TRUE)

# 3. Visualize
as.data.frame(pca_res$x) %>%
  rownames_to_column("Library") %>%
  separate(Library, into = c("Fraction", "Splicing", "BiolRep"), sep = "_") %>%
  ggplot(aes(x = PC1, y = PC2, fill = interaction(Fraction,Splicing))) +
  geom_point(size = 4,shape=21) +
  geom_polygon(stat = "ellipse", level=0.68, alpha=0.2, color="black")+
  geom_text(aes(label = BiolRep), vjust = -1) + # Helps identify which rep is the outlier
  labs(fill="Library",title="PCA of Library CPMs")+
  theme(plot.title = element_text(hjust=0.5))
  
  
################################################################################
#Expression PCA
# 1. Transform: Expression is already in log2 space
pca_input <- d %>% 
  pivot_wider(id_cols = c(Element,Allele),
              names_from = c(Fraction,Splicing,BiolRep),
              values_from = Expression) %>% tidyr::unite(.,Element,c("Element","Allele")) %>%
  column_to_rownames("Element") %>%
  select(1:24) %>%
  t() # PCA expects samples in rows, elements in columns
#Fill NAs with 0
pca_input[is.na(pca_input)] <- 0

# 2. Run PCA
pca_res <- prcomp(pca_input, scale. = TRUE)

# 3. Visualize
as.data.frame(pca_res$x) %>%
  rownames_to_column("Library") %>%
  separate(Library, into = c("Fraction", "Splicing", "BiolRep"), sep = "_") %>%
  ggplot(aes(x = PC1, y = PC2, fill = interaction(Fraction,Splicing))) +
  geom_point(size = 4,shape=21) +
  geom_polygon(stat = "ellipse", level=0.68, alpha=0.2,color="black")+
  geom_text(aes(label = BiolRep), vjust = -1) + # Helps identify which rep is the outlier
  labs(fill="Library",title="PCA of Library Expression")+
  theme(plot.title = element_text(hjust=0.5))

# It appears t11 biolrep 2 and t0 pre biolrep 3 are outliers.  Exclude them from the analysis.
```

## Reference vs. Shuffled
### Generate Ref vs. Shuf Welch's T-Test Function
```{r refshuf_function}
# We are going to repeat the refshuf analysis for this dataset


welch_shuftest <- function(data_table,element,allele,null_values){
  d <- filter(data_table,Allele==allele,Element==element)
  res <- t.test(d$Expression,null_values, var.equal=FALSE)
  dif <- res$estimate[1] - res$estimate[2]
  pval <- res$p.value
  return(c(element,dif,pval))
}
```

### Generate list of elements to test
```{r testable}
#Remove offending biolreps from list based on PCA and correlation heatmap
d_new <- filter(d,
            !(Splicing=="pre" & BiolRep==3),
            !(Fraction=="11hr" & BiolRep==2))
#Get new list of elements to test, knowing that 0hr Pre 3 and 11 hr post 2 are excluded
# Generate lists of elements to test by counting number of times the element appears
# (had an expression value calculated for how many biolreps)
pre <- as.data.table(table(filter(d_new,Fraction=="0hr",Splicing=="pre", Counts>49)[,c("Element","Allele")]))
pre <- pivot_wider(pre,1,names_from = "Allele",values_from = "N")
pre <- filter(pre,alt>2,ref>2)
post <- as.data.table(table(filter(d_new,Fraction=="0hr",Splicing=="post", Counts>49)[,c("Element","Allele")]))
post <- pivot_wider(post,1,names_from = "Allele",values_from = "N")
post <- filter(post,alt>2,ref>2)
five <- as.data.table(table(filter(d_new,Fraction=="5hr",Splicing=="post", Counts>49)[,c("Element","Allele")]))
five <- pivot_wider(five,1,names_from = "Allele",values_from = "N")
five <- filter(five,alt>2,ref>2)
eleven <- as.data.table(table(filter(d_new,Fraction=="11hr",Splicing=="post", Counts>49)[,c("Element","Allele")]))
eleven <- pivot_wider(eleven,1,names_from = "Allele",values_from = "N")
eleven <- filter(eleven,alt>2,ref>2)
library(purrr)
library(dplyr)
elements <- list(pre$Element,post$Element,five$Element,eleven$Element) %>% purrr::reduce(intersect)
```

### Pre-Splicing Ref vs. Shuf
```{r pre_refshuf}
# Pre-Splicing
nulls <- as.data.table(filter(d,Allele=="shuf",Fraction=="0hr",Splicing=="pre") %>% group_by(Element) %>%
    summarize(Expression=mean(Expression)))
intron_shuf <- lapply(elements, function(element) {
  welch_shuftest(data_table = filter(d,Fraction=="0hr",Splicing=="pre"), element = element, allele = "ref",null_values=nulls$Expression)
})
intron_shuf <- as.data.table(do.call(rbind,intron_shuf))
setnames(intron_shuf,,c("Element","logFC","p.val"))
intron_shuf$fdr <- p.adjust(intron_shuf$p.val,method = "fdr")
intron_shuf <- full_join(intron_shuf,fread("../../../code/mpra_tools/fullannotationtable.txt"))[,!c("Pos","Ref","Alt","Type","Pheno","library_seq_alt","library_seq_shuf")]
fwrite(intron_shuf,"Pre-Splicing_RefvsShuf_Welchs.tsv")


### Post-Splicing Ref vs. Shuf
```{r post_refshuf}
# Post-Splicing
nulls <- as.data.table(filter(d,Allele=="shuf",Fraction=="0hr",Splicing=="post") %>% group_by(Element) %>%
    summarize(Expression=mean(Expression)))
exon_shuf <- lapply(elements, function(element) {
  welch_shuftest(data_table = filter(d,Fraction=="0hr",Splicing=="post"), element = element, allele = "ref",null_values=nulls$Expression)
})
exon_shuf <- as.data.table(do.call(rbind,exon_shuf))
setnames(exon_shuf,,c("Element","logFC","p.val"))
exon_shuf$fdr <- p.adjust(exon_shuf$p.val,method = "fdr")
exon_shuf <- full_join(exon_shuf,fread("../../../code/mpra_tools/fullannotationtable.txt"))[,!c("Pos","Ref","Alt","Type","Pheno","library_seq_alt","library_seq_shuf")]
fwrite(exon_shuf,"Post-Splicing_RefvsShuf_Welchs.tsv")
```

### Stability Ref vs. Shuf
```{r stability_refshuf}
# Stability
# Join tables
stab_table <- as.data.table(inner_join(filter(d,Fraction=="0hr",Splicing=="pre"),
                         filter(d,Fraction=="0hr",Splicing=="post"),
                         by=c("Element","Allele","BiolRep","Fraction")))
stab_table[,Expression := Expression.y-Expression.x]
nulls <- as.data.table(filter(stab_table,Allele=="shuf") %>% group_by(Element) %>%
    summarize(Expression=mean(Expression)))
stab_shuf <- lapply(elements, function(element) {
  welch_shuftest(data_table = stab_table, element = element, allele = "ref",null_values=nulls$Expression)
})
stab_shuf <- as.data.table(do.call(rbind,stab_shuf))
setnames(stab_shuf,,c("Element","logFC","p.val"))
stab_shuf$fdr <- p.adjust(stab_shuf$p.val,method = "fdr")
stab_shuf <- full_join(stab_shuf,fread("../../../code/mpra_tools/fullannotationtable.txt"))[,!c("Pos","Ref","Alt","Type","Pheno","library_seq_alt","library_seq_shuf")]
fwrite(stab_shuf,"Stability_RefvsShuf_Welchs.tsv")
```

## Reference vs. Variant
### Linear Model on Pre- and Post-Splicing Ref vs Alt
```{r pre_refvsalt}
#Let's look at the post-only and pre-only ref/alt allele effects and correlate them
pre <- as.data.table(table(filter(d_new,Fraction=="0hr",Splicing=="pre", Counts>49)[,c("Element","Allele")]))
pre <- pivot_wider(pre,1,names_from = "Allele",values_from = "N")
pre <- filter(pre,alt>2,ref>2)
post <- as.data.table(table(filter(d_new,Fraction=="0hr",Splicing=="post", Counts>49)[,c("Element","Allele")]))
post <- pivot_wider(post,1,names_from = "Allele",values_from = "N")
post <- filter(post,alt>2,ref>2)
elements <- list(pre$Element,post$Element) %>% purrr::reduce(intersect)

post_res <- mpratools.lm(filter(d_new,Fraction=="0hr",Splicing=="post",Element %in% elements),
                             Expression ~ Allele)
pre_res <- mpratools.lm(filter(d_new,Fraction=="0hr",Splicing=="pre", Element %in% elements),
                             Expression ~ Allele)
post_res <- left_join(post_res,fread("../../../code/mpra_tools/fullannotationtable.txt"))
fwrite(post_res,"PostSplicing_RefAlt_LM.txt",sep="\t")
pre_res <- left_join(pre_res,fread("../../../code/mpra_tools/fullannotationtable.txt"))
fwrite(pre_res,"PreSplicing_RefAlt_LM.txt",sep="\t")
```

### Linear Interaction Models on Pre/Post t0
```{r run_LMs, include=FALSE}
#Remove offending biolreps from list based on PCA and correlation heatmap
d_new <- filter(d,
            !(Splicing=="pre" & BiolRep==3),
            !(Fraction=="11hr" & BiolRep==2))
#Get new list of elements to test, knowing that 0hr Pre 3 and 11 hr post 2 are excluded
# Generate lists of elements to test by counting number of times the element appears
# (had an expression value calculated for how many biolreps)
pre <- as.data.table(table(filter(d_new,Fraction=="0hr",Splicing=="pre", Counts>49)[,c("Element","Allele")]))
pre <- pivot_wider(pre,1,names_from = "Allele",values_from = "N")
pre <- filter(pre,alt>2,ref>2)
post <- as.data.table(table(filter(d_new,Fraction=="0hr",Splicing=="post", Counts>49)[,c("Element","Allele")]))
post <- pivot_wider(post,1,names_from = "Allele",values_from = "N")
post <- filter(post,alt>2,ref>2)
five <- as.data.table(table(filter(d_new,Fraction=="5hr",Splicing=="post", Counts>49)[,c("Element","Allele")]))
five <- pivot_wider(five,1,names_from = "Allele",values_from = "N")
five <- filter(five,alt>2,ref>2)
eleven <- as.data.table(table(filter(d_new,Fraction=="11hr",Splicing=="post", Counts>49)[,c("Element","Allele")]))
eleven <- pivot_wider(eleven,1,names_from = "Allele",values_from = "N")
eleven <- filter(eleven,alt>2,ref>2)
library(purrr)
library(dplyr)
elements <- list(pre$Element,post$Element,five$Element,eleven$Element) %>% purrr::reduce(intersect)
#That's a lot, good.

stability <- mpratools.lm.int( 
  filter(d_new,Fraction=="0hr",Element %in% elements),
  Expression ~ Allele * Splicing
)

stability <- left_join(stability,fread("../../../code/mpra_tools/fullannotationtable.txt"))
fwrite(stability,"Stability_RefAlt_LM.txt",sep="\t")
```

## ActinomycinD Allelic Ratios
### Allelic Ratio vs Time Modeling For t0 to t5
```{r actd_allelic_ratio}
# The data are super noisy.  Try allelic ratio at each time point like this:
# 1)Calculate the Allelic Ratio within each time point - t0_rt = log2(Variant_t0/reference_t0)
#   and t5 = log2(Variant_t5/Reference_t5)
# 2)Calculate the change in ratio delta_rt = t5_rt - t0_rt
# By using the Reference allele as the internal control for each variant within 
# the same tube, you bypass all the total-pool-shrinkage issues.
# Do this on counts only, NOT CPM, across biolreps.  Then the LM will be like
# AllelicRatio ~ Time, or even AllelicRatio ~ Time + (1 | BiolRep-Time)

Actd_Ratio <- pivot_wider(
  filter(d,
         Fraction %in% c("0hr","5hr"),
         Splicing=="post",!Allele %in% c("shuf","shuf2")),
  id_cols = c("Element","BiolRep"),
  names_from = c("Fraction","Allele"),
  values_from = "Counts")

# Eliminate rows that had a count less than 49 anywhere
Actd_Ratio <- as.data.table(Actd_Ratio %>%
  filter(if_all(3:6, ~ .x > 49)))

#Get a list of elements to test
elements <- filter(
  as.data.table(table(Actd_Ratio$Element)),
  N>2)$V1

Actd_Ratio <- filter(Actd_Ratio,Element %in% elements)

# 3769, good

Actd_Ratio[, Log2Ratio_T0 := log2(`0hr_alt`/`0hr_ref`)]
Actd_Ratio[, Log2Ratio_T5 := log2(`5hr_alt`/`5hr_ref`)]

#Now put it in long format for testing
Actd_Ratio <- as.data.table(pivot_longer(
  Actd_Ratio[,c(1,2,7,8)],
  cols = 3:4,
  names_to = "Time",
  values_to = "Log2Ratio"))

#Make the Time a Factor
Actd_Ratio[,Time := as.factor(Time)]

# Run the model for every element
Actd_Ratio_Results <- Actd_Ratio %>%
  group_by(Element) %>%
  do(tidy(lm(Log2Ratio ~ Time, data = .))) %>%
  ungroup() %>%
  # Filter for the Time coefficient (the difference of means)
  filter(term == "TimeLog2Ratio_T5") %>%
  # Clean up the output
  select(Element, 
         Delta_Log2Ratio = estimate, 
         p_value = p.value, 
         std_error = std.error) %>%
  # Add Adjusted P-values for multiple testing
  mutate(fdr = p.adjust(p_value, method = "BH"))

# Sort by the most significant stability changes
Actd_Ratio_Results <- Actd_Ratio_Results %>% arrange(p_value)

Actd_Ratio_Results <- left_join(Actd_Ratio_Results,fread("../../../code/mpra_tools/fullannotationtable.txt"))
fwrite(Actd_Ratio_Results,"ActD_Zero-to-Five_RefvsAlt_LM.tsv",sep="\t")
```

### Allelic Ratio vs Time Modeling For t0 to t11
```{r t11_allelic_ratio}

Actd_Ratio <- as.data.table(pivot_wider(
  filter(d,Fraction %in% c("0hr","11hr"),
         Splicing=="post",
         !Allele %in% c("shuf","shuf2"),
         !(Fraction=="11hr" & BiolRep==2)),
  id_cols = c("Element","BiolRep"),
  names_from = c("Fraction","Allele"),
  values_from = "Counts"))

# Eliminate rows that had a count less than 49 anywhere
Actd_Ratio <- as.data.table(Actd_Ratio %>%
  filter(if_all(3:6, ~ .x > 49)))

#Get a list of elements to test
elements <- filter(
  as.data.table(table(Actd_Ratio$Element)),
  N>2)$V1

# 3767, still really good

Actd_Ratio <- filter(Actd_Ratio,Element %in% elements)

Actd_Ratio[, Log2Ratio_T0 := log2(`0hr_alt`/`0hr_ref`)]
Actd_Ratio[, Log2Ratio_T11 := log2(`11hr_alt`/`11hr_ref`)]

#Now put it in long format for testing
Actd_Ratio <- as.data.table(pivot_longer(
  Actd_Ratio[,c(1,2,7:8)],
  cols = 3:4,
  names_to = "Time",
  values_to = "Log2Ratio"))

#Make the Time a Factor
Actd_Ratio[,Time := as.factor(Time)]

# Run the model for every element
Actd_Ratio_Results_0to11 <- Actd_Ratio %>%
  group_by(Element) %>%
  do(tidy(lm(Log2Ratio ~ Time, data = .))) %>%
  ungroup() %>%
  # Filter for the Time coefficient (the difference of means)
  filter(term == "TimeLog2Ratio_T11") %>%
  # Clean up the output
  select(Element, 
         Delta_Log2Ratio = estimate, 
         p_value = p.value, 
         std_error = std.error) %>%
  # Add Adjusted P-values for multiple testing
  mutate(fdr = p.adjust(p_value, method = "BH"))

# Sort by the most significant stability changes
Actd_Ratio_Results_0to11 <- Actd_Ratio_Results_0to11 %>% arrange(p_value)
Actd_Ratio_Results_0to11 <- left_join(Actd_Ratio_Results_0to11,fread("../../../code/mpra_tools/fullannotationtable.txt"))
fwrite(Actd_Ratio_Results_0to11,"ActD_Zero-to-Eleven_RefvsAlt_LM.tsv",sep="\t")
```

##Plotting ISOMPRA Stability Log2FC Against Allelic Ratio for Significant Variants
```{r Plots}

Actd_Ratio_Results <- fread("./ActD_Zero-to-Five_RefvsAlt_LM.tsv")
Actd_Ratio_Results_0to11 <- fread("./ActD_Zero-to-Eleven_RefvsAlt_LM.tsv")
stability <- fread("./Stability_RefAlt_LM.txt")

full <- list(
  Actd_Ratio_Results[, c(1, 2, 3, 5, 17)],
  Actd_Ratio_Results_0to11[, c(1, 2, 3, 5, 17)],
  stability[, c(1:12, 25)]
) %>% 
  purrr::reduce(~ left_join(.x, .y, by = c("Element", "Symbol")))

#Filter for variants that were nominally significant in both t0:t5 and t0:t11
elements <- list(
           filter(full,
                  abs(p_value.x)<0.05,
                  abs(p_value.y)<0.05)$Element
         ) %>% purrr::reduce(intersect)

truedecay <- filter(full, Element %in% elements)

zerofivestat <- cor.test(
  truedecay$Delta_Log2Ratio.x,
  -truedecay$logFC_interaction,
  method = "spearman")

zeroelevenstat <- cor.test(
  truedecay$Delta_Log2Ratio.y,
  -truedecay$logFC_interaction,
  method = "spearman")

xlims <- c(min(truedecay$logFC_interaction),max(truedecay$logFC_interaction))
ylims <- c(
  min(
    min(truedecay$Delta_Log2Ratio.x),min(truedecay$Delta_Log2Ratio.y)),
  max(
    max(truedecay$Delta_Log2Ratio.x),max(truedecay$Delta_Log2Ratio.y))
  )

p1 <- ggplot(truedecay, aes(x = -logFC_interaction, y = Delta_Log2Ratio.x)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  labs(x="ISOMPRA log2FC",
       y="ActD Allelic Ratio log2FC 0hr to 5hr")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_vline(xintercept = 0)+
  geom_hline(yintercept = 0)+
  geom_label_repel(aes(label=Symbol))+
  annotate("label", x = -1, y = 1.5,
           label = paste0("rho == ", round(zerofivestat$estimate, 2), 
                          " ~~~ p == ", round(zerofivestat$p.value, 4)),
           parse = TRUE)+
  xlim(-2,2)+
  ylim(-2,2)
print(p1)

p2 <- ggplot(truedecay, aes(x = -logFC_interaction, y = Delta_Log2Ratio.y)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", color = "blue", se = FALSE) +
  labs(x="ISOMPRA log2FC",
       y="ActD Allelic Ratio log2FC 0hr to 11hr")+
  theme(plot.title=element_text(hjust=0.5))+
  geom_vline(xintercept = 0)+
  geom_hline(yintercept = 0)+
  geom_label_repel(aes(label=Symbol))+
  annotate("label", x = -1, y = 1.5,
           label = paste0("rho == ", round(zeroelevenstat$estimate, 2), 
                          " ~~~ p == ", round(zeroelevenstat$p.value, 4)),
           parse = TRUE)+
  xlim(-2,2)+
  ylim(-2,2)
print(p2)

#Place this is Figure S7 of the figure_making folder
dir.create("../../../figure_making/Figures/")
dir.create("../../../figure_making/Figures/Figure_S7")
export_graph_start("../../../figure_making/Figures/Figure_S7/ActD logFC vs ISOMPRA logFC.pdf",sideways = TRUE,width = 2,height=1.5)
combined_plot <- p1+p2
combined_plot +
  plot_annotation(
    title = 'ActinomycinD Log2(Allelic Ratio) Analysis',
    subtitle = 'Comparing 0hr-5hr vs 0hr-11hr logFCs against ISOMPRA logFC',
    theme = theme(plot.title = element_text(size = 18, hjust = 0.5, face = "bold"),
                  plot.subtitle = element_text(size = 18, hjust = 0.5))
  )
export_graph_end()



```

##Individual BoxPlots of Selected Variants
```{r boxplots}
# Plot Individual Boxplots for variants with spike-in counts added
# Spike-in Counts were not necessary for the allelic ratio/pre-post analyses

d <- fread("../../../data/formatted/ActD_Trials_08Apr2026/formatted.txt")

d <- d %>%
  separate_wider_regex(
    cols = Element,
    patterns = c(
      Element = ".*",        # Match as much as possible (greedy)
      "_",                   # Match the last underscore
      Allele = "[^_]+"       # Match everything remaining that isn't an underscore
    )
  )
d <- as.data.table(d)

controls <- fread("./counts_report.txt")
sampleguide <- fread("./Samplemap.csv")
controls <- left_join(controls,sampleguide[,c(1,10)],join_by("Filename"=="FASTQ"))

setnames(controls,,c("Filename","PRE_mut","SRE_mut","Library"))
controls$Library <- str_replace_all(controls$Library, "((5hr|11hr)_[0-9]+)", "\\1_post")
controls <- controls[,2:4]
controls <- pivot_longer(controls,1:2,names_to = "Element", values_to = "Counts")
controls <- separate(controls,Library,c("Fraction","BiolRep","Splicing"))
controls$BiolRep <- as.numeric(controls$BiolRep)
d_wcontrols <- bind_rows(controls,d)
d_wcontrols <- d_wcontrols %>% group_by(Fraction,BiolRep,Splicing) %>% mutate(CPM=(Counts/sum(Counts))*1e6)
#Grab the DNA info
dna <- fread("../../../data/formatted/Library1and2BCLess_ISOMPRA_6-30-2022/formatted_Reads_compiled_06-30-2022.tsv") %>% filter(Fraction=="DNA")
dna <- dna %>% group_by(Element,Allele) %>% summarize(CPM=mean(CPM)) %>% as.data.table()
#Put the DNA info in the table, adjoining to correct element
d_wcontrols <- as.data.table(left_join(d_wcontrols,dna,by=c("Element","Allele")))
#Calculate expression
d_wcontrols[, Expression := log2(CPM.x/CPM.y)]
#Get rid of non-finite values
d_wcontrols <- filter(d_wcontrols,is.finite(Expression))

export_graph_start("../../../figure_making/Figures/Figure_S7/Boxplots for Truedecay_withSpikeIn.pdf",width=1.5,sideways = TRUE)
for (i in 1:nrow(truedecay)){
  # Separate these into ISOMPRA plots first, then ActD
  df <- as.data.table(filter(d_wcontrols,
                             Element == truedecay$Element[i],
                             Fraction == "0hr",
                             Allele %in% c("ref","alt")
                      ))
  df[,Allele := factor(Allele,levels = c("ref","alt"),
                         labels=c("Reference","Variant"))]
  df[,Splicing := factor(Splicing,levels = c("pre","post"),
                         labels=c("Pre-Splicing","Post-Splicing"))]
  p1 <- ggplot(df, aes(x = Splicing, y = Expression, color = Allele)) +
    # 1. Boxplots grouped by BOTH Allele and Splicing
    geom_boxplot(aes(group = interaction(Splicing, Allele)), 
                 position = position_dodge(width = 0.8), outlier.shape = NA) + 
    
    # 2. Points grouped by BOTH (to match the dodge)
    geom_point(aes(group = interaction(Splicing, Allele)), 
               position = position_dodge(width = 0.8), alpha = 0.5) +
    
    # 3. The Line grouped ONLY by Allele (to bridge the Splicing gap)
    stat_summary(aes(group = Allele), 
                 fun = median, geom = "line", 
                 position = position_dodge(width = 0.8), 
                 linewidth = 1) +
    
    labs(title = "ISOMPRA Profile",
         y = "log2(RNA/DNA)") +
    theme(plot.title = element_text(hjust=0.5))
  # Now the ActD
  df <- as.data.table(filter(d_wcontrols,
                             Element == truedecay$Element[i],
                             Splicing == "post",
                             Allele %in% c("ref","alt"),
                             !(Fraction=="11hr" & BiolRep==2),
                             !(Fraction=="0hr" & BiolRep==3 & Splicing=="post")
                             
                      ))
  df[,Allele := factor(Allele,levels = c("ref","alt"),
                         labels=c("Reference","Variant"))]
  df[,Time := factor(Fraction,levels = c("0hr","5hr","11hr"),
                         labels=c("t0","t5","t11"))]
  p2 <- ggplot(df, aes(x = Time, y = Expression, color = Allele)) +
    # 1. Boxplots grouped by BOTH Allele and Splicing
    geom_boxplot(aes(group = interaction(Time, Allele)), 
                 position = position_dodge(width = 0.8), outlier.shape = NA) + 
    
    # 2. Points grouped by BOTH (to match the dodge)
    geom_point(aes(group = interaction(Time, Allele)), 
               position = position_dodge(width = 0.8), alpha = 0.5) +
    
    # 3. The Line grouped ONLY by Allele (to bridge the Splicing gap)
    stat_summary(aes(group = Allele), 
                 fun = median, geom = "line", 
                 position = position_dodge(width = 0.8), 
                 linewidth = 1) +
    
    labs(title = "Actinomycin D Profile",
         y = "log2(RNA/DNA)") +
    theme(plot.title = element_text(hjust=0.5))
  
  combined_plot <- p1+p2
  
  print(combined_plot + 
    plot_annotation(
      title = truedecay$Symbol[i],
      subtitle = truedecay$Element[i],
      theme = theme(
        plot.title = element_text(hjust = 0.5, face = "bold", size = 24),
        plot.subtitle = element_text(hjust = 0.5, size = 18)
      )
    )
  )
  

}
export_graph_end()

```

## Investigating ActD Nominally Signficant Case Enrichment
```{r}
zero_to_five <- fread("./ActD_Zero-to-Five_RefvsAlt_LM.tsv")
zero_to_eleven <- fread("./ActD_Zero-to-Eleven_RefvsAlt_LM.tsv")

# 1. Get total counts of elements designed in your library per Phenotype group
# (Replace 'zero_to_five' with your actual data table variable name)
total_cases    <- length(zero_to_five[Pheno == "case"]$Element)
total_controls <- length(zero_to_five[Pheno == "control"]$Element)

# 2. Calculate the non-significant background
sig_cases   <- length(zero_to_five[Pheno == "case" & p_value<0.05]$Element)
sig_controls <- length(zero_to_five[Pheno == "control" & p_value<0.05]$Element)

nonsig_cases    <- total_cases - sig_cases
nonsig_controls <- total_controls - sig_controls

# 3. Build the 2x2 contingency table
contingency_matrix_ActD <- matrix(
  c(sig_cases, nonsig_cases,  # Cases (Sig, NonSig)
    sig_controls,  nonsig_controls), # Controls (Sig, NonSig)
  nrow = 2, byrow = TRUE,
  dimnames = list(Pheno = c("Case", "Control"),
                  Significance = c("Sig", "NonSig"))
)

fisher.test(contingency_matrix_ActD, alternative = "greater")

#Also rerun on pre- and post-splice
pre <- fread("./PreSplicing_RefAlt_LM.txt")
post <- fread("./PostSplicing_RefAlt_LM.txt")
stab <- fread("./Stability_RefAlt_LM.txt")
# 1. Get total counts of elements designed in your library per Phenotype group
# (Replace 'zero_to_five' with your actual data table variable name)
total_cases    <- length(pre[Pheno == "case"]$Element)
total_controls <- length(pre[Pheno == "control"]$Element)

# 2. Calculate the non-significant background
sig_cases   <- length(pre[Pheno == "case" & fdr<0.1]$Element)
sig_controls <- length(pre[Pheno == "control" & fdr<0.1]$Element)

nonsig_cases    <- total_cases - sig_cases
nonsig_controls <- total_controls - sig_controls

# 3. Build the 2x2 contingency table
contingency_matrix_pre <- matrix(
  c(sig_cases, nonsig_cases,  # Cases (Sig, NonSig)
    sig_controls,  nonsig_controls), # Controls (Sig, NonSig)
  nrow = 2, byrow = TRUE,
  dimnames = list(Pheno = c("Case", "Control"),
                  Significance = c("Sig", "NonSig"))
)

fisher.test(contingency_matrix_pre, alternative = "greater")

# 1. Get total counts of elements designed in your library per Phenotype group
# (Replace 'zero_to_five' with your actual data table variable name)
total_cases    <- length(post[Pheno == "case"]$Element)
total_controls <- length(post[Pheno == "control"]$Element)

# 2. Calculate the non-significant background
sig_cases   <- length(post[Pheno == "case" & fdr<0.1]$Element)
sig_controls <- length(post[Pheno == "control" & fdr<0.1]$Element)

nonsig_cases    <- total_cases - sig_cases
nonsig_controls <- total_controls - sig_controls

# 3. Build the 2x2 contingency table
contingency_matrix_post <- matrix(
  c(sig_cases, nonsig_cases,  # Cases (Sig, NonSig)
    sig_controls,  nonsig_controls), # Controls (Sig, NonSig)
  nrow = 2, byrow = TRUE,
  dimnames = list(Pheno = c("Case", "Control"),
                  Significance = c("Sig", "NonSig"))
)

fisher.test(contingency_matrix_post, alternative = "greater")

# 1. Get total counts of elements designed in your library per Phenotype group
# (Replace 'zero_to_five' with your actual data table variable name)
total_cases    <- length(stab[Pheno == "case"]$Element)
total_controls <- length(stab[Pheno == "control"]$Element)

# 2. Calculate the non-significant background
sig_cases   <- length(stab[Pheno == "case" & fdr_allele<0.1 & fdr_interaction>0.1]$Element)
sig_controls <- length(stab[Pheno == "control" & fdr_allele<0.1 & fdr_interaction>0.1]$Element)

nonsig_cases    <- total_cases - sig_cases
nonsig_controls <- total_controls - sig_controls

# 3. Build the 2x2 contingency table
contingency_matrix_stab <- matrix(
  c(sig_cases, nonsig_cases,  # Cases (Sig, NonSig)
    sig_controls,  nonsig_controls), # Controls (Sig, NonSig)
  nrow = 2, byrow = TRUE,
  dimnames = list(Pheno = c("Case", "Control"),
                  Significance = c("Sig", "NonSig"))
)

fisher.test(contingency_matrix_stab, alternative = "greater")


```

